---
title:  |
    | 
    | London Labour Market Update
date: "`r format(Sys.Date(),'%B %Y')`"
output:
  html_document:
    toc: true
    css: 'Y:\Labour Market Update R Project\MARKDOWN\GLAstyle.css'
# knit: (function(inputFile, encoding) {rmarkdown::render(inputFile,encoding=encoding,output_file=file.path(dirname(inputFile),"LMU.html")) })
---



```{r, echo=FALSE}

  ### Paths
wd <- setwd("Y:\\Labour Market Update R Project\\MARKDOWN")
INPUT <- paste0(wd,"/INPUT/")
INTERMEDIATE <- paste0(wd,"/INTERMEDIATE/")
OUTPUT <- paste0(wd,"/OUTPUT/")
MARKDOWN_IMAGES <- paste0(wd,"/IMAGES/")

  ### Header
   htmltools::img(src = knitr::image_uri(paste0(wd ,"/", "GLAEconomics.png")), 
                  alt = 'GLA Economics logo', 
                  style = 'position:absolute; align:right; top:0; right:0; width:300px; padding:10px;')
   


```

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE,scipen=999)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

#################################################################
###  Load packages and set paths
#################################################################

devtools::install_github("Greater-London-Authority/gglaplot")
#devtools::install_github("Greater-London-Authority/ldndatar", auth_token = "96e66bb601f49f62f0bb9bdcb73a849ece358ad1")
remotes::install_github("wilkelab/ggtext")

# install.packages("extrafont")
# install.packages('svglite')

library(extrafont)
library(remotes)
# remotes::install_version("Rttf2pt1", version = "1.3.8")
# extrafont::font_import()


library("knitr")
library("tidyverse")
library("ggplot2")
library("ggthemes")
library("nomisr")
library("devtools")
library("remotes")
library("scales")
library("gglaplot")
library("data.table")
library("janitor")
library("lubridate")
library("readr")
#library("ldndatar")
library("ggrepel")
library("plotly")
library("magrittr")
library("zoo")
library("kableExtra")
library("leaflet")
library("rgdal")

library(flextable)
library(officer) # for fp_text




### Geographic codes for UK, London and Boroughs.
### See "Borough codes.csv" in source folder for reference. 

Group <- c(2013265927,2092957697,1811939540,1811939541,1811939542,1811939543,1811939544,1811939526,1811939527,1811939545,1811939546,1811939547,
           1811939548,1811939528,1811939529,1811939530,1811939549,1811939550,1811939551,1811939552,1811939531,1811939532,1811939553,1811939533,
           1811939534,1811939554,1811939535,1811939555,1811939556,1811939536,1811939557,1811939537,1811939558,1811939538,1811939539)

################################################################################
### Presets for presenting figures etc.
################################################################################

signif_thous = 3 #How many significant figures to use when figures are in thousands
signif_mil = 2 # For millions
signif_perc = 2 #For percentages
signif_pp = 3 #For percentage points

digits_mil = 1
digits_perc = 1

################################################################################
### Functions
################################################################################

  ### Define function to always present percentages with one decimal and figures without decimals
perc_form = function(x, d=1) sprintf(paste0("%1.",d,"f"), x) 

value_form = function(x,s=2,d= -1) format(signif(round(as.numeric(x), d),s), big.mark=",")


  ### Function to make conditional text based on values
condi_text <- function(x,t=c("increase","noun","rise","growth","up","above")) { #Define names by the positive change
  if(t=="increase") {y = case_when(x<0 ~ "decreased",
                            x>0 ~ "increased")
                            return(y)}
  else if(t=="noun") {y = case_when(x<0 ~ "a decrease",
                            x>0 ~ "an increase")
                            return(y)}
  else if(t=="rise") {y = case_when(x<0 ~ "fell",
                            x>0 ~ "rose")
                            return(y)}
  else if(t=="growth") {y = case_when(x<0 ~ "shrank",
                            x>0 ~ "grew")
                            return(y)}
  else if(t=="up") {y = case_when(x<0 ~ "down",
                            x>0 ~ "up")
                            return(y)}
  else if(t=="above") {y = case_when(x<0 ~ "below",
                            x>0 ~ "above")
                            return(y)}
  else {return("[NB: wrong value for t]")}
}

### Function to remove negative sign from character strings (needed as formatted numbers are not numerical)
abs2 <- function(x) {
  y = gsub("-","",x)
  return(y)
}

################################################################################
### Downloads and data processing
################################################################################

  #.............................................................................
  ### London borough geographical codes
  #.............................................................................
  London_LA_codes <-readxl::read_excel(path = paste0(INPUT,"London borough codes.xlsx"), sheet = "Codes") %>% clean_names()

  #.............................................................................
  ### Fundction for dowloading headline stats
  #.............................................................................

HeadlineDownload <- function( data_series = "NM_59_1", 
                              time_period = c("1996-01","latest"), 
                              save_intermediate = TRUE,
                              geography = Group) {
  
  ### Download both absolute and percentage figures from NOMIS
  
  raw_nomis <- nomis_get_data( id = data_series, 
                               geography = geography,
                               time = time_period)
  
  ### Store downloads as dataframes
  
  lfs_stats <- raw_nomis %>% 
    mutate(DATE = as.Date(paste0(DATE, "-01"))) %>% 
    filter(MEASURES %in% c(20207,20100) & !is.na(OBS_VALUE)) %>%
    select( "DATE", "DATE_NAME", "SEX_NAME", "GEOGRAPHY_NAME", "ECONOMIC_ACTIVITY_NAME", "VALUE_TYPE_NAME", "OBS_VALUE","MEASURES","MEASURES_NAME") %>% 
     pivot_wider( names_from = ECONOMIC_ACTIVITY_NAME, values_from = OBS_VALUE) %>%  
    clean_names()

  ### Save intermediate dataframes if specified
  
  if( save_intermediate == TRUE) {
    lfs_stats %>% 
    fwrite(file = paste0(INTERMEDIATE,Sys.Date(),"_LFS_update.csv"))

    }
  
  data.table::fwrite( as.data.frame(colnames(lfs_stats)), file = paste0(INTERMEDIATE,"/",Sys.Date(),"_colnames.csv"))
  
  return( lfs_stats)
  
}

  #.............................................................................
  ### Download LFS stats, automatically clean and save
  #.............................................................................
  
lfs_stats <- HeadlineDownload( time_period = c("2020-01","latest"))



ClaimantCountDownload <- function( data_series = "NM_162_1",
                              time_period = c("1996-01","latest"),
                              save_intermediate = TRUE,
                              geography = Group,
                              line_or_bar = "line") {
  
  ### Download the claimant count data from NOMIS
  
  raw_nomis <- nomis_get_data(
                id = data_series, 
                geography = Group,
                time = time_period) 
  
  ### Clean data
  
  claimant_count_stats <- raw_nomis %>% 
    mutate( DATE = as.Date(paste0(DATE, "-01"))) %>% 
    filter( !is.na(OBS_VALUE) & MEASURES == 20100) %>% 
    pivot_wider( names_from = MEASURE_NAME, values_from = OBS_VALUE) %>%
    clean_names() %>% 
    select( date, date_name, geography_name, gender_name, age_name, last_col(4):last_col()) %>% 
    rename( sex_name = gender_name)
  
  if( line_or_bar == "line")
  {
    claimant_count_stats <- claimant_count_stats %>%
      filter( age_name == "All categories: Age 16+")
  }
  
  
  ### Save intermediate dataframes if specified
  
  if( save_intermediate == TRUE) {
    fwrite(claimant_count_stats, file = paste0(INTERMEDIATE,Sys.Date(),"_",data_series,"_claimant.csv"))
  }
  
 return( claimant_count_stats) 

}
  #.............................................................................
  ### Additional claimant count data
  #.............................................................................
  
  # NB: for now we need to use a manual downloaded file from Nomis as API is not working.

  claimant_count_stats <- ClaimantCountDownload( data_series = "NM_162_1", time_period = c("2019-12","latest"), save_intermediate = TRUE, line_or_bar = "bar")
  
 # raw_claim <- download.file( url = "https://www.nomisweb.co.uk/api/v01/dataset/NM_162_1.data.csv?geography=1807745125...1807745129,1807745111,1807745112,1807745130...1807745133,1807745113...1807745115,1807745134...1807745137,1807745116,1807745117,1807745138,1807745118,1807745119,1807745139,1807745120,1807745140,1807745141,1807745121,1807745142,1807745122,1807745143,1807745123,1807745124,2092957697,2013265927&date=latestMINUS22-latest&gender=0...2&age=0...2,10,11,3,12...16,4,17...20&measure=1...4&measures=20100&signature=NPK-ffca5e5a461529986a4082:0x6a7726c4d1c34de0ce2aca1289257fd9ef1cc613"), mode = "wb")

# claimant_count_stats_o <-read.csv(paste0(INTERMEDIATE,"2021-11-16_NM_162_1_claimant.csv")) %>% 
#     arrange(date, geography_name,sex_name,age_name)

  # claimant_count_stats <-read.csv(paste0(INPUT,"2021-12-10_claims_NM_162_1.csv")) %>% 
  #   mutate( DATE = as.Date(paste0(DATE, "-01"))) %>% 
  #   filter( !is.na(OBS_VALUE) & MEASURES == 20100) %>% 
  #   pivot_wider( names_from = MEASURE_NAME, values_from = OBS_VALUE) %>%
  #   clean_names() %>% 
  #   select( date, date_name, geography_name, gender_name, age_name, last_col(4):last_col()) %>% 
  #   rename( sex_name = gender_name) %>% 
  #   arrange(date, geography_name,sex_name,age_name)

  #.............................................................................
  ### Download PAYE data, clean and save
  #.............................................................................
  
  raw_paye <- download.file( url = "https://www.ons.gov.uk/file?uri=%2femploymentandlabourmarket%2fpeopleinwork%2fearningsandworkinghours%2fdatasets%2frealtimeinformationstatisticsreferencetableseasonallyadjusted%2fcurrent/rtisadec2021.xlsx",
                             destfile = paste0(INPUT,Sys.Date(),"_raw_paye.xlsx"),
                             mode = "wb")
  
  paye_stats <- readxl::read_excel(path = paste0(INPUT,Sys.Date(),"_raw_paye.xlsx"), sheet = "7. Employees (NUTS1)", skip = 6) %>% 
                  select( Date, London, UK) %>%
                  clean_names() %>% 
                  mutate( new_date = dmy(paste0("01",date)),
                          lon_change_since_feb20 = case_when( new_date <= as.Date("2020-02-01") ~ NaN,
                                                        new_date > as.Date("2020-02-01") ~ london - london[ new_date == "2020-02-01"]),
                          lon_percentage_change_since_feb20 = case_when( new_date <= as.Date("2020-02-01") ~ NaN,
                                                        new_date > as.Date("2020-02-01") ~ 100*(london - london[ new_date == "2020-02-01"])/london[ new_date == "2020-02-01"]),
                          lon_month_change = case_when( new_date - min(new_date) < 365 ~ NaN,
                                                                    new_date - min(new_date) > 365 ~ 
                                                                      (london - lag(london, n = 1))),
                          lon_month_percentage_change = case_when( new_date - min(new_date) < 365 ~ NaN,
                                                                    new_date - min(new_date) > 365 ~ 
                                                                      100*(london - lag(london, n = 1))/lag(london, n = 1)),
                          uk_change_since_feb20 = case_when( new_date <= as.Date("2020-02-01") ~ NaN,
                                                           new_date > as.Date("2020-02-01") ~ uk - uk[ new_date == "2020-02-01"]),
                          uk_percentage_change_since_feb20 = case_when( new_date <= as.Date("2020-02-01") ~ NaN,
                                                           new_date > as.Date("2020-02-01") ~ 100*(uk - uk[ new_date == "2020-02-01"])/uk[ new_date == "2020-02-01"]),
                          lon_annual_change = case_when( new_date - min(new_date) < 365 ~ NaN,
                                                                    new_date - min(new_date) > 365 ~ 
                                                                      (london - lag(london, n = 12))),
                          lon_annual_percentage_change = case_when( new_date - min(new_date) < 365 ~ NaN,
                                                                    new_date - min(new_date) > 365 ~ 
                                                                      100*(london - lag(london, n = 12))/lag(london, n = 12)),
                          uk_annual_percentage_change = case_when( new_date - min(new_date) < 365 ~ NaN,
                                                                    new_date - min(new_date) > 365 ~ 
                                                                      100*(uk - lag(uk, n = 12))/lag(uk, n = 12)))
  
  paye_nuts2_stats <- readxl::read_excel(path = paste0(INPUT,Sys.Date(),"_raw_paye.xlsx"), sheet = "11. Employees (NUTS2)", skip = 6) %>% 
        mutate( new_date = dmy(paste0("01",Date)))  %>% 
        merge(paye_stats,by="new_date")   %>% 
    rename(London=london) %>% 
    select( Date, new_date, "Inner London - West",	"Inner London - East",	"Outer London - East and North East",	"Outer London - South",	"Outer London - West and North West", "London","UK") %>%
    pivot_longer(cols="Inner London - West":UK,names_to = "area", values_to = "emp_level") %>% 
    group_by(area) %>% 
    mutate( 
      change_perc_since_feb20 = case_when( new_date <= as.Date("2020-02-01") ~ NaN,
                                      new_date > as.Date("2020-02-01") ~ 100*(emp_level - emp_level[ new_date == "2020-02-01"])/(emp_level[ new_date == "2020-02-01"])),
      annual_change = case_when(new_date - min(new_date) < 365 ~ NaN,
                                      new_date - min(new_date) > 365 ~ 100*(emp_level - lag(emp_level, n= 12))/lag(emp_level, n=12))) %>% 
    clean_names()
  
  #fwrite( paye_stats, file = paste0(INTERMEDIATE,Sys.Date(),"_paye.csv"), na = "NaN")
  
  #### Each quarter, process data on PAYE at borough level and merge with geographical codes. 
  #### NB: MUST BE UPDATED MANUALLY TO NEWER DATA

  paye_la_stats <- readxl::read_excel(path = paste0(INPUT,"2021-12-14_raw_paye.xlsx"), sheet = "19. Employees (LA)", skip = 7) %>% 
    mutate(new_date = dmy(paste0("01",Date))) %>% 
    select( Date, new_date, "City of London", "Barking and Dagenham", "Barnet", "Bexley", "Brent", "Bromley", "Camden", "Croydon", "Ealing", "Enfield", "Greenwich", "Hackney", "Hammersmith and Fulham", "Haringey", "Harrow", "Havering", "Hillingdon", "Hounslow", "Islington", "Kensington and Chelsea", "Kingston upon Thames", "Lambeth", "Lewisham", "Merton", "Newham", "Redbridge", "Richmond upon Thames", "Southwark", "Sutton", "Tower Hamlets", "Waltham Forest", "Wandsworth", "Westminster") %>% 
    pivot_longer(cols="City of London":"Westminster",names_to = "area", values_to = "emp_level") %>% 
    group_by(area) %>% 
    mutate( 
      change_perc_since_feb20 = case_when( new_date <= as.Date("2020-02-01") ~ NaN,
                                           new_date > as.Date("2020-02-01") ~ 100*(emp_level - emp_level[ new_date == "2020-02-01"])/(emp_level[ new_date == "2020-02-01"])),
      annual_change = case_when(new_date - min(new_date) < 365 ~ NaN,
                                new_date - min(new_date) > 365 ~ 100*(emp_level - lag(emp_level, n= 12))/lag(emp_level, n=12))) %>% 
    clean_names() %>%  merge(London_LA_codes,by='area') %>% 
    arrange(area,new_date)
  
  
  #.............................................................................
  ### Download Workforce Jobs data
  #.............................................................................
  
  wfj_stats <- nomis_get_data(id = "NM_130_1", time = c("2010-01", "latest"), geography = Group,
                              industry=c(37748736,150994945:150994964)) %>% 
    clean_names()   %>% 
    select(c("date","date_name","geography","geography_name","industry","industry_code","industry_name","item","item_name","measures","measures_name","obs_value","record_count")) %>% 
    mutate(new_date = dmy(paste0("01",date_name)),
           quarter =  paste0("Q",quarter(new_date)),
           quarter_text = case_when(quarter=="Q1" ~ "first",
                                  quarter=="Q2" ~ "second",
                                  quarter=="Q3" ~ "third",
                                  quarter=="Q4" ~ "fourth")) %>% 
    group_by(industry_name, geography_name, item_name,measures_name) %>% #grouping makes calculating lag values easier
    mutate( 
      change_perc_since_dec19 = case_when( new_date <= as.Date("2019-12-01") | measures_name == "Percent" ~ NaN,
                                      new_date > as.Date("2019-12-01") ~ 100*(obs_value - obs_value[ new_date == "2019-12-01"])/(obs_value[ new_date == "2019-12-01"])),
      change_since_dec19 = case_when( new_date <= as.Date("2019-12-01") | measures_name == "Percent" ~ NaN,
                                      new_date > as.Date("2019-12-01") ~ (obs_value - obs_value[ new_date == "2019-12-01"])),
      change_perc_quar = case_when(  measures_name == "Percent" ~ NaN,
                                    TRUE ~ 100*(obs_value - lag(obs_value, n = 1))/(lag(obs_value, n = 1))),
      change_quar = case_when(  measures_name == "Percent" ~ NaN,
                                    TRUE ~ (obs_value - lag(obs_value, n = 1))),
      industry_name = case_when(industry_name == "Total" ~ "Total",
                     TRUE ~  str_sub(industry_name,5,-1))) %>% #trims the name 
    mutate( industry_name_simple = case_when(
      industry_name == "Public administration and defence; compulsory social security" ~ "Public admin & defence",
      industry_name == "Human health and social work activities" ~ "Health",
      industry_name == "Financial and insurance activities" ~ "Finance & insurance",
      industry_name == "Professional, scientific and technical activities" ~ "Professional services",
      industry_name == "Wholesale and retail trade; repair of motor vehicles and motorcycles" ~ "Retail",
      industry_name == "Other service activities" ~ "Other services",
      industry_name == "Transportation and storage" ~ "Transport & storage",
      industry_name == "Arts, entertainment and recreation" ~ "Arts & recreation",
      industry_name == "Information and communication" ~ "Information & communication",
      industry_name == "Administrative and support service activities" ~ "Administration",
      industry_name == "Accommodation and food service activities" ~ "Hospitality",
      industry_name == "Water supply; sewerage, waste management and remediation activities" ~ "Water",
      TRUE ~ industry_name)) %>% 
    group_by( geography_name,new_date,item_name) %>% # To get ranks of industries within date-geography groups
    mutate(perc_change_dec19_within_date_rank = case_when( industry_name == "Total" ~ NA_integer_, 
                                         TRUE  ~ dense_rank(desc(change_perc_since_dec19))), 
            perc_change_quart_within_date_rank = case_when( industry_name == "Total" ~ NA_integer_, 
                                         TRUE  ~ dense_rank(desc(change_perc_quar)))) %>% 
    ungroup() #ungroup to ease future manipulations
    
  wfj_stats<-wfj_stats %>% arrange(date,geography,item,measures,industry)
  
  #wfj_stats <- clean_names(wfj_stats,case="snake") #Converting column names to lower

  fwrite(wfj_stats, file = paste0(INTERMEDIATE,Sys.Date(),"_wfj.csv"), na = "NaN")
  
  #-----------------------------------------------------------------------------
  ### Define GLA line chart function
  #-----------------------------------------------------------------------------
  
  GLALineChart <- function( data_set = lfs_stats, ### This function has been written to work specifically with the NM_59_1 dataset
                        lfs_or_claims = "lfs",
                        x_var = date, ### Assumed that this will be a line chart with dates on the x-axis
                        y_var = NULL, ### See document "[today's date]colnames.csv" for variables available for charting
                        geography = c("London", "United Kingdom"),
                        suffix = "%",
                        y_limits = c(0,100),
                        nudge_y = NULL,
                        title = NULL,
                        subtitle = paste0("Latest data for period ", new_release),
                        caption = "",
                        chart_name = NULL) {
      
    pal <- gla_pal(gla_theme = "default", palette_type = "highlight", n = c(1, 1))
    theme_set(theme_gla(gla_theme = "default"))
    
    ### Create dataset for charting
    
    if( lfs_or_claims == "lfs") {
    
    for_charting <- data_set %>% 
      filter( geography_name %in% geography & value_type_name == "Level" & sex_name == "Total" & measures == 20207)
    } else {
    for_charting <- data_set %>% 
      filter( geography_name %in% geography & sex_name == "Total" & !is.na({{y_var}}))
    }
    ### Set x-axis format
  
    ### Plot charts
    
    new_release <- for_charting$date_name[for_charting$date == max(for_charting$date)]
    
    glaplot <- for_charting %>%
      ggplot(mapping = aes(x = {{x_var}}, y = {{y_var}}, 
                           colour = geography_name, group = geography_name,
                           text = paste(
                          geography_name, "\n",
                          format({{x_var}},'%B %Y'), "\n",
                          "Rate: ", perc_form({{y_var}}),"%", "\n",
                          sep = ""))) +
      ggla_line(aes(size= geography_name)) +
      scale_size_manual(values = c(4 * mm_to_pt, 2 * mm_to_pt)) +
      scale_colour_manual(values = pal) +
      ggla_highlight(filter_type = "end") +
      ggla_highlight(mapping = aes(label = paste0({{y_var}}, "%")),
                     geom = GeomGLATextHighlight, filter_type = "end",  size = 4.5,
                     position = position_nudge(y = nudge_y),check_overlap = TRUE)+
      coord_cartesian(clip = 'off') +
      scale_y_continuous(expand = c(0, 0), labels = dollar_format(prefix = "", 
                                                                  suffix = suffix, 
                                                                  largest_with_cents = 1), 
                                                                  limits = y_limits) +
      scale_x_date( date_breaks = "2 months",
                    date_labels = "%b'%y",
                    expand = expansion( mult = c(0.1,0.02))) +
      theme(plot.margin = unit(c(1,1,1,1), "cm"))+
      labs(title = title,
           subtitle = subtitle,
           caption = caption) +
      theme(plot.caption = element_text(color = rgb(166,166,166,maxColorValue = 255)))
    
    ggsave( paste0(MARKDOWN_IMAGES,Sys.Date(),"_",chart_name,".svg"), device = "svg", width = 8, height = 8, units = "in")
    ggsave( paste0(MARKDOWN_IMAGES,Sys.Date(),"_",chart_name,".png"), device = "png", width = 8, height = 8, units = "in")
    
    return(glaplot)
    
  }

```

``` {r, main_figures ,echo=FALSE}
  ### Create all the figures needed for the section below
  # Name format:
  # _d_: difference, _p_: difference in percent, _pp_: difference in percentage points, _uk: UK level stat, London otherwise

# Paye
paye_d_month <- value_form(paye_stats %>% filter( new_date == max(new_date)) %>% select(lon_month_change),s= signif_thous)

paye_p_month <- perc_form(paye_stats %>% filter( new_date == max(new_date)) %>% select(lon_month_percentage_change),d = digits_perc)

paye_newmonth <- paye_stats %>% filter( new_date == max(new_date)) %>% mutate(new_date=format(new_date,'%B %Y')) %>%  select(new_date)

paye_lastmonth <- paye_stats  %>% mutate(lag_month = lag(new_date, n = 1)) %>% mutate(lag_month2 = format(lag_month, '%B')) %>% filter( new_date == max(new_date)) %>%  select(lag_month2)

paye_p_feb20 <- perc_form(paye_stats %>% filter( new_date == max(new_date)) %>% select(lon_percentage_change_since_feb20),d = digits_perc)

# LFS

lfs_newmonth <- lfs_stats %>% mutate(month = format(date, '%B %Y')) %>% filter(sex_name == "Total"  & geography_name == "London" & value_type_name == "Level" & date == max(date) & measures == 20207) %>%  select(month)
  
lfs_lastmonth <- lfs_stats  %>% mutate(lag_month = ymd(date) %m+% - months(2)) %>% mutate(lag_month2 = format(lag_month, '%B')) %>% filter(sex_name == "Total"  & geography_name == "London" & value_type_name == "Level" & date == max(date) & measures == 20207) %>%  select(lag_month2)


# Unem

unem_rate <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London" & value_type_name == "Level" & measures == 20207) %>% select(total_unemployed_aged_16_and_over),d = digits_perc)
  
unem_pp_quart <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London" & value_type_name == "Change on quarter" & measures == 20207) %>% select(total_unemployed_aged_16_and_over),d = digits_perc)


unem_pp_year <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London" & value_type_name == "Change on year" & measures == 20207)  %>% select(total_unemployed_aged_16_and_over),d = digits_perc)
  
unem_rate_uk <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom" & value_type_name == "Level" & measures == 20207) %>% select(total_unemployed_aged_16_and_over),d = digits_perc)


# Employment

emp_rate <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London" & value_type_name == "Level" & measures == 20207)  %>% select(total_in_employment_aged_16_to_64),d = digits_perc)

emp_pp_quart <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London"  & value_type_name == "Change on quarter" & measures == 20207)  %>% select(total_in_employment_aged_16_to_64),d = digits_perc)

emp_pp_year <-perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London"  & value_type_name == "Change on year" & measures == 20207) %>% select(total_in_employment_aged_16_to_64),d = digits_perc)


# WFJ

wfj_year <-  wfj_stats %>% filter(new_date == max(new_date)) %>% filter(row_number()==1) %>% mutate(year=year(new_date)) %>% select(year)

wfj_newquarter <- wfj_stats %>% ungroup()  %>% filter(new_date == max(new_date)) %>% filter(row_number()==1) %>% select(quarter_text) 
  
wfj_oldquarter <-  wfj_stats %>% filter(new_date == (ymd(max(new_date)) %m+% - months(3))) %>% filter(row_number()==1) %>% select(quarter_text)

wfj_d_quart <- value_form(wfj_stats%>% ungroup()  %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd(max(new_date)) %m+% - months(3)) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value),s = signif_thous)

wfj_p_quart <- perc_form(100*(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd(max(new_date)) %m+% - months(3)) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value))/(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd(max(new_date)) %m+% - months(3)) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value)),d = digits_perc)

wfj_d_dec19 <- value_form(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd("2019-12-01")) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value),s = signif_thous)

wfj_p_dec19 <- perc_form(100*(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd("2019-12-01")) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value))/(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd("2019-12-01")) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value)),d = digits_perc)

```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>
<!-- Summary of findings - MANUALLY ADJUST -->

<br/>

## Summary of key points

<br/>

* __Payrolled employees__ `r condi_text(paye_d_month,"rise")` by around `r abs2(paye_d_month)`,  or `r abs2(paye_p_month)`%, between `r paye_lastmonth` and `r paye_newmonth` and is `r abs2(paye_p_feb20)`% `r condi_text(paye_p_feb20,"above")` pre-pandemic (February 2020) levels. This was both the largest monthly percentage increase and the highest level of employees on record (since July 2014).
* __Unemployment rate__ estimated at `r unem_rate`% for the three months ending `r lfs_newmonth`, `r condi_text(unem_pp_quart,"up")` `r abs2(unem_pp_quart)` percentage points (pp)
on the quarter and `r condi_text(unem_pp_year,"up")` `r abs2(unem_pp_year)`pp on the same period in the previous year (while the UK average was `r unem_rate_uk`%). 
* __Employment rate__ estimated at `r emp_rate`% for the three months ending `r lfs_newmonth`, `r condi_text(emp_pp_quart,"up")` `r abs2(emp_pp_quart)`pp on the previous quarter and `r condi_text(emp_pp_year,"up")` `r abs2(emp_pp_year)`pp from a year earlier.
* __Workforce jobs__ `r condi_text(wfj_d_quart,"increase")` by `r abs2(wfj_d_quart)` or `r abs2(wfj_p_quart)`% between the `r wfj_oldquarter` and `r  wfj_newquarter` quarter of `r wfj_year`, and is `r abs2(wfj_d_dec19)`, or `r abs2(wfj_p_dec19)`%, `r condi_text(wfj_d_dec19,"above")` pre-pandemic (December 2019) levels. While employee jobs have recovered, self-employed jobs remain 16.7% below pre-pandemic levels.

<a href="#top">Back to top</a>

<!-- Standard intro page with caveats -->

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

## Note on end of furlough

<br/>

According to official HMRC statistics, there were still around 231,000 London staff on furlough at the [end of September 2021](https://data.london.gov.uk/download/gla-economics-covid-19-labour-market-analysis/a1fdeb85-1918-4b45-a384-6e0cad70b469/Official%20Furlough%20stats%20November%202021%20-%20London%20Datastore.pdf), with a higher furlough take-up rate in the capital when the scheme closed.

The latest payrolled employee and claimant count estimates for November 2021 provide an early indication of the
impact of the end of the scheme on the labour market in London.*

Both measures show relatively steady trends in the latest month, with a significant increase in the number of employees
paid through payroll and a small decline in the number of people claiming unemployment related benefits.

These latest estimates may not yet reflect post-furlough redundancies. For example, according to the [ONS](https://www.ons.gov.uk/employmentandlabourmarket/peopleinwork/employmentandemployeetypes/bulletins/uklabourmarket/latest):
<div style="margin-left: 1em;">
‘It is possible those made redundant at the end of the furlough scheme will be included in the RTI data for a few further months, while they work out their notice period.’
</div>

However, responses to a recent [ONS business survey](https://www.ons.gov.uk/businessindustryandtrade/business/businessservices/bulletins/businessinsightsandimpactontheukeconomy/4november2021) suggest that only a small share of furloughed employees
were made redundant recently. Early indications suggest this may also be true for London.

<a href="#top">Back to top</a>

<foot>*Note: the latest payrolled employees estimate is based on incomplete information and may be subject to revision. </foot>

<!-- Standard intro page with caveats -->

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

## Measuring the labour market

<br/>

The ONS has published its latest [labour market update](https://www.ons.gov.uk/employmentandlabourmarket/peopleinwork/employmentandemployeetypes) covering Labour Force Survey (LFS) data for the three months ending `r lfs_newmonth`. This bulletin presents the latest headline labour market estimates for London.    
 
Please note:    
 
* Many of the statistics presented here are estimates based on surveys, and as such have a margin of error – known as sampling variability.  
* They also cover different reference periods or count dates, and have therefore been impacted differently by the spread of COVID-19.   
 
More information and previous GLA Economics analyses can be found on our COVID-19 labour market analysis [page](https://data.london.gov.uk/dataset/gla-economics-covid-19-labour-market-analysis).   

More information on the data used and a glossary of terms can be found on the [ONS website](https://www.ons.gov.uk/employmentandlabourmarket/peopleinwork/employmentandemployeetypes/methodologies/aguidetolabourmarketstatistics).

``` {r,paye figures, echo=FALSE}

paye_level <- value_form(paye_stats %>% filter( new_date == max(new_date)) %>% mutate(lon_round=london/1000000) %>% select(lon_round),s=signif_mil, d=digits_mil)

paye_d_year <- value_form(paye_stats %>% filter( new_date == max(new_date))  %>% select(lon_annual_change),s=signif_thous)
  
paye_p_year <- perc_form(paye_stats %>% filter( new_date == max(new_date)) %>% select(lon_annual_percentage_change),d = digits_perc)

paye_lastyearmonth <- paye_stats  %>% mutate(lag_month = lag(new_date, n = 12)) %>% mutate(lag_month2 = format(lag_month, '%B %Y')) %>% filter( new_date == max(new_date)) %>%  select(lag_month2)
  
paye_p_year_uk <- perc_form(paye_stats %>% filter( new_date == max(new_date)) %>% select(uk_annual_percentage_change),d = digits_perc)

paye_d_feb20 <- value_form(paye_stats %>% filter( new_date == max(new_date))  %>% select(lon_change_since_feb20),s=signif_thous)

paye_p_feb20 <- perc_form(paye_stats %>% filter( new_date == max(new_date)) %>% select(lon_percentage_change_since_feb20),d = digits_perc)

paye_d_feb20_uk <- value_form(paye_stats %>% filter( new_date == max(new_date))  %>% select(uk_change_since_feb20),s=signif_thous)

paye_p_feb20_uk <- perc_form(paye_stats %>% filter( new_date == max(new_date)) %>% select(uk_percentage_change_since_feb20),d = digits_perc)

```

<a href="#top">Back to top</a>

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

## Payrolled employees {.tabset}

<br/>

The count of payrolled employees from HMRC’s Pay As You Earn (PAYE) RTI dataset offers a timely measure of labour market trends:

* Early estimates indicate that there were around `r paye_level` million payrolled employees living in London in  `r paye_newmonth`, `r condi_text(paye_d_month,"noun")` of around `r abs2(paye_d_month)`  or `r abs2(paye_p_month)`% on the previous month. 

* The latest London estimate represents `r condi_text(paye_d_year,"noun")` of `r abs2(paye_d_year)` or `r abs2(paye_p_year)`% on the previous year (`r paye_lastyearmonth`) against a UK average increase of `r paye_p_year_uk`% (see charts).

* Relative to pre-pandemic levels (February 2020), the number of payrolled employees in London is `r condi_text(paye_d_feb20,"up")` by `r abs2(paye_d_feb20)` or `r abs2(paye_p_feb20)`%. This compares to `r condi_text(paye_p_feb20_uk,"noun")` of `r abs2(paye_p_feb20_uk)`%  across the UK on average.

<a href="#top">Back to top</a>

### Change on previous year

```{r PAYE chart, echo=FALSE,fig.cap = payecap, out.width='100%' }

pal <- gla_pal(gla_theme = "default", palette_type = "highlight", n = c(1, 1))
theme_set(theme_gla(gla_theme = "default"))

paye_stats_chart <- paye_stats %>%
                    pivot_longer(cols = ends_with("_change"), names_to = "geography", values_to = "change" ) %>%
                    mutate( geography_name = case_when( grepl("lon",geography) ~ "London",
                                                        grepl("uk",geography) ~ "United Kingdom")) %>%
                    mutate(var_type = sub(".*?_",'',geography)) %>% 
                                      filter( new_date >= "2020-01-01" & var_type == "annual_percentage_change") %>%
                    ggplot(mapping = aes(x = new_date, y = change,
                                          colour = geography_name, group = geography_name,
                           text = paste(
                          geography_name, "\n",
                          format(new_date,'%B %Y'), "\n",
                          "Change: ", perc_form(change),"%", "\n",
                          sep = ""))) +
                    ggla_line(aes(size= geography_name)) +
                    scale_size_manual(values = c(4 * mm_to_pt, 2 * mm_to_pt)) +
                    scale_colour_manual(values = pal) +
                   # ggla_highlight(filter_type = "end") +
                    # ggla_highlight(mapping = aes(label = paste0(format(change, digits = 1),"%")),
                    #                geom = GeomGLATextHighlight, filter_type = "end",  size = 4.5,
                    #                position = position_nudge(y = c(-1.3,+0.5)))+
                    coord_cartesian(clip = 'off') +
                    geom_hline(aes(yintercept=0), colour="gray45") +
                    scale_y_continuous(expand = c(0, 0), labels = dollar_format(prefix = "",
                                                                                suffix = "%",
                                                                                largest_with_cents = 1),
                                       limits = c(-6,7)) +
                    scale_x_date( date_breaks = "2 months",
                                  date_labels = "%b'%y",
                                  expand = expansion( mult = c(0.1,0.02))) +
                    theme(plot.margin = unit(c(1,1,1,1), "cm"))+
                    labs(title =  "Payrolled employees, change on previous year",
                         subtitle = paste0("Latest data for period ", paye_stats$date[paye_stats$new_date == max(paye_stats$new_date)]),
                         caption = "\nSource: HM Revenue and Customs – Pay As You Earn Real Time Information.\n\nNote: Estimates are based on where employees live.")+
                    theme(plot.caption = element_text(color = rgb(166,166,166,maxColorValue = 255)))

ggsave( paste0(MARKDOWN_IMAGES,Sys.Date(),"_","paye",".svg"), device = "svg", width = 8, height = 8, units = "in")
ggsave( paste0(MARKDOWN_IMAGES,Sys.Date(),"_","paye",".png"), device = "png", width = 8, height = 8, units = "in")

                     
#knitr::include_graphics(paste0(MARKDOWN_IMAGES,Sys.Date(),"_","paye",".svg"))

ggplotly(paye_stats_chart,tooltip = "text") %>% 
  ggla_plotly_settings()    %>% 
  layout(title = list(text = paste0("<b>","Payrolled employees, change on previous year","</b>",
                                    "<br>",
                                    "<sup>",
                                    paste0("Latest data for period ", paye_stats$date[paye_stats$new_date == max(paye_stats$new_date)]),
                                    "</sup>",
                                    "<br>"),
         font = list(size = 22),
         y = .95, xref = "plot"),
         xaxis = list(tickfont = list(size = 15)),
         yaxis = list(tickfont = list(size = 15)),
         legend = list(font = list(size = 15)),
         hovermode = "x") 

payecap <- paste0("Source: HM Revenue and Customs – Pay As You Earn Real Time Information.","<br>","<br>", "Note: Estimates are based on where employees live.")

```


<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

### Change since February 2020, NUTS2 regions

```{r paye bar feb20, , echo=FALSE,fig.cap = payebarcapfeb, out.width='100%'}


pal <- gla_pal(gla_theme = "default", palette_type = "highlight", n = c(1, 1))
theme_set(theme_gla(gla_theme = "default"))

paye_stats_barchart_feb20 <- paye_nuts2_stats %>%
  filter( new_date == max(new_date)) %>%
  ungroup() %>% 
  mutate(chart_ranking = case_when( area == "UK" ~ 1,
                                    area == "London" ~ 2,
                                    TRUE ~ dense_rank(desc(change_perc_since_feb20))+2)) %>% 
  ggplot(mapping = aes(x =  reorder(area, -chart_ranking), y = change_perc_since_feb20,
                       colour = area, group = area, fill=area,
                       text = paste(
                         area, "\n",
                         "Change: ", perc_form(change_perc_since_feb20),"%", "\n",
                         sep = ""))) +
  geom_bar(stat="identity",width=0.5)+
  geom_text(aes(label=paste(format(round(change_perc_since_feb20,1), digits = 2), "%"), y = change_perc_since_feb20-0.2*((change_perc_since_feb20)/abs(change_perc_since_feb20))),color=c("white","white","white","white","white","white","black"))+ #cannot use vjust with ggplotly
  scale_color_manual(values = c(pal[1],pal[1],pal[1],pal[1],pal[1],pal[1],pal[2]), aesthetics = "colour") +
  scale_fill_manual(values = c(pal[1],pal[1],pal[1],pal[1],pal[1],pal[1],pal[2]), aesthetics =  "fill") +
  coord_cartesian(clip = 'off') +
  geom_hline(aes(yintercept=0), colour="gray45") +
  scale_y_continuous(labels = dollar_format(prefix = "",
                                            suffix = "%",
                                            largest_with_cents = 1),
                     limits = c(-2,3)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  
  theme(plot.margin = unit(c(1,1,1,1), "cm"),axis.text=element_text(size=12))+
  labs(title =  "Payrolled employees, change since February 2020 by NUTS2 region",
       subtitle = paste0("Latest data for period ", paye_nuts2_stats$date[paye_nuts2_stats$new_date == max(paye_nuts2_stats$new_date)]),
       caption = "") + 
  theme(legend.position = "none")

ggsave( paste0(MARKDOWN_IMAGES,Sys.Date(),"_","paye_nuts2",".svg"), device = "svg", width = 16, height = 8, units = "in")
ggsave( paste0(MARKDOWN_IMAGES,Sys.Date(),"_","paye_nuts2",".png"), device = "png", width = 16, height = 8, units = "in")


#knitr::include_graphics(paste0(MARKDOWN_IMAGES,Sys.Date(),"_","paye_nuts2",".png"))



ggplotly(paye_stats_barchart_feb20,tooltip = "text") %>% 
  ggla_plotly_settings(legend=FALSE)   %>%
  layout(title = list(text = paste0("<b>","Payrolled employees, change since February 2020 by NUTS2 region","</b>",
                                    "<br>",
                                    "<sup>",
                                    paste0("Latest data for period ", paye_nuts2_stats$date[paye_nuts2_stats$new_date == max(paye_nuts2_stats$new_date)]),
                                    "</sup>",
                                    "<br>"),
                      font = list(size = 22),
                      y = .95, xref = "plot"),
         xaxis = list(tickfont = list(size = 15)),
         yaxis = list(tickfont = list(size = 15)),
         legend = list(font = list(size = 15))) 

payebarcapfeb <- paste0("Source: HM Revenue and Customs – Pay As You Earn Real Time Information.","<br>","<br>", "Note: Estimates are based on where employees live.")

```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

### Change on previous year, NUTS2 regions
```{r paye bar, , echo=FALSE,fig.cap = payebarcap, out.width='100%' }


pal <- gla_pal(gla_theme = "default", palette_type = "highlight", n = c(1, 1))
theme_set(theme_gla(gla_theme = "default"))

paye_stats_barchart <- paye_nuts2_stats %>%
  filter(new_date == max(new_date)) %>%
  ungroup() %>% 
  mutate(chart_ranking = case_when( area == "UK" ~ 1,
                                    area == "London" ~ 2,
                                    TRUE ~ dense_rank(desc(annual_change))+2)) %>% 
  ggplot(mapping = aes(x =  reorder(area, -chart_ranking), y = annual_change,
                       colour = area, group = area, fill=area,
         text = paste(
           area, "\n",
           "Change: ", perc_form(annual_change),"%", "\n",
           sep = ""))) +
  geom_text(aes(label=paste0(format(round(annual_change,1), digits = 2), "%"), y = annual_change-0.2*min(annual_change)), color=c("white","white","white","white","white","white","black"))+ #cannot use vjust with ggplotly
  geom_bar(stat="identity",width=0.5)+
  scale_color_manual(values = c(pal[1],pal[1],pal[1],pal[1],pal[1],pal[1],pal[2]), aesthetics = "colour") +
  scale_fill_manual(values = c(pal[1],pal[1],pal[1],pal[1],pal[1],pal[1],pal[2]), aesthetics = "fill") +
  scale_size_manual(values = c(4 * mm_to_pt, 2 * mm_to_pt)) +
  coord_cartesian(clip = 'off') +
  geom_hline(aes(yintercept=0), colour="gray45") +
  scale_y_continuous(labels = dollar_format(prefix = "",
                                            suffix = "%",
                                            largest_with_cents = 1),
                     limits = c(0,8)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  
  theme(plot.margin = unit(c(1,1,1,1), "cm"),axis.text=element_text(size=12))+
  labs(title =  "Payrolled employees, change on previous year by NUTS2 region",
       subtitle = paste0("Latest data for period ", paye_nuts2_stats$date[paye_nuts2_stats$new_date == max(paye_nuts2_stats$new_date)]),
       caption = "\nSource: HM Revenue and Customs – Pay As You Earn Real Time Information.\n\nNote: Estimates are based on where employees live.") + 
  theme(legend.position = "none") +
  theme(plot.caption = element_text(color = rgb(166,166,166,maxColorValue = 255)))

ggsave( paste0(MARKDOWN_IMAGES,Sys.Date(),"_","paye_nuts2_V2",".svg"), device = "svg", width = 8, height = 8, units = "in")
ggsave( paste0(MARKDOWN_IMAGES,Sys.Date(),"_","paye_nuts2_V2",".png"), device = "png", width = 8, height = 8, units = "in")


#knitr::include_graphics(paste0(MARKDOWN_IMAGES,Sys.Date(),"_","paye_nuts2_V2",".svg"))


ggplotly(paye_stats_barchart,tooltip = "text") %>% 
  ggla_plotly_settings(legend=FALSE)   %>%
  layout(title = list(text = paste0("<b>","Payrolled employees, change on previous year by NUTS2 region","</b>",
                                    "<br>",
                                    "<sup>",
                                    paste0("Latest data for period ", paye_nuts2_stats$date[paye_nuts2_stats$new_date == max(paye_nuts2_stats$new_date)]),
                                    "</sup>",
                                    "<br>"),
                      font = list(size = 22),
                      y = .95, xref = "plot"),
         xaxis = list(tickfont = list(size = 15)),
         yaxis = list(tickfont = list(size = 15)),
         legend = list(font = list(size = 15))) 

payebarcap <- paste0("Source: HM Revenue and Customs – Pay As You Earn Real Time Information.","<br>","<br>", "Note: Estimates are based on where employees live.")


```

``` {r paye la stats,echo=FALSE}

paye_la_max_val <- perc_form(paye_la_stats %>% filter(new_date == max(new_date)) %>% filter(change_perc_since_feb20 == max(change_perc_since_feb20)) %>% select(change_perc_since_feb20),d = digits_perc)

paye_la_max_area <- paye_la_stats %>% filter(new_date == max(new_date)) %>% filter(change_perc_since_feb20 == max(change_perc_since_feb20)) %>% select(area)

paye_la_min_val <- perc_form(paye_la_stats %>% filter(new_date == max(new_date)) %>% filter(change_perc_since_feb20 == min(change_perc_since_feb20)) %>% select(change_perc_since_feb20),d = digits_perc)

paye_la_min_area <- paye_la_stats %>% filter(new_date == max(new_date)) %>% filter(change_perc_since_feb20 == min(change_perc_since_feb20)) %>% select(area)

paye_la_count_below_feb20 <- paye_la_stats %>% filter(new_date == max(new_date)) %>% mutate(count_below_feb20 = sum(change_perc_since_feb20<0)) %>% filter(row_number()==1) %>%  select(count_below_feb20)

paye_la_count_tot <- paye_la_stats %>% filter(new_date == max(new_date)) %>% add_count() %>% filter(row_number()==1) %>%  select(n)

```


<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

##  Change in payrolled employees by local authority

<br/>

There is also considerable variation in the change in payrolled employees by London authority.

* Compared to February 2020, the number of payrolled employees living in `r paye_la_max_area` is `r condi_text(paye_la_max_val,"up")` by `r abs2(paye_la_max_val)`% in `r paye_newmonth`, while the number living in `r paye_la_min_area` is `r condi_text(paye_la_min_val,"up")` by `r abs2(paye_la_min_val)`%. 

* Overall, the number of payrolled employees are still below pre-pandemic levels in `r paye_la_count_below_feb20` out of `r paye_la_count_tot` London authorities (including City of London, see map).

<a href="#top">Back to top</a>


```{r results="asis", echo=FALSE}
# This code is needed to force background of Leaflet map to be white
cat("
<style>
.leaflet-container {
    background: #FFF;
}
</style>
")
```

``` {r paye borough map, echo=FALSE,fig.cap = map_cap}

# Produces interactive map of boroughs with change in payrolled employees

#Read the shape file from the drive
boroughs <- readOGR(dsn = "Y:/Labour Market Update R Project/London map data/statistical-gis-boundaries-london/ESRI",
                    layer = "London_Borough_Excluding_MHW",
                    verbose = FALSE) %>%
  spTransform(CRS("+proj=longlat +datum=WGS84"))

# Prepare data
paye_la_stats_geo <- paye_la_stats %>%
  mutate(tooltip= paste("Change: ", perc_form(change_perc_since_feb20),"%", "\n", sep = "")) %>% 
  rename(GSS_CODE=gss_code, value=change_perc_since_feb20) %>% #Necessary since the shape files use upper caps variable names
  filter(new_date == max(new_date)) 

#Make a function to make things easier

map_function <- function(data, polygons, bins, palette, map_title){

  #Join data to polygons
  polygons@data <- left_join(polygons@data, data, by="GSS_CODE")
  
  #Create the palette
  pal <- colorBin(palette, domain = polygons$value, bins = bins)
  
  #Make tool tip
  labels <- sprintf(
    "<strong>%s</strong><br/>%s",
    polygons$NAME, polygons$tooltip
  ) %>% lapply(htmltools::HTML)

  
  #Make map
  map_output <- leaflet() %>%
    addPolygons(data = polygons,
                fillColor = ~pal(value),
                weight = 1,
                opacity = 1,
                color = "grey",
                fillOpacity = 0.8,
                label = labels,
                labelOptions = labelOptions(
                  style = list("font-weight" = "normal", padding = "3px 8px"),
                  textsize = "15px",
                  direction = "auto")) %>%
    addLegend(pal = pal, values = polygons$value, opacity = 0.8,
              title = map_title,
              position = "bottomright", labFormat = labelFormat(between = "% to ", suffix="%"))
  
  return(map_output)
}

# NB: THE BINS BELOW MUST BE CHANGED WHEN VALUES CHANGE!
bins <- c(-8,-6,-4,-2, 0,2,4,6,8)

# Colour palette, title and captions
map_palette <- gla_pal(palette_type = "diverging", main_colours = c("red","blue"), n = 8, remove_margin = "both")
map_title <- paste0("Change since","<br>" ,"February 2020 (%)")
map_cap <- paste0("Source: HM Revenue and Customs – Pay As You Earn Real Time Information. Contains Ordnance Survey data Crown copyright and database right [2015]","<br>","<br>", "Note: Estimates are based on where employees live.")

#Run Map
map_function(paye_la_stats_geo, boroughs, bins, map_palette, map_title)

```

``` {r, emp_figures, echo=FALSE}

emp_rate_uk <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom" & value_type_name == "Level" & measures == 20207)  %>% select(total_in_employment_aged_16_to_64),d = digits_perc)

emp_pp_quart_uk <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom"  & value_type_name == "Change on quarter" & measures == 20207)  %>% select(total_in_employment_aged_16_to_64),d = digits_perc)

emp_pp_year_uk <-perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom"  & value_type_name == "Change on year" & measures == 20207) %>% select(total_in_employment_aged_16_to_64),d = digits_perc)

```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

##  Employment rate

<br/>

Employment measures the number of people aged 16 and over in paid work and those who had a job that they were temporarily away from. The employment rate is the proportion of people aged between 16 and 64 years who are in employment.

For `r lfs_lastmonth` to `r lfs_newmonth`:

* The 16-64 employment rate in London was estimated at `r emp_rate`%; this represents `r condi_text(emp_pp_quart,"noun")` of `r abs2(emp_pp_quart)` percentage points (pp) on the previous quarter and
`r condi_text(emp_pp_year,"noun")` of `r abs2(emp_pp_year)`pp from a year earlier.

* The overall UK employment rate was estimated at `r emp_rate_uk`%, `r condi_text(emp_pp_quart_uk,"noun")` of `r abs2(emp_pp_quart_uk)`pp on the quarter and `r condi_text(emp_pp_year_uk,"noun")` of `r abs2(emp_pp_year_uk)`pp on the year.

<a href="#top">Back to top</a>

```{r employ chart, echo=FALSE,fig.cap = empcap, out.width='100%'}
### The below should be the relevant chart
employment_rate <- GLALineChart(data_set = lfs_stats, y_var = total_in_employment_aged_16_to_64, 
                         title = "Employment rate (% of working age population)",
                         caption = "\nSource: ONS Labour Force Survey.\n\nNote: The margin of error is +/- 1.3% for London and +/- 0.4% for the UK.",
                         chart_name = "employment", y_limits = c(70,80), nudge_y = c(-0.5, +0.5))


#knitr::include_graphics(paste0(MARKDOWN_IMAGES,Sys.Date(),"_","employment",".svg"))

ggplotly(employment_rate,tooltip = "text") %>% 
  ggla_plotly_settings()    %>% 
  layout(title = list(text = paste0("<b>","Employment rate (% of working age population)","</b>",
                                    "<br>"),
         font = list(size = 22),
         y = .95, xref = "plot"),
         xaxis = list(tickfont = list(size = 15)),
         yaxis = list(tickfont = list(size = 15)),
         legend = list(font = list(size = 15)),
         hovermode = "x") 

empcap <- paste0("Source: ONS Labour Force Survey.","<br>","<br>", "Note: The margin of error is +/- 1.4% for London and +/- 0.4% for the UK.")


```

``` {r, unem_figures, echo=FALSE}

unem_pp_quart_uk <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom" & value_type_name == "Change on quarter" & measures == 20207) %>% select(total_unemployed_aged_16_and_over),d = digits_perc)


unem_pp_year_uk <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom" & value_type_name == "Change on year" & measures == 20207)  %>% select(total_unemployed_aged_16_and_over),d = digits_perc)

```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

##  Unemployment rate

<br/>

Unemployment measures people without a job who have been actively seeking work within the last four weeks and are available to start work within the next two weeks. The unemployment rate is the proportion of the economically active population who are unemployed.

For `r lfs_lastmonth` to `r lfs_newmonth`:

* The unemployment rate for London was estimated at `r unem_rate`%; this represents `r condi_text(unem_pp_quart,"noun")` of `r abs2(unem_pp_quart)`pp on the previous quarter and `r condi_text(unem_pp_year,"noun")` of `r abs2(unem_pp_year)`pp from a year earlier.

* The UK unemployment rate was estimated at `r unem_rate_uk`%, `r condi_text(unem_pp_quart_uk,"noun")` of `r abs2(unem_pp_quart_uk)`pp from the previous quarter and `r condi_text(unem_pp_year_uk,"noun")` of `r abs2(unem_pp_year_uk)`pp on the year.

<a href="#top">Back to top</a>

```{r unemployed chart, echo=FALSE,fig.cap = unempcap, out.width='100%'}

unemployment_rate <- GLALineChart( data_set = lfs_stats, y_var = total_unemployed_aged_16_and_over, 
                             title = "Unemployment rate (% of economically active)",
                             caption = "\nSource: ONS Labour Force Survey.\n\nNote: The margin of error is +/- 0.8% for London and +/- 0.2% for the UK.",
                             chart_name = "unemployment", y_limits = c(0,8), nudge_y = 0.5)

#knitr::include_graphics(paste0(MARKDOWN_IMAGES,Sys.Date(),"_","unemployment",".svg"))

ggplotly(unemployment_rate,tooltip = "text") %>% 
  ggla_plotly_settings()    %>% 
  layout(title = list(text = paste0("<b>","Unemployment rate (% of economically active)","</b>",
                                    "<br>"),
         font = list(size = 22),
         y = .95, xref = "plot"),
         xaxis = list(tickfont = list(size = 15)),
         yaxis = list(tickfont = list(size = 15)),
         legend = list(font = list(size = 15)),
         hovermode = "x") 

unempcap <- paste0("Source: ONS Labour Force Survey.","<br>","<br>", "Note: The margin of error is +/- 0.8% for London and +/- 0.2% for the UK.")



```


```{r inact_figures,echo=FALSE}

inact_rate <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London" & value_type_name == "Level" & measures == 20207)  %>% select(total_economically_inactive_aged_16_to_64),d= digits_perc)
  
inact_pp_quart <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London" & value_type_name == "Change on quarter" & measures == 20207) %>% select(total_economically_inactive_aged_16_to_64),d= digits_perc)
  
inact_pp_year <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London" & value_type_name == "Change on year" & measures == 20207)  %>% select(total_economically_inactive_aged_16_to_64),d= digits_perc)

inact_rate_uk <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom" & value_type_name == "Level" & measures == 20207)  %>% select(total_economically_inactive_aged_16_to_64),d= digits_perc)
  
inact_pp_quart_uk <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom" & value_type_name == "Change on quarter" & measures == 20207) %>% select(total_economically_inactive_aged_16_to_64),d= digits_perc)
  
inact_pp_year_uk <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom" & value_type_name == "Change on year" & measures == 20207)  %>% select(total_economically_inactive_aged_16_to_64),d= digits_perc)

```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

## Economic inactivity rate

<br/>

The economic inactivity rate is the proportion of 16 to 64 year olds not in work and either not looking for or unable to work. This group includes some students, people who are looking after family/home, and people who are too ill to work (most of whom are long-term sick). 

For `r lfs_lastmonth` to `r lfs_newmonth`:

* The rate of economic inactivity in London was estimated at `r inact_rate`%; this is unchanged from the previous quarter and `r condi_text(inact_pp_year,"noun")` of `r abs2(inact_pp_year)`pp from a year earlier.

* The UK rate of economic inactivity was `r inact_rate_uk`%. This represents `r condi_text(inact_pp_quart_uk,"noun")` of `r abs2(inact_pp_quart_uk)`pp on the previous quarter and `r condi_text(inact_pp_year_uk,"noun")` of `r abs2(inact_pp_year_uk)`pp on the year.

<a href="#top">Back to top</a>

```{r inactive chart, echo=FALSE,fig.cap = inactcap, out.width='100%'}

econ_inactive <- GLALineChart( data_set = lfs_stats, y_var = total_economically_inactive_aged_16_to_64, 
                            title =  "Economic inactivity (% of working age population)",
                             caption = "\nSource: ONS Labour Force Survey.\n\nNote: The margin of error is not published for London, the UK margin is +/- 0.4%.",
                            chart_name = "inact", y_limits = c(15,25), nudge_y = c(-0.5, +0.5))

#knitr::include_graphics(paste0(MARKDOWN_IMAGES,Sys.Date(),"_","inact",".svg"))


ggplotly(econ_inactive,tooltip = "text") %>% 
  ggla_plotly_settings()    %>% 
  layout(title = list(text = paste0("<b>","Economic inactivity (% of working age population)","</b>",
                                    "<br>"),
         font = list(size = 22),
         y = .95, xref = "plot"),
         xaxis = list(tickfont = list(size = 15)),
         yaxis = list(tickfont = list(size = 15)),
         legend = list(font = list(size = 15)),
         hovermode = "x") 

inactcap <- paste0("Source: ONS Labour Force Survey.","<br>","<br>", "Note: The margin of error is not published for London, the UK margin is +/- 0.4%.")

```

``` {r, echo=FALSE}

claimant_count_stats_long <- claimant_count_stats %>% pivot_longer(cols=claimant_count:claimants_as_a_proportion_of_the_workforce,names_to = "measure_type", values_to = "measure_value",values_drop_na = TRUE) %>% 
  group_by(geography_name,age_name,sex_name,measure_type) %>% 
  mutate(change_month = measure_value - lag(measure_value,n=1),
         change_month_percent = 100*(measure_value - lag(measure_value,n=1))/lag(measure_value,n=1),
         change_year = measure_value - lag(measure_value,n=12),
         change_year_percent = 100*(measure_value - lag(measure_value,n=12))/lag(measure_value,n=12)) %>% 
  ungroup

```

``` {r claim_figures, echo=FALSE}

claim_period <- claimant_count_stats_long %>%  filter(date == max(date)) %>% filter(row_number()==1) %>% mutate(date = format(date, '%B %Y'))  %>% select(date)

claim_level <- value_form(claimant_count_stats_long %>% filter( sex_name == "Total" & geography_name == "London" & date == max(date) & age_name == "All categories: Age 16+" & measure_type == "claimant_count") %>% select(measure_value),s = signif_thous)
  
claim_rate <- perc_form(claimant_count_stats_long %>% filter( sex_name == "Total" & geography_name == "London" & date == max(date) & age_name == "All categories: Age 16+" & measure_type == "claimants_as_a_proportion_of_residents_aged_16_64") %>% select(measure_value),d = digits_perc)

claim_d_year <- value_form(claimant_count_stats_long %>% filter( sex_name == "Total" & geography_name == "London" & date == max(date) & age_name == "All categories: Age 16+" & measure_type == "claimant_count") %>% select(change_year),s = signif_thous)
  
claim_p_year <- perc_form(claimant_count_stats_long %>% filter( sex_name == "Total" & geography_name == "London" & date == max(date) & age_name == "All categories: Age 16+" & measure_type == "claimant_count") %>% select(change_year_percent),d = digits_perc)
  
claim_d_year_uk <- value_form(claimant_count_stats_long %>% filter( sex_name == "Total" & geography_name == "United Kingdom" & date == max(date) & age_name == "All categories: Age 16+" & measure_type == "claimant_count") %>% select(change_year),s = signif_thous)
  
claim_p_year_uk <- perc_form(claimant_count_stats_long %>% filter( sex_name == "Total" & geography_name == "United Kingdom" & date == max(date) & age_name == "All categories: Age 16+" & measure_type == "claimant_count") %>% select(change_year_percent),d = digits_perc)

```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

## Claimant count {.tabset}

<br/>

The claimant count is a measure of the number of people claiming benefits principally for the reason of being unemployed. It includes people claiming Jobseeker’s Allowance and those claiming Universal Credit who are required to seek work.

The latest data for `r claim_period` show that:

* There were `r claim_level` people claiming unemployment related benefits in London - equal to a claimant unemployment rate (as a proportion of residents aged 16 to 64) of `r claim_rate`%.

* The number of claimants in London has `r condi_text(claim_d_year,"increase")` by `r abs2(claim_d_year)`, or `r abs2(claim_p_year)`%, on the year, compared to `r condi_text(claim_d_year,"noun")` of `r abs2(claim_d_year_uk)`, or `r abs2(claim_p_year_uk)`%, in the UK overall.

<a href="#top">Back to top</a>

### Rate of residents aged 16-64
```{r claimant chart, echo=FALSE,fig.cap = claimcap, out.width='100%'}

### Create claimant count charts

claimant_count_percent <- GLALineChart( data_set = claimant_count_stats, y_var = claimants_as_a_proportion_of_residents_aged_16_64, 
                                suffix = "%", lfs_or_claims = "claims", title =  "Claimant count (as a % of residents aged 16 to 64)",
                                caption = "\nSource: ONS Claimant Count by sex and age (NSA).\n\nNote: may include some employed claimants on low hours or earnings.",
                               chart_name = "claimant_count_per", y_limits = c(0,10), nudge_y = 0.5)

#knitr::include_graphics(paste0(MARKDOWN_IMAGES,Sys.Date(),"_","claimant_count_per",".svg"))

ggplotly(claimant_count_percent,tooltip = "text") %>% 
  ggla_plotly_settings()    %>% 
  layout(title = list(text = paste0("<b>","Claimant count (as a % of residents aged 16 to 64)","</b>",
                                    "<br>"),
         font = list(size = 22),
         y = .95, xref = "plot"),
         xaxis = list(tickfont = list(size = 15)),
         yaxis = list(tickfont = list(size = 15)),
         legend = list(font = list(size = 15)),
         hovermode = "x") 

claimcap <- paste0("Source: ONS Claimant Count by sex and age (NSA).","<br>","<br>", "Note: may include some employed claimants on low hours or earnings.")

  ### Manually plot total London claimants

new_release_text <- claimant_count_stats$date_name[claimant_count_stats$date == max(claimant_count_stats$date)]


claimant_count_total <- claimant_count_stats %>%
                        filter( geography_name == "London" & sex_name == "Total" & !is.na(claimant_count) 
                                & age_name == "All categories: Age 16+") %>% 
                        ggplot(mapping = aes(x = date, y = claimant_count, 
                                             colour = geography_name, group = geography_name)) +
                        ggla_line(aes(size= geography_name)) +
                        scale_size_manual(values = c(4 * mm_to_pt, 2 * mm_to_pt)) +
                        scale_colour_manual(values = pal) +
                        ggla_highlight(filter_type = "end") +
                        ggla_highlight(mapping = aes(label = format(claimant_count, big.mark = ",")),
                                       geom = GeomGLATextHighlight, filter_type = "end",  size = 4.5,
                                       position = position_nudge(y = 20000),check_overlap = TRUE)+
                        coord_cartesian(clip = 'off') +
                        ggla_axisat0() +
                        scale_y_continuous(expand = c(0, 0), labels = dollar_format(prefix = "", 
                                                                                    suffix = "", 
                                                                                    largest_with_cents = 1), 
                                           limits = c(0,750000)) +
                        scale_x_date( date_breaks = "2 months",
                                      date_labels = "%b'%y",
                                      expand = expansion( mult = c(0.1,0.02))) +
                        theme(plot.margin = unit(c(1,1,1,1), "cm"))+
                        labs(title =  "Claimant count in London",
                             subtitle = paste0("Latest data for period ", new_release_text),
                             caption = "\nSource: ONS Claimant Count by sex and age (NSA).\n\nNote: May include some employed claimants on low hours or earnings.")+
                       theme(plot.caption = element_text(color = rgb(166,166,166,maxColorValue = 255)))

ggsave( paste0(MARKDOWN_IMAGES,Sys.Date(),"_","claimant_count",".svg"), device = "svg", width = 8, height = 8, units = "in")
ggsave( paste0(MARKDOWN_IMAGES,Sys.Date(),"_","claimant_count",".png"), device = "png", width = 8, height = 8, units = "in") 

# knitr::include_graphics(paste0(MARKDOWN_IMAGES,Sys.Date(),"_","claimant_count",".svg"))

  ### Claimant count bar charts (edit variable names)

new_release <- max(claimant_count_stats$date)
year_ago <- ymd(new_release) - years(1)


```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

### Year-on-year percentage change

``` {r claimant increase, echo = FALSE,fig.cap = claimbarcap, out.width='100%'}

    ### Increases in claimants by age
pal <- gla_pal(gla_theme = "default", palette_type = "highlight", n = c(1, 1))
theme_set(theme_gla(gla_theme = "default"))

claimant_count_bar <- claimant_count_stats %>%
  group_by(sex_name, geography_name, date) %>% 
  mutate(chart_ranking = case_when( age_name == "All categories: Age 16+" ~ 1,
                                    TRUE ~ rank((age_name))+1)) %>% 
  group_by( age_name, sex_name,geography_name) %>% 
  mutate( annual_change = case_when( date <= as.Date(year_ago) ~ NaN,
                                     date > as.Date(year_ago) ~ 100*(claimant_count - claimant_count[ date == as.Date(year_ago)])/claimant_count[ date == as.Date(year_ago)])) %>% 
  ungroup() %>% 
   filter( sex_name == "Total" & !is.na(annual_change) & (date == max(date)) & 
            !(age_name %in% c("Aged 16-17", "Aged 18-21", "Aged 18-24", "Aged 50+" , "Age unknown (clerical claims)", "Aged 25-49", "Aged 65+"))
          & geography_name %in% c("London", "United Kingdom")) %>% 
  ggplot(mapping = aes(x = reorder(age_name,chart_ranking), 
                       y = annual_change, 
                       colour = geography_name , fill=geography_name,
                       text = paste(age_name, "\n",
                                    geography_name, "\n",
                                    "Change in count: ", perc_form(annual_change),"%", "\n",
                                    sep = ""))) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.4) +
  geom_hline(aes(yintercept=0), colour="gray45") +
  scale_colour_manual(values = pal) +
  scale_fill_manual(values = pal) +
  theme_set(theme_gla(gla_theme = "default")) +
  scale_y_continuous(limits = c(-40, 30), labels = dollar_format(prefix = "", 
                                                                 suffix = "%", 
                                                                 largest_with_cents = 1)) +
  theme(axis.text.x = element_text(angle = 90, hjust=1, vjust=0.5)) +
  theme(plot.margin = unit(c(1,1,1,1), "cm"))+
  labs(title = "Year-on-year percentage change in claimant count",
       subtitle = paste0("Latest data for period ", format(new_release,"%B %Y")),
       caption = "\nSource: ONS Claimant Count by sex and age (NSA).\n\nNote: May include some employed claimants on low hours or earnings.") +
  theme(plot.caption = element_text(color = rgb(166,166,166,maxColorValue = 255)))

ggsave( paste0(MARKDOWN_IMAGES,Sys.Date(),"_","claimant_age_increase.svg"), device = "svg", width = 8, height = 8, units = "in")
ggsave( paste0(MARKDOWN_IMAGES,Sys.Date(),"_","claimant_age_increase.png"), device = "png", width = 8, height = 8, units = "in")

#knitr::include_graphics(paste0(MARKDOWN_IMAGES,Sys.Date(),"_","claimant_age_increase",".svg"))

ggplotly(claimant_count_bar,tooltip = "text") %>% 
  ggla_plotly_settings()    %>% 
  layout(title = list(text = paste0("<b>","Year-on-year percentage change in claimant count","</b>",
                                    "<br>",
                                    "<sup>",
                                    paste0("Latest data for period ", format(new_release,"%B %Y")),
                                    "</sup>",
                                    "<br>"),
                      font = list(size = 22),
                      y = .95, xref = "plot"),
         xaxis = list(tickfont = list(size = 15), tickangle = 270),
         yaxis = list(tickfont = list(size = 15)),
         legend = list(font = list(size = 15))) 


claimbarcap <- paste0("Source: ONS Claimant Count by sex and age (NSA).","<br>","<br>", "Note: may include some employed claimants on low hours or earnings.")


```


``` {r wfj_figures, echo=FALSE}

wfj_newmonth <- wfj_stats %>%  filter(new_date == max(new_date)) %>% filter(row_number()==1) %>% mutate(new_date = format(new_date, '%B %Y'))  %>% select(new_date)

wfj_lastmonth <- wfj_stats %>% filter(new_date == (ymd(max(new_date)) %m+% - months(3))) %>% filter(row_number()==1) %>% mutate(new_date = format(new_date, '%B %Y'))  %>% select(new_date)

wfj_level <- value_form(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% mutate(obs_value=obs_value/1000000) %>% select(obs_value),s=signif_mil, d=digits_mil)


wfj_p_quart_uk <- perc_form(100*(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "United Kingdom" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd(max(new_date)) %m+% - months(3)) & geography_name == "United Kingdom" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value))/(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd(max(new_date)) %m+% - months(3)) & geography_name == "United Kingdom" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value)),d = digits_perc)


wfj_p_dec19_uk <- perc_form(100*(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "United Kingdom" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd("2019-12-01")) & geography_name == "United Kingdom" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value))/(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd("2019-12-01")) & geography_name == "United Kingdom" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value)),d = digits_perc)

wfj_emp_d_dec19 <- value_form(wfj_stats %>% filter(item_name == "employee jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "employee jobs" & new_date == (ymd("2019-12-01")) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value),s = signif_thous)
  
wfj_emp_p_dec19 <- perc_form(100*(wfj_stats %>% filter(item_name == "employee jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "employee jobs" & new_date == (ymd("2019-12-01")) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value))/(wfj_stats %>% filter(item_name == "employee jobs" & new_date == (ymd("2019-12-01")) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value)),d = digits_perc)

wfj_self_d_dec19 <- value_form(wfj_stats %>% filter(item_name == "self-employment jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "self-employment jobs" & new_date == (ymd("2019-12-01")) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value),s = signif_thous)
  
wfj_self_p_dec19 <- perc_form(100*(wfj_stats %>% filter(item_name == "self-employment jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "self-employment jobs" & new_date == (ymd("2019-12-01")) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value))/(wfj_stats %>% filter(item_name == "self-employment jobs" & new_date == (ymd("2019-12-01")) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value)),d = digits_perc)


```


<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

## Workforce Jobs {.tabset}

<br/>

### Number of workforce jobs

The Workforce Jobs series provides quarterly estimates of the number of jobs in the UK and is the ONS’ preferred source of jobs broken down by region of workplace and industry.

The latest data for `r wfj_newmonth` show that:

* The total number of workforce jobs in London was estimated at `r wfj_level` million, `r condi_text(wfj_p_quart,"noun")` of approximately 
`r abs2(wfj_p_quart)`% from `r wfj_lastmonth` against a UK average of `r wfj_p_quart_uk`%.  


* The latest estimate represents `r condi_text(wfj_d_dec19,"noun")` of approximately `r abs2(wfj_d_dec19)` or `r abs2(wfj_p_dec19)`% from December 2019 levels (the pre-pandemic peak in London) compared with `r condi_text(wfj_p_dec19_uk,"noun")` of `r abs2(wfj_p_dec19_uk)`% across the UK on average. 


* Total jobs includes both employee and self-employment jobs alongside other types of employment; since December 2019, the latest employee jobs estimate has `r condi_text(wfj_emp_d_dec19,"increase")` by 
`r abs2(wfj_emp_d_dec19)` (`r abs2(wfj_emp_p_dec19)`%)  while the self-employment jobs estimate has `r condi_text(wfj_self_d_dec19,"increase")` by `r abs2(wfj_self_d_dec19)` (`r abs2(wfj_self_p_dec19)`%).

<a href="#top">Back to top</a>

```{r workforce jobs trend, echo=FALSE,fig.cap = wfjcap, out.width='100%'}


pal <- gla_pal(gla_theme = "default", palette_type = "highlight", n = c(1, 1))
theme_set(theme_gla(gla_theme = "default"))

wfj_time_chart <- wfj_stats %>%
  filter( new_date >= "2010-03-01" & measures_name == "Value" & geography_name == "London" & industry_name == "Total" & item_name == "total workforce jobs") %>%
  ggplot(mapping = aes(x = new_date, y = obs_value/1000,
                       colour = geography_name, group = geography_name,
                       text = paste( format(new_date,'%B %Y'), "\n",
                                     "Jobs: ", value_form(obs_value/1000,s=3), "\n",
                                     sep = ""))) +
  ggla_line(aes(size= geography_name)) +
  scale_size_manual(values = c(4 * mm_to_pt, 2 * mm_to_pt)) +
  scale_colour_manual(values = pal) +
  ggla_highlight(filter_type = "end") +
  ggla_highlight(mapping = aes(label = paste0(format(obs_value/1000, digits = 1, big.mark=","))),
                 geom = GeomGLATextHighlight, filter_type = "end",  size = 4.5,
                 position = position_nudge(y = c(200,0)))+
  coord_cartesian(clip = 'off') +
  scale_y_continuous(expand = c(0, 0), labels = dollar_format(prefix = "",
                                                              largest_with_cents = 1),
                     limits = c(3000,7000)) +
  scale_x_date( date_breaks = "18 months",
                date_labels = "%b'%y",
                expand = expansion( mult = c(0.1,0.02))) +
  theme(plot.margin = unit(c(1,1,1,1), "cm"))+
  labs(title =  "Total number of Workforce jobs in London",
       subtitle = paste0("Seasonally adjusted (thousands), latest data for ", wfj_stats$date_name[wfj_stats$new_date == max(wfj_stats$new_date)]),
       caption = "\nSource: ONS Workforce Jobs.\n\nNote: The margin of error for all jobs is +/- 0.9% for London and +/- 0.3% for the UK.")+
  theme(plot.caption = element_text(color = rgb(166,166,166,maxColorValue = 255)))

ggsave( paste0(MARKDOWN_IMAGES,Sys.Date(),"_","wfj_jobs",".svg"), device = "svg", width = 8, height = 8, units = "in")
ggsave( paste0(MARKDOWN_IMAGES,Sys.Date(),"_","wfj_jobs",".png"), device = "png", width = 8, height = 8, units = "in")

                     
#knitr::include_graphics(paste0(MARKDOWN_IMAGES,Sys.Date(),"_","wfj_jobs",".svg"))

ggplotly(wfj_time_chart,tooltip = "text", dynamicticks=TRUE) %>% 
  ggla_plotly_settings()    %>% 
  layout(title = list(text = paste0("<b>","Total number of Workforce jobs in London","</b>",
                                    "<br>",
                                    "<sup>",
                                    paste0("Seasonally adjusted (thousands), latest data for ", wfj_stats$date_name[wfj_stats$new_date == max(wfj_stats$new_date)]),
                                    "</sup>",
                                    "<br>"),
                      font = list(size = 22),
                      y = .95, xref = "plot"),
         xaxis = list(tickfont = list(size = 15)),
         yaxis = list(tickfont = list(size = 15)),
         legend = list(font = list(size = 15))) 


wfjcap <- paste0("Source: ONS Workforce Jobs.","<br>","<br>", "Note: The margin of error for all jobs is +/- 0.9% for London and +/- 0.3% for the UK.")


```

``` {r,echo=FALSE}
# Find industries with lowest and highest growth in recent data

sector_thresh <- 25000

lowest_growth <- wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & obs_value>=sector_thresh & !is.na(perc_change_dec19_within_date_rank)) %>% slice_max(perc_change_dec19_within_date_rank,n=3) 

highest_growth <- wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & obs_value>=sector_thresh & !is.na(perc_change_dec19_within_date_rank)) %>% slice_min(perc_change_dec19_within_date_rank,n=3) 

lowest_growth_q <- wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & obs_value>=sector_thresh & !is.na(perc_change_quart_within_date_rank)) %>% slice_max(perc_change_quart_within_date_rank,n=3) 

highest_growth_q <- wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & obs_value>=sector_thresh & !is.na(perc_change_quart_within_date_rank)) %>% slice_min(perc_change_quart_within_date_rank,n=3) 

## Find industry names and growth figures
## by year

low_1_name <- lowest_growth %>% filter(row_number()==1) %>% select(industry_name_simple)
  
low_1_d <- value_form(lowest_growth %>% filter(row_number()==1) %>% select(change_since_dec19),s = signif_thous)
  
low_1_p <- perc_form(lowest_growth %>% filter(row_number()==1) %>% select(change_perc_since_dec19),d = digits_perc)

low_2_name <- lowest_growth %>% filter(row_number()==2) %>% select(industry_name_simple)
  
low_2_d <- value_form(lowest_growth %>% filter(row_number()==2) %>% select(change_since_dec19),s = signif_thous)
  
low_2_p <- perc_form(lowest_growth %>% filter(row_number()==2) %>% select(change_perc_since_dec19),d = digits_perc)

low_3_name <- lowest_growth %>% filter(row_number()==3) %>% select(industry_name_simple)
  
low_3_d <- value_form(lowest_growth %>% filter(row_number()==3) %>% select(change_since_dec19),s = signif_thous)
  
low_3_p <- perc_form(lowest_growth %>% filter(row_number()==3) %>% select(change_perc_since_dec19),d = digits_perc)


high_1_name <- highest_growth %>% filter(row_number()==1) %>% select(industry_name_simple)
  
high_1_d <- value_form(highest_growth %>% filter(row_number()==1) %>% select(change_since_dec19),s = signif_thous)
  
high_1_p <- perc_form(highest_growth %>% filter(row_number()==1) %>% select(change_perc_since_dec19),d = digits_perc)

high_2_name <- highest_growth %>% filter(row_number()==2) %>% select(industry_name_simple)
  
high_2_d <- value_form(highest_growth %>% filter(row_number()==2) %>% select(change_since_dec19),s = signif_thous)
  
high_2_p <- perc_form(highest_growth %>% filter(row_number()==2) %>% select(change_perc_since_dec19),d = digits_perc)

high_3_name <- highest_growth %>% filter(row_number()==3) %>% select(industry_name_simple)
  
high_3_d <- value_form(highest_growth %>% filter(row_number()==3) %>% select(change_since_dec19),s = signif_thous)
  
high_3_p <- perc_form(highest_growth %>% filter(row_number()==3) %>% select(change_perc_since_dec19),d = digits_perc)

```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

### Workforce jobs by sector

Between December 2019 and `r wfj_newmonth`*:

* `r low_1_name` recorded the largest percentage decrease in jobs
in London (`r low_1_d` or `r low_1_p`%); this was followed by `r low_2_name` (`r low_2_d` or `r low_2_p`%) and `r low_3_name`
(`r low_3_d` or `r low_3_p`%).


* The following industry groups shown in the chart saw the largest positive percentage change in jobs in London: 
  + `r high_1_name` (`r high_1_d` or `r high_1_p`%)
  + `r high_2_name` (`r high_2_d` or `r high_2_p`%)
  + `r high_3_name` (`r high_3_d` or `r high_3_p`%)
  
<a href="#top">Back to top</a>

<foot>*Note: excluding sectors with less than `r value_form(sector_thresh,s = signif_thous)` jobs in London. </foot>

```{r wfj sector, echo=FALSE,fig.cap = wfjseccap, out.width='100%'}

pal <- gla_pal(gla_theme = "default", palette_type = "highlight", n = c(1, 1))
theme_set(theme_gla(gla_theme = "default"))


# Select industries large in London, e.g. 20k+
lon_large_inds <- wfj_stats %>%
  filter( item_name == "total workforce jobs" & new_date == max(new_date) & measures_name == "Value" & obs_value>=sector_thresh 
          & geography_name =="London") %>% 
  select(industry_name_simple)

lon_large_inds <- unlist(lon_large_inds)

wjf_industry_bar <- wfj_stats %>%
  filter( item_name == "total workforce jobs" & new_date == max(new_date) & measures_name == "Value" & industry_name_simple %in% lon_large_inds
          & geography_name %in% c("London", "United Kingdom")) %>% 
  group_by(geography_name) %>% #To ensure chart ranking is correct, make rank based on London and assign within sector
  mutate(chart_ranking = case_when( 
    industry_name_simple == "Total" & geography_name=="London" ~ 1,
    !(industry_name_simple == "Total") & geography_name=="London" ~ dense_rank(desc(change_perc_since_dec19))+1,
    TRUE ~ 99))  %>% 
  group_by(industry_name_simple) %>% 
  mutate(chart_ranking=min(chart_ranking)) %>% 
  ungroup() %>% 
  arrange(chart_ranking) %>% 
  ggplot(mapping = aes(x =  factor(reorder(industry_name_simple, -chart_ranking)), 
                       y = change_perc_since_dec19, 
                       colour = geography_name , fill=geography_name, width = 0.5,
                       text = paste(industry_name_simple, "\n",
                                    geography_name, "\n",
                                    "Change: ", perc_form(change_perc_since_dec19),"%", "\n",
                                    sep = ""))) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_hline(aes(yintercept=0), colour="gray45") +
  geom_vline(xintercept=c(length(lon_large_inds) - 0.5),colour="gray45")+ #adds line below total
  scale_color_manual(values = pal, aesthetics = "colour")+
  scale_fill_manual(values = pal, aesthetics = "fill")+
  theme_set(theme_gla(gla_theme = "default", y_label_length=100)) + #GLA theme and removes lines below y-axis labels
  scale_y_continuous(limits = c(-18, 12), labels = dollar_format(prefix = "", 
                                                                 suffix = "%", 
                                                                 largest_with_cents = 1)) +
  theme(plot.margin = unit(c(1,1,1,1), "cm"))+
  scale_x_discrete(expand = c(0,0)) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_line( size=.5 )) + 
  coord_flip()+
  labs(title = paste0("Percentage change in Workforce Jobs by industry group"),
       subtitle = paste0("Seasonally adjusted, between December 2019 and ", wfj_stats$date_name[wfj_stats$new_date == max(wfj_stats$new_date)]),
       caption = "\nSource: ONS Workforce Jobs.\n\nNote: The margin of error for all jobs is +/- 0.9% for London and +/- 0.3% for the UK.")+
  theme(plot.caption = element_text(color = rgb(166,166,166,maxColorValue = 255)))

ggsave(paste0(MARKDOWN_IMAGES,Sys.Date(),"_","wfj_industries",".svg"), device = "svg", width = 8, height = 8, units = "in")
ggsave(paste0(MARKDOWN_IMAGES,Sys.Date(),"_","wfj_industries",".png"), device = "png", width = 8, height = 8, units = "in")


#knitr::include_graphics(paste0(MARKDOWN_IMAGES,Sys.Date(),"_","wfj_industries",".svg"))

ggplotly(wjf_industry_bar,tooltip = "text") %>% 
  ggla_plotly_settings()    %>% 
  layout(title = list(text = paste0("<b>","Percentage change in Workforce Jobs by industry group","</b>",
                                    "<br>",
                                    "<sup>",
                                    paste0("Seasonally adjusted, between December 2019 and ", wfj_stats$date_name[wfj_stats$new_date == max(wfj_stats$new_date)]),
                                    "</sup>",
                                    "<br>"),
                      font = list(size = 22),
                      y = .95, xref = "plot"),
         xaxis = list(tickfont = list(size = 15), showgrid=TRUE),
         yaxis = list(tickfont = list(size = 15), showgrid=FALSE),
         legend = list(font = list(size = 15)),
         margin = list(l = 160))


wfjseccap <- paste0("Source: ONS Workforce Jobs.","<br>","<br>", "Note: The margin of error for all jobs is +/- 0.9% for London and +/- 0.3% for the UK.")

```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

## Summary of headline indicators

<br/>


``` {r flextable indicators, echo = FALSE, ft.align="left", out.width='100%'}

#-------------------------------------------------------------------------------
#  Create the headline stats table as data frame
#-------------------------------------------------------------------------------


condi_arrow <- function(value) {
  arrow <- case_when( value < 0 ~ "u",
                      value == 0 ~ "v",
                      value > 0 ~ "t")
  return(arrow)
}

# Create rows to use as subheaders in table
sub_head_rows <- tibble("geography_name"=c("Employment","Unemployment","Economic Inactivity"),"indicator"=c("emp_","unem_","inact_"),tab_rank=1)

headline_rates <- lfs_stats %>% 
  filter(sex_name == "Total" & geography_name %in% c("London", "United Kingdom") & date == max(date)) %>% 
  mutate(measures_name = case_when(measures == 20100 ~ "num",
                                   measures == 20207 ~ "rate"),
         val_type = case_when(value_type_name == "Level" ~ "le",
                              value_type_name == "Change on quarter" ~ "cq",
                              value_type_name == "Change on year" ~ "cy"),
         geography_name = case_when(geography_name == "United Kingdom" ~ "UK" ,
                                   TRUE ~ "London")) %>% 
  select("date_name", "geography_name","measures_name","val_type","total_in_employment_aged_16_and_over","total_unemployed_aged_16_and_over",
         "total_economically_inactive_aged_16_and_over","total_in_employment_aged_16_to_64","total_economically_inactive_aged_16_to_64") %>% 
  pivot_longer(names_to = "obs_type_help", values_to = "obs_value", cols = "total_in_employment_aged_16_and_over":"total_economically_inactive_aged_16_to_64") %>% 
  mutate(obs_type = case_when(obs_type_help == "total_in_employment_aged_16_and_over" ~ "emp_16_plus",
                              obs_type_help == "total_unemployed_aged_16_and_over" ~ "unem_16_plus",
                              obs_type_help == "total_economically_inactive_aged_16_and_over" ~ "inact_16_plus",
                              obs_type_help == "total_in_employment_aged_16_to_64" ~ "emp_16_64",
                              obs_type_help == "total_economically_inactive_aged_16_to_64" ~ "inact_16_64")) %>%
  filter((measures_name=="num" & obs_type %in% c("emp_16_plus","unem_16_plus","inact_16_64")) | 
            (measures_name=="rate" & obs_type %in% c("emp_16_64","unem_16_plus","inact_16_64"))) %>% #Should make the right combination for table
  mutate(indicator =  str_extract(obs_type,".*?_")) %>% 
  mutate(col_name = paste(measures_name,val_type, sep = "_")) %>% 
  select("date_name","geography_name","obs_value","indicator","col_name") %>% 
  pivot_wider(names_from = "col_name", values_from = "obs_value") %>% 
  mutate(num_cq2 = condi_arrow(num_cq),
             num_cy2 = condi_arrow(num_cy),
             rate_cq2 = condi_arrow(rate_cq),
             rate_cy2 = condi_arrow(rate_cy),
         num_le = round(num_le/1000,0),
         num_cq = round(num_cq/1000,0),
         num_cy = round(num_cy/1000,0),
         tab_rank = case_when(geography_name == "London" ~ 2,
                              geography_name == "UK" ~ 3)) %>% 
  bind_rows(sub_head_rows) %>% 
  arrange(match(indicator,c("emp_","unem_","inact_")), tab_rank) %>% # Row order
  select("date_name","geography_name","indicator","num_le","num_cq","num_cq2","num_cy","num_cy2","rate_le","rate_cq","rate_cq2","rate_cy","rate_cy2") 
  
#-------------------------------------------------------------------------------
# Format table appearance manually
#-------------------------------------------------------------------------------

std_border = fp_border(width = 1)
no_border = fp_border(width = 0)

headline_rates_table <-  flextable(headline_rates, 
                                  col_keys=c("geography_name","num_le","num_cq","num_cq2","num_cy","num_cy2","rate_le","rate_cq","rate_cq2","rate_cy","rate_cy2")) %>% 
  colformat_num(
    big.mark = ",", decimal.mark = ".", digits=0) %>% 
  font(fontname='Arial', part = "all") %>% 
  # Setting colors and font of change variables to make indicator arrows
  font(j = c("num_cq2", "num_cy2", "rate_cq2", "rate_cy2"), fontname ='Marlett') %>% 
  color( i= ~ indicator %in% "emp_" & num_cq2 %in% "t", j = ~ num_cq2, color = "green") %>%  # Specifically for employment
  color( i= ~ indicator %in% "emp_" & num_cq2 %in% "u", j = ~ num_cq2, color = "red") %>% 
  color( i= ~ indicator %in% "emp_" & num_cy2 %in% "t", j = ~ num_cy2, color = "green") %>% 
  color( i= ~ indicator %in% "emp_" & num_cy2 %in% "u", j = ~ num_cy2, color = "red") %>% 
  color( i= ~ indicator %in% "emp_" & rate_cq2 %in% "t", j = ~ rate_cq2, color = "green") %>% 
  color( i= ~ indicator %in% "emp_" & rate_cq2 %in% "u", j = ~ rate_cq2, color = "red") %>% 
  color( i= ~ indicator %in% "emp_" & rate_cy2 %in% "t", j = ~ rate_cy2, color = "green") %>% 
  color( i= ~ indicator %in% "emp_" & rate_cy2 %in% "u", j = ~ rate_cy2, color = "red") %>%  
  color( i= ~ !(indicator %in% "emp_") & num_cq2 %in% "t", j = ~ num_cq2, color = "red") %>% # For the others
  color( i= ~ !(indicator %in% "emp_") & num_cq2 %in% "u", j = ~ num_cq2, color = "green") %>% 
  color( i= ~ !(indicator %in% "emp_") & num_cy2 %in% "t", j = ~ num_cy2, color = "red") %>% 
  color( i= ~ !(indicator %in% "emp_") & num_cy2 %in% "u", j = ~ num_cy2, color = "green") %>% 
  color( i= ~ !(indicator %in% "emp_") & rate_cq2 %in% "t", j = ~ rate_cq2, color = "red") %>% 
  color( i= ~ !(indicator %in% "emp_") & rate_cq2 %in% "u", j = ~ rate_cq2, color = "green") %>% 
  color( i= ~ !(indicator %in% "emp_") & rate_cy2 %in% "t", j = ~ rate_cy2, color = "red") %>% 
  color( i= ~ !(indicator %in% "emp_") & rate_cy2 %in% "u", j = ~ rate_cy2, color = "green") %>% 
  color( i= ~ num_cq2 %in% "v", j = ~ num_cq2, color = "orange") %>% # If equal
  color( i= ~ num_cy2 %in% "v", j = ~ num_cy2, color = "orange") %>% 
  color( i= ~ rate_cq2 %in% "v", j = ~ rate_cq2, color = "orange") %>% 
  color( i= ~ rate_cy2 %in% "v", j = ~ rate_cy2, color = "orange") %>% 
  #color( i= ~ geography_name %in% "London", j = ~ geography_name, color = rgb(109,167,222, maxColorValue = 255)) %>% 
  #color( i= ~ geography_name %in% "UK", j = ~ geography_name, color = rgb(204,204,204, maxColorValue = 255)) %>% 
  bold(i= ~ geography_name %in% c("Employment","Unemployment","Economic Inactivity"), j = ~ geography_name) %>% 
  add_header_row(colwidths = c(1 ,5, 5),
                 values = c(paste0("Headline estimates for ",headline_rates %>% filter(!is.na(date_name)) %>% filter(row_number()==1)  %>% select(date_name)), "Number* (thousands)","Rate** (%)")) %>% 
  align(part = "header", align = "center") %>% 
  set_header_labels( "geography_name"="","num_le"="Latest Estimate","num_cq"="Change on Quarter","num_cq2"="","num_cy"="Change on Year","num_cy2"="","rate_le"="Latest Estimate","rate_cq"="Change on Quarter","rate_cq2"="","rate_cy"="Change on Year","rate_cy2"="") %>% 
  bold(i=1, part = "header") %>% 
   add_footer_lines("Source: ONS Labour Force Survey. \n \nNotes: All figures are seasonally adjusted. *Numbers are based on total population (16+, male and female), except for inactivity which is 16-64. **Rates are based on working age population (16–64, male and female), except for the unemployment rate which is age 16+.") %>% 
  color( part = "footer", color = rgb(166,166,166,maxColorValue = 255)) %>%
      fontsize(size = 14, part = "all") %>% 
  # Cell widths
  width(j=1, width=6, unit="cm") %>% 
  width(j= ~ num_cq2 + num_cy2 + rate_cq2 + rate_cy2, width=1, unit="cm") %>% 
  # Merging of header cells
  merge_at(i=2,j=3:4,part="header") %>%  
  merge_at(i=2,j=5:6,part="header") %>%  
  merge_at(i=2,j=8:9,part="header") %>%  
  merge_at(i=2,j=10:11,part="header") %>%  
  merge_at(i=1:2,j=1,part="header") %>%  
  # Custom borders
  border_outer(part="header") %>% 
  border_outer(part="body") %>% 
  hline(i=1,part="header", border=no_border) %>% 
  hline(i= ~ geography_name %in% c("Employment","Unemployment","Economic Inactivity") ,part="body", border=std_border) %>% 
  hline(i= ~ geography_name %in% c("UK") ,part="body", border=std_border) %>% 
  vline(border = std_border, part = "all", j = ~ geography_name ) %>% 
  vline(border = std_border, part = "all", j = ~ num_cy2 ) %>% 
  fix_border_issues()

knit_print(headline_rates_table)


```

<a href="#top">Back to top</a>

</font>

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>