---
title:  "<br> London Labour Market Update"
date: "`r format(Sys.Date(),'%B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    css:  !expr "here::here('FORMATTING','GLAstyle.css')"
    includes:
      in_header: !expr "here::here('FORMATTING','favicon.html')"
      before_body: !expr "here::here('FORMATTING','header.html')"
      after_body: !expr "here::here('FORMATTING','footer.html')"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE,scipen=999)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

#################################################################
###  Load packages and set paths
#################################################################

#devtools::install_github("Greater-London-Authority/gglaplot")
#devtools::install_github("Greater-London-Authority/ldndatar", auth_token = "96e66bb601f49f62f0bb9bdcb73a849ece358ad1")
#remotes::install_github("wilkelab/ggtext")

library("here") # To set project folder dynamically

library("extrafont")
library("remotes")


library("knitr")
library("tidyverse")
library("ggplot2")
library("ggthemes")
library("nomisr")
library("devtools")
library("remotes")
library("scales")
library("gglaplot")
library("data.table")
library("janitor")
library("lubridate")
library("readr")
#library("ldndatar")
library("ggrepel")
library("plotly")
library("magrittr")
library("zoo")
library("kableExtra")
library("leaflet")
library("rgdal")
library(httr)

library("flextable")
library("officer") # for fp_text

library("svDialogs") #for pop-ups

### Paths

#setwd(wd)
#wd <- "Y:\\Labour Market Update R Project\\MARKDOWN"

INPUT <- paste0(here("INPUT"),"/")
INTERMEDIATE <- paste0(here("INTERMEDIATE"),"/")
IMAGES <- paste0(here("IMAGES"),"/")
FORMATTING <- paste0(here("FORMATTING"),"/")


### Geographic codes for UK, London and Boroughs.
### See "Borough codes.csv" in source folder for reference. 

Group <- c(2013265927,2092957697,1811939540,1811939541,1811939542,1811939543,1811939544,1811939526,1811939527,1811939545,1811939546,1811939547,
           1811939548,1811939528,1811939529,1811939530,1811939549,1811939550,1811939551,1811939552,1811939531,1811939532,1811939553,1811939533,
           1811939534,1811939554,1811939535,1811939555,1811939556,1811939536,1811939557,1811939537,1811939558,1811939538,1811939539)

################################################################################
### Presets for presenting figures etc.
################################################################################

signif_thous = 3 #How many significant figures to use when figures are in thousands
signif_mil = 3 # For millions
signif_perc = 2 #For percentages
signif_pp = 3 #For percentage points

digits_mil = 2
digits_perc = 1

################################################################################
### Functions
################################################################################

  ### Define function to always present percentages with one decimal and figures without decimals
perc_form = function(x, d=1) sprintf(paste0("%1.",d,"f"), x) 

value_form = function(x,s=2,d= -1) format(signif(round(as.numeric(x), d),s), big.mark=",")


  ### Function to make conditional text based on values
condi_text <- function(x,t=c("increase","noun","rise","growth","up","above")) { #Define names by the positive change
  if(t=="increase") {y = case_when(x<0 ~ "decreased",
                            x>0 ~ "increased")
                            return(y)}
  else if(t=="noun") {y = case_when(x<0 ~ "a decrease",
                            x>0 ~ "an increase")
                            return(y)}
  else if(t=="rise") {y = case_when(x<0 ~ "fell",
                            x>0 ~ "rose")
                            return(y)}
  else if(t=="growth") {y = case_when(x<0 ~ "shrank",
                            x>0 ~ "grew")
                            return(y)}
  else if(t=="up") {y = case_when(x<0 ~ "down",
                            x>0 ~ "up")
                            return(y)}
  else if(t=="above") {y = case_when(x<0 ~ "below",
                            x>0 ~ "above")
                            return(y)}
  else {return("[NB: wrong value for t]")}
}

### Function to remove negative sign from character strings (needed as formatted numbers are not numerical)
abs2 <- function(x) {
  y = gsub("-","",x)
  return(y)
}

### Helper function to ensure legend labels are placed correctly
reverse_legend_labels <- function(plotly_plot) {
  n_labels <- length(plotly_plot$x$data)
  plotly_plot$x$data[1:n_labels] <- plotly_plot$x$data[n_labels:1]
  plotly_plot
}

### Helper function to re-introduce simplified modebar allowing chart download
plotly_modebar <- function(plotly_plot) {
  plotly_plot <- plotly_plot %>% plotly::config(displayModeBar = TRUE) %>% # Allows menu bar such that image can be downloaded
  plotly::config(modeBarButtonsToRemove = c("zoomIn2d", "zoomOut2d","zoom2d","pan2d","autoScale2d","hoverClosestCartesian","hoverCompareCartesian","select2d","lasso2d")) %>% 
  plotly::config(displaylogo = FALSE)
  plotly_plot
}

### To help pull data
  pull_dtpt <- function(dt=NULL,flt=NULL,vr=NULL) {
    dtpt <- dt %>% 
      filter(eval(rlang::parse_expr(flt))) %>% 
      pull(var=vr)
    
    return(dtpt)
  }
 # THe function above can be used as follows:
    # pull_dtpt(dt=paye_nat_long,flt="nationality == 'EU' & region== 'London' & date_day == max(date_day)",vr="employee_count")

################################################################################
### Downloads and data processing
################################################################################

  #.............................................................................
  ### London borough geographical codes
  #.............................................................................
  London_LA_codes <-readxl::read_excel(path = paste0(INPUT,"London borough codes.xlsx"), sheet = "Codes") %>% clean_names()

  #.............................................................................
  ### Fundction for dowloading headline stats
  #.............................................................................
  
  HeadlineDownload <- function( data_series = "NM_59_1", 
                                time_period = c("1996-01","latest"), 
                                save_intermediate = TRUE,
                                geography = Group) {
    
    ### Download both absolute and percentage figures from NOMIS
    
    raw_nomis <- nomis_get_data( id = data_series, 
                                 geography = geography,
                                 time = time_period)
    
    ### Store downloads as dataframes
    
    lfs_stats <- raw_nomis %>% 
      mutate(DATE = as.Date(paste0(DATE, "-01"))) %>% 
      filter(MEASURES %in% c(20207,20100) & !is.na(OBS_VALUE)) %>%
      select( "DATE", "DATE_NAME", "SEX_NAME", "GEOGRAPHY_NAME", "ECONOMIC_ACTIVITY_NAME", "VALUE_TYPE_NAME", "OBS_VALUE","MEASURES","MEASURES_NAME") %>% 
      pivot_wider( names_from = ECONOMIC_ACTIVITY_NAME, values_from = OBS_VALUE) %>%  
      clean_names()
    
    ### Save intermediate dataframes if specified
    
    if( save_intermediate == TRUE) {
      lfs_stats %>% 
        fwrite(file = paste0(INTERMEDIATE,Sys.Date(),"_LFS_update.csv"))
      
    }
    
    data.table::fwrite( as.data.frame(colnames(lfs_stats)), file = paste0(INTERMEDIATE,"/",Sys.Date(),"_colnames.csv"))
    
    return( lfs_stats)
    
  }

  #.............................................................................
  ### Download LFS stats, automatically clean and save
  #.............................................................................
  
  lfs_stats <- HeadlineDownload( time_period = c("2020-01","latest"))

  #.............................................................................
  ### Claimant count data download
  #.............................................................................

  # Define function used
  
  ClaimantCountDownload <- function( data_series = "NM_162_1",
                                     time_period = c("1996-01","latest"),
                                     save_intermediate = TRUE,
                                     geography_v = Group,
                                     measures_v = 20100,
                                     line_or_bar = "line") {
    
    ### Download the claimant count data from NOMIS
    
    raw_nomis <- nomis_get_data(
      id = data_series, 
      geography = geography_v,
      measures = measures_v,
      time = time_period) 
    
    ### Clean data
    
    claimant_count_stats <- raw_nomis %>% 
      mutate( DATE = as.Date(paste0(DATE, "-01"))) %>% 
      filter( !is.na(OBS_VALUE) & MEASURES == measures_v) %>% 
      pivot_wider( names_from = MEASURE_NAME, values_from = OBS_VALUE) %>%
      clean_names() %>% 
      select( date, date_name, geography_name, gender_name, age_name, last_col(4):last_col()) %>% 
      rename( sex_name = gender_name) %>% 
      pivot_longer(cols=c(contains("claimant")),
                   names_to = "measure_name",
                   values_to = "measure_value",
                   values_drop_na=TRUE) # Re-reshape to get rid of NA and format as panel
    
    
    if( line_or_bar == "line")
    {
      claimant_count_stats <- claimant_count_stats %>%
        filter( age_name == "All categories: Age 16+")
    }
    
    
    ### Save intermediate dataframes if specified
    
    if( save_intermediate == TRUE) {
      fwrite(claimant_count_stats, file = paste0(INTERMEDIATE,Sys.Date(),"_",data_series,"_claimant.csv"))
    }
    
    return( claimant_count_stats) 
    
  }

  
  # ----- Download data and make wide and long version -----

  claimant_count_stats_long <- ClaimantCountDownload( data_series = "NM_162_1", 
                                                      time_period = c("2019-12","latest"),
                                                      geography_v = c(2013265927,2092957697), 
                                                      save_intermediate = TRUE, 
                                                      line_or_bar = "bar")  %>% 
    group_by(geography_name,age_name,sex_name,measure_name) %>% 
    mutate(change_month = measure_value - lag(measure_value , n = 1),
           change_month_percent = 100 * (measure_value - lag(measure_value, n = 1))/lag(measure_value, n = 1),
           change_year = measure_value - lag(measure_value, n = 12),
           change_year_percent = 100*(measure_value - lag(measure_value, n = 12))/lag(measure_value, n = 12)) %>% 
    ungroup

  claimant_count_stats_wide <- claimant_count_stats_long %>% # Wide used for some charts
    select(-c(contains("change"))) %>% 
    pivot_wider(names_from = "measure_name", values_from = "measure_value")
  
 # raw_claim <- download.file( url = "https://www.nomisweb.co.uk/api/v01/dataset/NM_162_1.data.csv?geography=1807745125...1807745129,1807745111,1807745112,1807745130...1807745133,1807745113...1807745115,1807745134...1807745137,1807745116,1807745117,1807745138,1807745118,1807745119,1807745139,1807745120,1807745140,1807745141,1807745121,1807745142,1807745122,1807745143,1807745123,1807745124,2092957697,2013265927&date=latestMINUS22-latest&gender=0...2&age=0...2,10,11,3,12...16,4,17...20&measure=1...4&measures=20100&signature=NPK-ffca5e5a461529986a4082:0x6a7726c4d1c34de0ce2aca1289257fd9ef1cc613"), mode = "wb")

# claimant_count_stats_o <-read.csv(paste0(INTERMEDIATE,"2021-11-16_NM_162_1_claimant.csv")) %>% 
#     arrange(date, geography_name,sex_name,age_name)

  # claimant_count_stats <-read.csv(paste0(INPUT,"2021-12-10_claims_NM_162_1.csv")) %>% 
  #   mutate( DATE = as.Date(paste0(DATE, "-01"))) %>% 
  #   filter( !is.na(OBS_VALUE) & MEASURES == 20100) %>% 
  #   pivot_wider( names_from = MEASURE_NAME, values_from = OBS_VALUE) %>%
  #   clean_names() %>% 
  #   select( date, date_name, geography_name, gender_name, age_name, last_col(4):last_col()) %>% 
  #   rename( sex_name = gender_name) %>% 
  #   arrange(date, geography_name,sex_name,age_name)

  #.............................................................................
  ### Download PAYE data, clean and save
  #.............................................................................
  
  # Check if there is a file this month, otherwise take previous month
  
  paye_online_path_now <- paste0("https://www.ons.gov.uk/file?uri=/employmentandlabourmarket/peopleinwork/earningsandworkinghours/datasets/realtimeinformationstatisticsreferencetableseasonallyadjusted/current/",
                             paste0("rtisa",tolower(format(Sys.Date(),'%b%Y'))),
                             ".xlsx")

  paye_online_path_prev <-  paste0("https://www.ons.gov.uk/file?uri=/employmentandlabourmarket/peopleinwork/earningsandworkinghours/datasets/realtimeinformationstatisticsreferencetableseasonallyadjusted/current/",
                                   paste0("rtisa",tolower(format(floor_date(Sys.Date(),"month")-months(1),'%b%Y'))),
                             ".xlsx")
  
  ## Function to assert URL exists
  urlFileExist <- function(url){
    HTTP_STATUS_OK <- 200
    hd <- httr::HEAD(url)
    status <- hd$all_headers[[1]]$status
    list(exists = status == HTTP_STATUS_OK)
  }
  
  # Actual control
  if (urlFileExist(paye_online_path_now)==TRUE) {
    paye_online_path <- paye_online_path_now
  }  else {
    paye_online_path <- paye_online_path_prev
  }
  
  #Download latest file
  raw_paye <- download.file( url = paye_online_path,
                             destfile = paste0(INTERMEDIATE,Sys.Date(),"_raw_paye.xlsx"),
                             mode = "wb")
  
  paye_stats <- readxl::read_excel(path = paste0(INTERMEDIATE,Sys.Date(),"_raw_paye.xlsx"), sheet = "7. Employees (NUTS1)", skip = 6) %>% 
                  select( Date, London, UK) %>%
                  clean_names() %>% 
                  mutate( new_date = dmy(paste0("01",date)),
                          lon_change_since_feb20 = case_when( new_date <= as.Date("2020-02-01") ~ NaN,
                                                        new_date > as.Date("2020-02-01") ~ london - london[ new_date == "2020-02-01"]),
                          lon_percentage_change_since_feb20 = case_when( new_date <= as.Date("2020-02-01") ~ NaN,
                                                        new_date > as.Date("2020-02-01") ~ 100*(london - london[ new_date == "2020-02-01"])/london[ new_date == "2020-02-01"]),
                          lon_month_change = case_when( new_date - min(new_date) < 365 ~ NaN,
                                                                    new_date - min(new_date) > 365 ~ 
                                                                      (london - lag(london, n = 1))),
                          lon_month_percentage_change = case_when( new_date - min(new_date) < 365 ~ NaN,
                                                                    new_date - min(new_date) > 365 ~ 
                                                                      100*(london - lag(london, n = 1))/lag(london, n = 1)),
                          uk_change_since_feb20 = case_when( new_date <= as.Date("2020-02-01") ~ NaN,
                                                           new_date > as.Date("2020-02-01") ~ uk - uk[ new_date == "2020-02-01"]),
                          uk_percentage_change_since_feb20 = case_when( new_date <= as.Date("2020-02-01") ~ NaN,
                                                           new_date > as.Date("2020-02-01") ~ 100*(uk - uk[ new_date == "2020-02-01"])/uk[ new_date == "2020-02-01"]),
                          lon_annual_change = case_when( new_date - min(new_date) < 365 ~ NaN,
                                                                    new_date - min(new_date) > 365 ~ 
                                                                      (london - lag(london, n = 12))),
                          lon_annual_percentage_change = case_when( new_date - min(new_date) < 365 ~ NaN,
                                                                    new_date - min(new_date) > 365 ~ 
                                                                      100*(london - lag(london, n = 12))/lag(london, n = 12)),
                          uk_annual_percentage_change = case_when( new_date - min(new_date) < 365 ~ NaN,
                                                                    new_date - min(new_date) > 365 ~ 
                                                                      100*(uk - lag(uk, n = 12))/lag(uk, n = 12)))
  
  paye_nuts2_stats <- readxl::read_excel(path = paste0(INTERMEDIATE,Sys.Date(),"_raw_paye.xlsx"), sheet = "11. Employees (NUTS2)", skip = 6) %>% 
        mutate( new_date = dmy(paste0("01",Date)))  %>% 
        merge(paye_stats,by="new_date")   %>% 
    rename(London=london) %>% 
    select( Date, new_date, "Inner London - West",	"Inner London - East",	"Outer London - East and North East",	"Outer London - South",	"Outer London - West and North West", "London","UK") %>%
    pivot_longer(cols="Inner London - West":UK,names_to = "area", values_to = "emp_level") %>% 
    group_by(area) %>% 
    mutate( 
      change_perc_since_feb20 = case_when( new_date <= as.Date("2020-02-01") ~ NaN,
                                      new_date > as.Date("2020-02-01") ~ 100*(emp_level - emp_level[ new_date == "2020-02-01"])/(emp_level[ new_date == "2020-02-01"])),
      annual_change = case_when(new_date - min(new_date) < 365 ~ NaN,
                                      new_date - min(new_date) > 365 ~ 100*(emp_level - lag(emp_level, n= 12))/lag(emp_level, n=12))) %>% 
    clean_names()

  
  ## -------------------------------------------------------------------------##
  # PAYE data has unique quarterly data on following months:
  ## - (a) By Age*NUTS1: 1,4,7,10
  ## - (b) By Industry*NUTS1: 2,5,8,11
  ## - (c) By LA: 3,6,9,12
  
  # These need to be updated as such:
  ## If division remainder is 0, then (c); if 1 then (a), if 2 then (b)
  ## If it is LMU release day, ensure data is updated

  current_month <- lubridate::month(Sys.Date())
  month_remainder <- current_month %% 3
  release_dates <- c("2022-02-15","2022-03-15","2022-04-12","2022-05-17","2022-06-14","2022-07-19","2022-08-16","2022-09-13","2022-10-11")
  
  if (Sys.Date() %in% lapply(release_dates,as.Date)) {
    
    if (month_remainder == 0) {
      dlg_message("Remember to update the following: PAYE LA data")$res
    } else if (month_remainder==1) {
      dlg_message("Remember to update the following: PAYE Age*NUTS1 data")$res
    } else if (month_remainder==2) {
      dlg_message("Remember to update the following: PAYE Industry*NUTS1 data")$res
    }
    
  }
  #### (a) Each quarter, process data on PAYE employee age on regional level
  #### NB: MUST BE UPDATED MANUALLY TO NEWER DATA  
  
  latest_paye_nuts1age_name <- "2022-01-18_raw_paye.xlsx"
  
    paye_age_stats_uk <- readxl::read_excel(path = paste0(INTERMEDIATE,latest_paye_nuts1age_name), sheet = "28. Employees (Age)", skip = 5) %>%
    mutate(new_date = dmy(paste0("01",Date))) %>% 
    select( Date, new_date, "0-17","18-24","25-34","35-49","50-64","65+") %>% 
    pivot_longer(cols="0-17":"65+",names_to = "age_group", values_to = "emp_level") %>% 
    group_by(age_group) %>% 
    mutate( 
      area = "United Kingdom",
      change_perc_since_feb20 = case_when( new_date <= as.Date("2020-02-01") ~ NaN,
                                           new_date > as.Date("2020-02-01") ~ 100*(emp_level - emp_level[ new_date == "2020-02-01"])/(emp_level[ new_date == "2020-02-01"])),
      annual_change = case_when(new_date - min(new_date) < 365 ~ NaN,
                                new_date - min(new_date) >= 365 ~ 100*(emp_level - lag(emp_level, n= 12))/lag(emp_level, n=12)),
      age_group = paste0("Aged: ", age_group)) %>% 
    clean_names() 
    
    
  paye_nuts1age_stats <- readxl::read_excel(path = paste0(INTERMEDIATE,latest_paye_nuts1age_name), sheet = "32. Employees (NUTS1Age)", skip = 6) %>%
    mutate(new_date = dmy(paste0("01",Date))) %>% 
    select( Date, new_date, "London: 0-17","London: 18-24","London: 25-34","London: 35-49","London: 50-64","London: 65+") %>% 
    pivot_longer(cols="London: 0-17":"London: 65+",names_to = "age_group", values_to = "emp_level") %>% 
    group_by(age_group) %>% 
    mutate( 
      age_group = gsub("London: ","",x=age_group),
      area = "London",
      change_perc_since_feb20 = case_when( new_date <= as.Date("2020-02-01") ~ NaN,
                                           new_date > as.Date("2020-02-01") ~ 100*(emp_level - emp_level[ new_date == "2020-02-01"])/(emp_level[ new_date == "2020-02-01"])),
      annual_change = case_when(new_date - min(new_date) < 365 ~ NaN,
                                new_date - min(new_date) >= 365 ~ 100*(emp_level - lag(emp_level, n= 12))/lag(emp_level, n=12)),
      age_group =paste0("Aged: ", age_group)) %>%
    clean_names()   %>%  rbind(paye_age_stats_uk) %>% 
    arrange(area,new_date)
  
  
  newest_period_paye_age <- format(max(paye_nuts1age_stats$new_date),"%B %Y")
  
  
  
  #### (b) Each quarter, process data on PAYE employees by industry at region level. 
  #### NB: MUST BE UPDATED MANUALLY TO NEWER DATA
  
  latest_paye_nuts1ind_name <- "2022-02-15_raw_paye.xlsx"
  
  paye_ind_stats_uk <- readxl::read_excel(path = paste0(INTERMEDIATE,latest_paye_nuts1ind_name), sheet = "23. Employees (Industry)", skip = 6) %>% 
    mutate(new_date = dmy(paste0("01",Date)),
           geography_name="United Kingdom") %>% 
    select(-c("UK"))  %>% 
    pivot_longer(cols="Agriculture, forestry and fishing":"Households and Extraterritorial",
                 names_to = "industry_name", values_to = "emp_level") %>% 
    clean_names() 
    
  
  paye_nuts1ind_stats <- readxl::read_excel(path = paste0(INTERMEDIATE,latest_paye_nuts1ind_name), sheet = "36. Employees (NUTS1Sector)", skip = 6) %>% 
    mutate(new_date = dmy(paste0("01",Date)),
           geography_name="London") %>% 
    select(c(Date,new_date,geography_name,contains("London")))  %>%
    rename_with(~str_replace(.x,pattern="London: ",replacement=""))%>% 
    pivot_longer(cols="Agriculture, forestry and fishing":"Households and Extraterritorial",
                 names_to = "industry_name", values_to = "emp_level") %>% 
    clean_names() %>% 
    rbind(paye_ind_stats_uk) %>% 
    group_by(geography_name,industry_name) %>% 
    mutate(
      change_perc_since_feb20 = case_when( new_date <= as.Date("2020-02-01") ~ NaN,
                                           new_date > as.Date("2020-02-01") ~ 100*(emp_level - emp_level[new_date == "2020-02-01"])/(emp_level[ new_date == "2020-02-01"])),
      annual_change = case_when(new_date - min(new_date) < 365 ~ NaN,
                                new_date - min(new_date) >= 365 ~ 100*(emp_level - lag(emp_level, n= 12))/lag(emp_level, n=12)),
      industry_name_simple = case_when(
        industry_name == "Public administration and defence; social security" ~ "Public admin & defence",
        industry_name == "Health and social work" ~ "Health",
        industry_name == "Finance and insurance" ~ "Finance & insurance",
        industry_name == "Professional, scientific and technical" ~ "Professional services",
        industry_name == "Wholesale and retail; repair of motor vehicles" ~ "Retail",
        industry_name == "Other service activities" ~ "Other services",
        industry_name == "Transportation and storage" ~ "Transport & storage",
        industry_name == "Arts, entertainment and recreation" ~ "Arts & recreation",
        industry_name == "Information and communication" ~ "Information & communication",
        industry_name == "Administrative and support services" ~ "Administration",
        industry_name == "Accommodation and food service activities" ~ "Hospitality",
        industry_name == "Water supply, sewerage and waste" ~ "Water",
        TRUE ~ industry_name)) %>% 
    arrange(new_date,geography_name,industry_name_simple) %>% 
    relocate(date,new_date,geography_name,industry_name_simple) %>% 
    ungroup()
  
  newest_period_paye_ind <- format(max(paye_nuts1ind_stats$new_date),"%B %Y")
  
  #### (c) Each quarter, process data on PAYE at borough level and merge with geographical codes. 
  #### NB: MUST BE UPDATED MANUALLY TO NEWER DATA
  
  latest_paye_locauth_name <- "2022-03-15_raw_paye.xlsx"

  paye_la_stats <- readxl::read_excel(path = paste0(INTERMEDIATE,latest_paye_locauth_name), sheet = "19. Employees (LA)", skip = 7) %>% 
    mutate(new_date = dmy(paste0("01",Date))) %>% 
    select( Date, new_date, "City of London", "Barking and Dagenham", "Barnet", "Bexley", "Brent", "Bromley", "Camden", "Croydon", "Ealing", "Enfield", "Greenwich", "Hackney", "Hammersmith and Fulham", "Haringey", "Harrow", "Havering", "Hillingdon", "Hounslow", "Islington", "Kensington and Chelsea", "Kingston upon Thames", "Lambeth", "Lewisham", "Merton", "Newham", "Redbridge", "Richmond upon Thames", "Southwark", "Sutton", "Tower Hamlets", "Waltham Forest", "Wandsworth", "Westminster") %>% 
    pivot_longer(cols="City of London":"Westminster",names_to = "area", values_to = "emp_level") %>% 
    group_by(area) %>% 
    mutate( 
      change_perc_since_feb20 = case_when( new_date <= as.Date("2020-02-01") ~ NaN,
                                           new_date > as.Date("2020-02-01") ~ 100*(emp_level - emp_level[ new_date == "2020-02-01"])/(emp_level[ new_date == "2020-02-01"])),
      annual_change = case_when(new_date - min(new_date) < 365 ~ NaN,
                                new_date - min(new_date) >= 365 ~ 100*(emp_level - lag(emp_level, n= 12))/lag(emp_level, n=12))) %>% 
    clean_names() %>%  merge(London_LA_codes,by='area') %>% 
    arrange(area,new_date)
  
  newest_period_paye_la <- format(max(paye_la_stats$new_date),"%B %Y")
  
  
  ## -------------------------------------------------------------------------##
  ## PAYE supplement with employees by nationality
  
  paye_nat <- readxl::read_excel(path = paste0(INTERMEDIATE,"nuts1bynationalitynsaandsamar2022.xlsx"), sheet = "(2.5) NUTS1 by Nat - SA", skip = 8) %>% 
    pivot_longer(cols = !Date, names_to = c("nationality","region"), names_sep = "_",values_to = "employee_count") %>% 
    mutate(date_day = lubridate::my(Date)) %>% 
    clean_names()
  
  paye_nat_uk <- paye_nat %>% 
    group_by(nationality,date,date_day) %>% 
    summarise(employee_count = sum(employee_count)) %>% 
    mutate(region= "United Kingdom")
  
  paye_nat_long  <-  paye_nat %>% 
    rbind(paye_nat_uk) %>% 
    arrange(date_day,region,nationality) %>% 
    mutate(index_feb20 = case_when(date_day<"2020-02-01" ~ NA_real_,
                                   TRUE ~ 100* (employee_count)/(employee_count[date_day=="2020-02-01"])),
           d_change_feb20 =  case_when(date_day<="2020-02-01" ~ NA_real_,
                                       TRUE ~ employee_count-employee_count[date_day=="2020-02-01"]),
           p_change_feb20 =  case_when(date_day<="2020-02-01"  ~ NA_real_,
                                       TRUE ~ 100* (employee_count-employee_count[date_day=="2020-02-01"])/(employee_count[date_day=="2020-02-01"])))
  
  
  #.............................................................................
  ### Download Workforce Jobs data
  #.............................................................................
  
  
  wfj_stats <- nomis_get_data(id = "NM_130_1", time = c("2010-01", "latest"), geography = Group,
                              industry=c(37748736,150994945:150994964)) %>% 
    clean_names()   %>% 
    select(c("date","date_name","geography","geography_name","industry","industry_code","industry_name","item","item_name","measures","measures_name","obs_value","record_count")) %>% 
    mutate(new_date = dmy(paste0("01",date_name)),
           quarter =  paste0("Q",quarter(new_date)),
           quarter_text = case_when(quarter=="Q1" ~ "first",
                                    quarter=="Q2" ~ "second",
                                    quarter=="Q3" ~ "third",
                                    quarter=="Q4" ~ "fourth")) %>% 
    group_by(industry_name, geography_name, item_name,measures_name) %>% #grouping makes calculating lag values easier
    mutate( 
      change_perc_since_dec19 = case_when( new_date <= as.Date("2019-12-01") | measures_name == "Percent" ~ NaN,
                                           new_date > as.Date("2019-12-01") ~ 100*(obs_value - obs_value[ new_date == "2019-12-01"])/(obs_value[ new_date == "2019-12-01"])),
      change_since_dec19 = case_when( new_date <= as.Date("2019-12-01") | measures_name == "Percent" ~ NaN,
                                      new_date > as.Date("2019-12-01") ~ (obs_value - obs_value[ new_date == "2019-12-01"])),
      change_perc_quar = case_when(  measures_name == "Percent" ~ NaN,
                                     TRUE ~ 100*(obs_value - lag(obs_value, n = 1))/(lag(obs_value, n = 1))),
      change_quar = case_when(  measures_name == "Percent" ~ NaN,
                                TRUE ~ (obs_value - lag(obs_value, n = 1))),
      industry_name = case_when(industry_name == "Total" ~ "Total",
                                TRUE ~  str_sub(industry_name,5,-1))) %>% #trims the name 
    mutate( industry_name_simple = case_when(
      industry_name == "Public administration and defence; compulsory social security" ~ "Public admin & defence",
      industry_name == "Human health and social work activities" ~ "Health",
      industry_name == "Financial and insurance activities" ~ "Finance & insurance",
      industry_name == "Professional, scientific and technical activities" ~ "Professional services",
      industry_name == "Wholesale and retail trade; repair of motor vehicles and motorcycles" ~ "Retail",
      industry_name == "Other service activities" ~ "Other services",
      industry_name == "Transportation and storage" ~ "Transport & storage",
      industry_name == "Arts, entertainment and recreation" ~ "Arts & recreation",
      industry_name == "Information and communication" ~ "Information & communication",
      industry_name == "Administrative and support service activities" ~ "Administration",
      industry_name == "Accommodation and food service activities" ~ "Hospitality",
      industry_name == "Water supply; sewerage, waste management and remediation activities" ~ "Water",
      TRUE ~ industry_name)) %>% 
    group_by( geography_name,new_date,item_name) %>% # To get ranks of industries within date-geography groups
    mutate(perc_change_dec19_within_date_rank = case_when( industry_name == "Total" ~ NA_integer_, 
                                                           TRUE  ~ dense_rank(desc(change_perc_since_dec19))), 
           perc_change_quart_within_date_rank = case_when( industry_name == "Total" ~ NA_integer_, 
                                                           TRUE  ~ dense_rank(desc(change_perc_quar)))) %>% 
    ungroup()%>%  #ungroup to ease future manipulations
    arrange(date,geography,item,measures,industry)
  
  #wfj_stats <- clean_names(wfj_stats,case="snake") #Converting column names to lower

  fwrite(wfj_stats, file = paste0(INTERMEDIATE,Sys.Date(),"_wfj.csv"), na = "NaN")
  
  #-----------------------------------------------------------------------------
  ### Define GLA line chart function
  #-----------------------------------------------------------------------------
  
  GLALineChart <- function( data_set = lfs_stats, ### This function has been written to work specifically with the NM_59_1 dataset
                        lfs_or_claims = "lfs",
                        x_var = date, ### Assumed that this will be a line chart with dates on the x-axis
                        y_var = NULL, ### See document "[today's date]colnames.csv" for variables available for charting
                        geography = c("London", "United Kingdom"),
                        suffix = "%",
                        y_limits = c(0,100),
                        nudge_y = NULL,
                        title = NULL,
                        subtitle = paste0("Latest data for period ", new_release),
                        caption = "",
                        chart_name = NULL) {
      
    pal <- gla_pal(gla_theme = "default", palette_type = "highlight", n = c(1, 1))
    theme_set(theme_gla(gla_theme = "default"))
    
    ### Create dataset for charting
    
    if( lfs_or_claims == "lfs") {
    
    for_charting <- data_set %>% 
      filter( geography_name %in% geography & value_type_name == "Level" & sex_name == "Total" & measures == 20207)
    } else {
    for_charting <- data_set %>% 
      filter( geography_name %in% geography & sex_name == "Total" & !is.na({{y_var}}))
    }
    ### Set x-axis format
  
    ### Plot charts
    
    new_release <- for_charting$date_name[for_charting$date == max(for_charting$date)]
    
    glaplot <- for_charting %>%
      ggplot(mapping = aes(x = {{x_var}}, y = {{y_var}}, 
                           colour = geography_name, group = geography_name,
                           text = paste(
                          geography_name, "\n",
                          format({{x_var}},'%B %Y'), "\n",
                          "Rate: ", perc_form({{y_var}}),"%", "\n",
                          sep = ""))) +
      ggla_line(aes(size= geography_name)) +
      scale_size_manual(values = c(4 * mm_to_pt, 2 * mm_to_pt)) +
      scale_colour_manual(values = pal) +
      ggla_highlight(filter_type = "end") +
      ggla_highlight(mapping = aes(label = paste0({{y_var}}, "%")),
                     geom = GeomGLATextHighlight, filter_type = "end",  size = 4.5,
                     position = position_nudge(y = nudge_y),check_overlap = TRUE)+
      coord_cartesian(clip = 'off') +
      scale_y_continuous(expand = c(0, 0), labels = dollar_format(prefix = "", 
                                                                  suffix = suffix, 
                                                                  largest_with_cents = 1), 
                                                                  limits = y_limits) +
      scale_x_date( date_breaks = "2 months",
                    date_labels = "%b'%y",
                    expand = expansion( mult = c(0.1,0.02))) +
      theme(plot.margin = unit(c(1,1,1,1), "cm"))+
      labs(title = title,
           subtitle = subtitle,
           caption = caption) +
      theme(plot.caption = element_text(color = rgb(166,166,166,maxColorValue = 255)))
    
    ggsave( paste0(IMAGES,Sys.Date(),"_",chart_name,".svg"), device = "svg", width = 8, height = 8, units = "in")
    ggsave( paste0(IMAGES,Sys.Date(),"_",chart_name,".png"), device = "png", width = 8, height = 8, units = "in")
    
    return(glaplot)
    
  }

```

```{r, main_figures ,echo=FALSE}
  ### Create all the figures needed for the section below
  # Name format:
  # _d_: difference, _p_: difference in percent, _pp_: difference in percentage points, _uk: UK level stat, London otherwise

# Paye
paye_d_month <- value_form(paye_stats %>% filter( new_date == max(new_date)) %>% select(lon_month_change),s= signif_thous)

paye_p_month <- perc_form(paye_stats %>% filter( new_date == max(new_date)) %>% select(lon_month_percentage_change),d = digits_perc)

paye_newmonth <- paye_stats %>% filter( new_date == max(new_date)) %>% mutate(new_date=format(new_date,'%B %Y')) %>%  select(new_date)

paye_lastmonth <- paye_stats  %>% mutate(lag_month = lag(new_date, n = 1)) %>% mutate(lag_month2 = format(lag_month, '%B')) %>% filter( new_date == max(new_date)) %>%  select(lag_month2)

paye_p_feb20 <- perc_form(paye_stats %>% filter( new_date == max(new_date)) %>% select(lon_percentage_change_since_feb20),d = digits_perc)

# LFS

lfs_newmonth <- lfs_stats %>% mutate(month = format(date, '%B %Y')) %>% filter(sex_name == "Total"  & geography_name == "London" & value_type_name == "Level" & date == max(date) & measures == 20207) %>%  select(month)
  
lfs_lastmonth <- lfs_stats  %>% mutate(lag_month = ymd(date) %m+% - months(2)) %>% mutate(lag_month2 = format(lag_month, '%B')) %>% filter(sex_name == "Total"  & geography_name == "London" & value_type_name == "Level" & date == max(date) & measures == 20207) %>%  select(lag_month2)


# Unem

unem_rate <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London" & value_type_name == "Level" & measures == 20207) %>% select(total_unemployed_aged_16_and_over),d = digits_perc)
  
unem_pp_quart <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London" & value_type_name == "Change on quarter" & measures == 20207) %>% select(total_unemployed_aged_16_and_over),d = digits_perc)


unem_pp_year <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London" & value_type_name == "Change on year" & measures == 20207)  %>% select(total_unemployed_aged_16_and_over),d = digits_perc)
  
unem_rate_uk <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom" & value_type_name == "Level" & measures == 20207) %>% select(total_unemployed_aged_16_and_over),d = digits_perc)


# Employment

emp_rate <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London" & value_type_name == "Level" & measures == 20207)  %>% select(total_in_employment_aged_16_to_64),d = digits_perc)

emp_pp_quart <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London"  & value_type_name == "Change on quarter" & measures == 20207)  %>% select(total_in_employment_aged_16_to_64),d = digits_perc)

emp_pp_year <-perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London"  & value_type_name == "Change on year" & measures == 20207) %>% select(total_in_employment_aged_16_to_64),d = digits_perc)

emp_rate_uk <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom" & value_type_name == "Level" & measures == 20207)  %>% select(total_in_employment_aged_16_to_64),d = digits_perc)

# Inactivity

inact_rate <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London" & value_type_name == "Level" & measures == 20207)  %>% select(total_economically_inactive_aged_16_to_64),d = digits_perc)

inact_pp_quart <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London"  & value_type_name == "Change on quarter" & measures == 20207)  %>% select(total_economically_inactive_aged_16_to_64),d = digits_perc)

inact_pp_year <-perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London"  & value_type_name == "Change on year" & measures == 20207) %>% select(total_economically_inactive_aged_16_to_64),d = digits_perc)

inact_rate_uk <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom" & value_type_name == "Level" & measures == 20207)  %>% select(total_economically_inactive_aged_16_to_64),d = digits_perc)

# WFJ

wfj_year <-  wfj_stats %>% filter(new_date == max(new_date)) %>% filter(row_number()==1) %>% mutate(year=year(new_date)) %>% select(year)

wfj_newquarter <- wfj_stats %>% ungroup()  %>% filter(new_date == max(new_date)) %>% filter(row_number()==1) %>% select(quarter_text) 
  
wfj_oldquarter <-  wfj_stats %>% filter(new_date == (ymd(max(new_date)) %m+% - months(3))) %>% filter(row_number()==1) %>% select(quarter_text)

wfj_d_quart <- value_form(wfj_stats%>% ungroup()  %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd(max(new_date)) %m+% - months(3)) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value),s = signif_thous)

wfj_p_quart <- perc_form(100*(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd(max(new_date)) %m+% - months(3)) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value))/(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd(max(new_date)) %m+% - months(3)) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value)),d = digits_perc)

wfj_d_dec19 <- value_form(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd("2019-12-01")) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value),s = signif_thous)

wfj_p_dec19 <- perc_form(100*(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd("2019-12-01")) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value))/(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd("2019-12-01")) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value)),d = digits_perc)

```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<!-- Summary of findings - MANUALLY ADJUST -->

<br/>

## Summary of key points

<br/>

-   The most timely estimate of **payrolled employees** shows  `r condi_text(paye_d_month,"noun")` of around `r abs2(paye_d_month)`, or `r abs2(paye_p_month)`%, in London between `r paye_lastmonth` and `r paye_newmonth` and is `r abs2(paye_p_feb20)`% `r condi_text(paye_p_feb20,"above")` pre-pandemic (February 2020) levels. 

-   The **employment rate** in London was estimated at `r emp_rate`% for the three months ending `r lfs_newmonth`, `r condi_text(emp_pp_quart,"up")` `r abs2(emp_pp_quart)` percentage points (pp) on the previous quarter and `r condi_text(emp_pp_year,"up")` `r abs2(emp_pp_year)`pp on the same period in the previous year. London's employment rate remains close to the UK average (`r emp_rate_uk`%).

-   The **unemployment rate** continued to fall from its pandemic peak but is still higher than the UK average. London's unemployment rate was estimated at `r unem_rate`%, `r condi_text(unem_pp_quart,"up")` `r abs2(unem_pp_quart)`pp on the quarter and `r condi_text(unem_pp_year,"up")` `r abs2(unem_pp_year)`pp from a year earlier. The UK average was `r unem_rate_uk`%.

-   London's **inactivity rate** (the measure of those not looking and/or not available to work) was estimated at `r inact_rate`%. While this was `r condi_text(inact_pp_year,"up")` `r abs2(inact_pp_year)`pp on the previous year, the inactivity rate `r condi_text(inact_pp_quart,"rise")` `r abs2(inact_pp_quart)`pp on the previous quarter and remains lower than the UK-wide estimate of `r inact_rate_uk`%.

-    The total number of **workforce jobs** `r condi_text(wfj_d_quart,"increase")` by `r abs2(wfj_d_quart)`, or `r abs2(wfj_p_quart)`%, between the `r wfj_oldquarter` and `r  wfj_newquarter` quarter of `r wfj_year`, and is `r abs2(wfj_d_dec19)`, or `r abs2(wfj_p_dec19)`%, `r condi_text(wfj_d_dec19,"above")` pre-pandemic (December 2019) levels. While employee jobs have recovered, self-employed jobs remain 15.5% below pre-pandemic levels.

<a href="#top">Back to top</a>

<!-- Standard intro page with caveats -->

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>


## Measuring the labour market

<br/>

The ONS has published its latest [labour market update](https://www.ons.gov.uk/employmentandlabourmarket/peopleinwork/employmentandemployeetypes) covering Labour Force Survey (LFS) data for the three months ending `r lfs_newmonth`. This bulletin presents the latest headline labour market estimates for London.

Please note:

-   Many of the statistics presented here are estimates based on surveys, and as such have a margin of error -- known as sampling variability.

-   They also cover different reference periods or count dates, and have therefore been impacted differently by the spread of COVID-19.

More information and previous GLA Economics analyses can be found on our labour market analysis [page](https://data.london.gov.uk/dataset/gla-economics-covid-19-labour-market-analysis).

More information on the data used and a glossary of terms can be found on the [ONS website](https://www.ons.gov.uk/employmentandlabourmarket/peopleinwork/employmentandemployeetypes/methodologies/aguidetolabourmarketstatistics).

```{r,paye figures, echo=FALSE}

paye_level <- value_form(paye_stats %>% filter( new_date == max(new_date)) %>% mutate(lon_round=london/1000000) %>% select(lon_round),s=signif_mil, d=digits_mil)

paye_d_year <- value_form(paye_stats %>% filter( new_date == max(new_date))  %>% select(lon_annual_change),s=signif_thous)
  
paye_p_year <- perc_form(paye_stats %>% filter( new_date == max(new_date)) %>% select(lon_annual_percentage_change),d = digits_perc)

paye_lastyearmonth <- paye_stats  %>% mutate(lag_month = lag(new_date, n = 12)) %>% mutate(lag_month2 = format(lag_month, '%B %Y')) %>% filter( new_date == max(new_date)) %>%  select(lag_month2)
  
paye_p_year_uk <- perc_form(paye_stats %>% filter( new_date == max(new_date)) %>% select(uk_annual_percentage_change),d = digits_perc)

paye_d_feb20 <- value_form(paye_stats %>% filter( new_date == max(new_date))  %>% select(lon_change_since_feb20),s=signif_thous)

paye_p_feb20 <- perc_form(paye_stats %>% filter( new_date == max(new_date)) %>% select(lon_percentage_change_since_feb20),d = digits_perc)

paye_d_feb20_uk <- value_form(paye_stats %>% filter( new_date == max(new_date))  %>% select(uk_change_since_feb20),s=signif_thous)

paye_p_feb20_uk <- perc_form(paye_stats %>% filter( new_date == max(new_date)) %>% select(uk_percentage_change_since_feb20),d = digits_perc)

```

<a href="#top">Back to top</a>

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

## Payrolled employees {.tabset}

<br/>

The count of payrolled employees from HMRC's Pay As You Earn (PAYE) RTI dataset offers a timely measure of labour market trends:

-   Early estimates indicate that there were around `r paye_level` million payrolled employees living in London in `r paye_newmonth`, `r condi_text(paye_d_month,"noun")` of around `r abs2(paye_d_month)`, or `r abs2(paye_p_month)`%, on the previous month.

-   The latest London estimate represents `r condi_text(paye_d_year,"noun")` of `r abs2(paye_d_year)`, or `r abs2(paye_p_year)`%, on the previous year (`r paye_lastyearmonth`) against a UK average increase of `r paye_p_year_uk`% (see charts).

-   Relative to pre-pandemic (February 2020) levels, the number of payrolled employees in London was `r condi_text(paye_d_feb20,"up")` by `r abs2(paye_d_feb20)`, or `r abs2(paye_p_feb20)`%. This compares to `r condi_text(paye_p_feb20_uk,"noun")` of `r abs2(paye_p_feb20_uk)`% across the UK on average.

<a href="#top">Back to top</a>

### Change on previous year

```{r PAYE chart, echo=FALSE,fig.cap = payecap, out.width='100%' }

pal <- gla_pal(gla_theme = "default", palette_type = "highlight", n = c(1, 1))
theme_set(theme_gla(gla_theme = "default"))
  
  paye_stats_chart <- paye_stats %>%
    pivot_longer(cols = ends_with("_change"), names_to = "geography", values_to = "change" ) %>%
    mutate( geography_name = case_when( grepl("lon",geography) ~ "London",
                                        grepl("uk",geography) ~ "United Kingdom")) %>%
    mutate(var_type = sub(".*?_",'',geography)) %>% 
    filter( new_date >= "2020-01-01" & var_type == "annual_percentage_change") %>%
    ggplot(mapping = aes(x = new_date, y = change,
                         colour = geography_name, group = geography_name,
                         text = paste(
                           geography_name, "\n",
                           format(new_date,'%B %Y'), "\n",
                           "Change: ", perc_form(change),"%", "\n",
                           sep = ""))) +
    ggla_line(aes(size= geography_name)) +
    scale_size_manual(values = c(4 * mm_to_pt, 2 * mm_to_pt)) +
    scale_colour_manual(values = pal) +
     ggla_highlight(filter_type = "end") +
     ggla_highlight(mapping = aes(label = paste0(format(change, digits = 1),"%")),
                    geom = GeomGLATextHighlight, filter_type = "end",  size = 4.5,
                    position = position_nudge(y = c(+0.4,+0.4)))+
    coord_cartesian(clip = 'off') +
    geom_hline(aes(yintercept=0), colour="gray45") +
    scale_y_continuous(expand = c(0, 0), labels = dollar_format(prefix = "",
                                                                suffix = "%",
                                                                largest_with_cents = 1),
                       limits = c(-6,8)) +
    scale_x_date( date_breaks = "2 months",
                  date_labels = "%b'%y",
                  expand = expansion( mult = c(0.1,0.02))) +
    theme(plot.margin = unit(c(1,1,1,1), "cm"))+
    labs(title =  "Payrolled employees, change on previous year",
         subtitle = paste0("Latest data for period ", paye_stats$date[paye_stats$new_date == max(paye_stats$new_date)]),
         caption = "\nSource: HM Revenue and Customs – Pay As You Earn Real Time Information.\n\nNote: Estimates are based on where employees live.")+
    theme(plot.caption = element_text(color = rgb(166,166,166,maxColorValue = 255)))

ggsave( paste0(IMAGES,Sys.Date(),"_","paye",".svg"), device = "svg", width = 8, height = 8, units = "in")
ggsave( paste0(IMAGES,Sys.Date(),"_","paye",".png"), device = "png", width = 8, height = 8, units = "in")

                     
#knitr::include_graphics(paste0(IMAGES,Sys.Date(),"_","paye",".svg"))

ggplotly(paye_stats_chart,tooltip = "text") %>% 
  ggla_plotly_settings()    %>% 
  layout(title = list(text = paste0("<b>","Payrolled employees, change on previous year","</b>",
                                    "<br>",
                                    "<sup>",
                                    paste0("Latest data for period ", paye_stats$date[paye_stats$new_date == max(paye_stats$new_date)]),
                                    "</sup>",
                                    "<br>"),
                      font = list(size = 22),
                      y = .95, xref = "plot"),
         xaxis = list(tickfont = list(size = 15)),
         yaxis = list(tickfont = list(size = 15)),
         legend = list(font = list(size = 15),title=list(text="")),
         hovermode = "x") %>% 
  plotly_modebar

payecap <- paste0("Source: HM Revenue and Customs – Pay As You Earn Real Time Information.","<br>","<br>", "Note: Estimates are based on where employees live.")

```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

### Change since February 2020, NUTS2 regions

```{r paye bar feb20, , echo=FALSE,fig.cap = payebarcapfeb, out.width='100%'}


pal <- gla_pal(gla_theme = "default", palette_type = "highlight", n = c(1, 1))
theme_set(theme_gla(gla_theme = "default"))

paye_stats_barchart_feb20 <- paye_nuts2_stats %>%
  filter( new_date == max(new_date)) %>%
  ungroup() %>% 
  mutate(chart_ranking = case_when( area == "UK" ~ 1,
                                    area == "London" ~ 2,
                                    TRUE ~ dense_rank(desc(change_perc_since_feb20))+2)) %>% 
  ggplot(mapping = aes(x =  reorder(area, -chart_ranking), y = change_perc_since_feb20,
                       colour = area, group = area, fill=area,
                       text = paste(
                         area, "\n",
                         "Change: ", perc_form(change_perc_since_feb20),"%", "\n",
                         sep = ""))) +
  geom_bar(stat="identity",width=0.5)+
  geom_text(aes(label=paste(format(round(change_perc_since_feb20,1), digits = 2), "%"),
                y = change_perc_since_feb20-0.13*((change_perc_since_feb20)/abs(change_perc_since_feb20))),
            color=c("white","white","white","white","white","white","black"))+ #cannot use vjust with ggplotly
  scale_color_manual(values = c(pal[1],pal[1],pal[1],pal[1],pal[1],pal[1],pal[2]), aesthetics = "colour") +
  scale_fill_manual(values = c(pal[1],pal[1],pal[1],pal[1],pal[1],pal[1],pal[2]), aesthetics =  "fill") +
  coord_cartesian(clip = 'off') +
  geom_hline(aes(yintercept=0), colour="gray45") +
  scale_y_continuous(labels = dollar_format(prefix = "",
                                            suffix = "%",
                                            largest_with_cents = 1),
                     limits = c(0,3)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme(plot.margin = unit(c(1,1,1,1), "cm"),axis.text=element_text(size=12))+
  labs(title =  "Payrolled employees, change since February 2020 by NUTS2 region",
       subtitle = paste0("Latest data for period ", paye_nuts2_stats$date[paye_nuts2_stats$new_date == max(paye_nuts2_stats$new_date)]),
       caption = "\nSource: HM Revenue and Customs – Pay As You Earn Real Time Information.\n\nNote: Estimates are based on where employees live.") + 
  theme(legend.position = "none",
        plot.caption = element_text(color = rgb(166,166,166,maxColorValue = 255)))

ggsave( paste0(IMAGES,Sys.Date(),"_","paye_nuts2_feb20",".svg"), device = "svg", width = 8, height = 8, units = "in")
ggsave( paste0(IMAGES,Sys.Date(),"_","paye_nuts2_feb20",".png"), device = "png", width = 8, height = 8, units = "in")


#knitr::include_graphics(paste0(IMAGES,Sys.Date(),"_","paye_nuts2",".png"))



ggplotly(paye_stats_barchart_feb20,tooltip = "text") %>% 
  ggla_plotly_settings(legend=FALSE)   %>%
  layout(title = list(text = paste0("<b>","Payrolled employees, change since February 2020 by NUTS2 region","</b>",
                                    "<br>",
                                    "<sup>",
                                    paste0("Latest data for period ", paye_nuts2_stats$date[paye_nuts2_stats$new_date == max(paye_nuts2_stats$new_date)]),
                                    "</sup>",
                                    "<br>"),
                      font = list(size = 22),
                      y = .95, xref = "plot"),
         xaxis = list(tickfont = list(size = 15)),
         yaxis = list(tickfont = list(size = 15)),
         legend = list(font = list(size = 15),title=list(text=""))) %>% 
  plotly_modebar

payebarcapfeb <- paste0("Source: HM Revenue and Customs – Pay As You Earn Real Time Information.","<br>","<br>", "Note: Estimates are based on where employees live.")


```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

### Change on previous year, NUTS2 regions

```{r paye bar, , echo=FALSE,fig.cap = payebarcap, out.width='100%' }


pal <- gla_pal(gla_theme = "default", palette_type = "highlight", n = c(1, 1))
theme_set(theme_gla(gla_theme = "default"))

paye_stats_barchart <- paye_nuts2_stats %>%
  filter(new_date == max(new_date)) %>%
  ungroup() %>% 
  mutate(chart_ranking = case_when( area == "UK" ~ 1,
                                    area == "London" ~ 2,
                                    TRUE ~ dense_rank(desc(annual_change))+2)) %>% 
  ggplot(mapping = aes(x =  reorder(area, -chart_ranking), y = annual_change,
                       colour = area, group = area, fill=area,
         text = paste(
           area, "\n",
           "Change: ", perc_form(annual_change),"%", "\n",
           sep = ""))) +
  geom_text(aes(label=paste0(format(round(annual_change,1), digits = 2), "%"), 
                y = annual_change-0.15*min(annual_change)), 
            color=c("white","white","white","white","white","white","black"))+ #cannot use vjust with ggplotly
  geom_bar(stat="identity",width=0.5)+
  scale_color_manual(values = c(pal[1],pal[1],pal[1],pal[1],pal[1],pal[1],pal[2]), aesthetics = "colour") +
  scale_fill_manual(values = c(pal[1],pal[1],pal[1],pal[1],pal[1],pal[1],pal[2]), aesthetics = "fill") +
  scale_size_manual(values = c(4 * mm_to_pt, 2 * mm_to_pt)) +
  coord_cartesian(clip = 'off') +
  geom_hline(aes(yintercept=0), colour="gray45") +
  scale_y_continuous(labels = dollar_format(prefix = "",
                                            suffix = "%",
                                            largest_with_cents = 1),
                     limits = c(0,10)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  
  theme(plot.margin = unit(c(1,1,1,1), "cm"),axis.text=element_text(size=12))+
  labs(title =  "Payrolled employees, change on previous year by NUTS2 region",
       subtitle = paste0("Latest data for period ", paye_nuts2_stats$date[paye_nuts2_stats$new_date == max(paye_nuts2_stats$new_date)]),
       caption = "\nSource: HM Revenue and Customs – Pay As You Earn Real Time Information.\n\nNote: Estimates are based on where employees live.") + 
  theme(legend.position = "none") +
  theme(plot.caption = element_text(color = rgb(166,166,166,maxColorValue = 255)))

ggsave( paste0(IMAGES,Sys.Date(),"_","paye_nuts2_yoy",".svg"), device = "svg", width = 8, height = 8, units = "in")
ggsave( paste0(IMAGES,Sys.Date(),"_","paye_nuts2_yoy",".png"), device = "png", width = 8, height = 8, units = "in")


#knitr::include_graphics(paste0(IMAGES,Sys.Date(),"_","paye_nuts2_V2",".svg"))


ggplotly(paye_stats_barchart,tooltip = "text") %>% 
  ggla_plotly_settings(legend=FALSE)   %>%
  layout(title = list(text = paste0("<b>","Payrolled employees, change on previous year by NUTS2 region","</b>",
                                    "<br>",
                                    "<sup>",
                                    paste0("Latest data for period ", paye_nuts2_stats$date[paye_nuts2_stats$new_date == max(paye_nuts2_stats$new_date)]),
                                    "</sup>",
                                    "<br>"),
                      font = list(size = 22),
                      y = .95, xref = "plot"),
         xaxis = list(tickfont = list(size = 15)),
         yaxis = list(tickfont = list(size = 15)),
         legend = list(font = list(size = 15),title=list(text=""))) %>% 
  plotly_modebar

payebarcap <- paste0("Source: HM Revenue and Customs – Pay As You Earn Real Time Information.","<br>","<br>", "Note: Estimates are based on where employees live.")


```



<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

## Additional payrolled employee statistics {.tabset}

<br/>

The tabs below are each updated once per quarter, as new data is released by the ONS/HMRC.

<br/>

### Change by local authority

```{r paye la stats,echo=FALSE}

  paye_la_max_val <- perc_form(paye_la_stats %>% filter(new_date == max(new_date)) %>% filter(change_perc_since_feb20 == max(change_perc_since_feb20)) %>% select(change_perc_since_feb20),d = digits_perc)
  
  paye_la_max_area <- paye_la_stats %>% filter(new_date == max(new_date)) %>% filter(change_perc_since_feb20 == max(change_perc_since_feb20)) %>% select(area)
  
  paye_la_min_val <- perc_form(paye_la_stats %>% filter(new_date == max(new_date)) %>% filter(change_perc_since_feb20 == min(change_perc_since_feb20)) %>% select(change_perc_since_feb20),d = digits_perc)
  
  paye_la_min_area <- paye_la_stats %>% filter(new_date == max(new_date)) %>% filter(change_perc_since_feb20 == min(change_perc_since_feb20)) %>% select(area)
  
  paye_la_count_below_feb20 <- paye_la_stats %>% filter(new_date == max(new_date)) %>% mutate(count_below_feb20 = sum(change_perc_since_feb20<0)) %>% filter(row_number()==1) %>%  select(count_below_feb20)
  
  paye_la_count_tot <- paye_la_stats %>% filter(new_date == max(new_date)) %>% add_count() %>% filter(row_number()==1) %>%  select(n)

```

<br/>

There is considerable variation in the change in payrolled employees by London authority. Comparing **`r newest_period_paye_la`** to pre-pandemic (February 2020) levels:

-   The number of payrolled employees living in `r paye_la_max_area` was `r condi_text(paye_la_max_val,"up")` by `r abs2(paye_la_max_val)`%, while the number living in `r paye_la_min_area` was `r condi_text(paye_la_min_val,"up")` by `r abs2(paye_la_min_val)`%.

-   Overall, the number of payrolled employees were still below pre-pandemic levels in `r paye_la_count_below_feb20` out of `r paye_la_count_tot` London authorities (including City of London, see map).

<a href="#top">Back to top</a>

```{r results="asis", echo=FALSE}
# This code is needed to force background of Leaflet map to be white
cat("
<style>
.leaflet-container {
    background: #FFF;
}
</style>
")
```

```{r paye borough map, echo=FALSE,fig.cap = map_cap, out.width='100%' }
  
  # Produces interactive map of boroughs with change in payrolled employees
  
  #Read the shape file from the drive
  boroughs <- readOGR(dsn =  here("INPUT","statistical-gis-boundaries-london","ESRI"),
                      layer = "London_Borough_Excluding_MHW",
                      verbose = FALSE) %>%
    spTransform(CRS("+proj=longlat +datum=WGS84"))
  
  # Prepare data
  paye_la_stats_geo <- paye_la_stats %>%
    mutate(tooltip= paste("Change: ", perc_form(change_perc_since_feb20),"%", "\n", sep = "")) %>% 
    rename(GSS_CODE=gss_code, value=change_perc_since_feb20) %>% #Necessary since the shape files use upper caps variable names
    filter(new_date == max(new_date)) 
  
  #Make a function to make things easier
  
  map_function <- function(data, polygons, bins, palette, map_title){
  
    #Join data to polygons
    polygons@data <- left_join(polygons@data, data, by="GSS_CODE")
    
    #Create the palette
    pal <- colorBin(palette, domain = polygons$value, bins = bins)
    
    #Make tool tip
    labels <- sprintf(
      "<strong>%s</strong><br/>%s",
      polygons$NAME, polygons$tooltip
    ) %>% lapply(htmltools::HTML)
  
    
    #Make map
    map_output <- leaflet() %>%
      addPolygons(data = polygons,
                  fillColor = ~pal(value),
                  weight = 1,
                  opacity = 1,
                  color = "grey",
                  fillOpacity = 0.8,
                  label = labels,
                  labelOptions = labelOptions(
                    style = list("font-weight" = "normal", padding = "3px 8px"),
                    textsize = "15px",
                    direction = "auto")) %>%
      addLegend(pal = pal, values = polygons$value, opacity = 0.8,
                title = map_title,
                position = "bottomright", labFormat = labelFormat(between = "% to ", suffix="%"))
    
    return(map_output)
  }
  
  # NB: THE BINS BELOW MUST BE CHANGED WHEN VALUES CHANGE!
  bins <- c(-8,-6,-4,-2, 0,2,4,6,8)
  
  # Colour palette, title and captions
  map_palette <- gla_pal(palette_type = "diverging", main_colours = c("red","blue"), n = 8, remove_margin = "both")
  map_title <- paste0("Change since","<br>" ,"February 2020 (%)")
  map_cap <- paste0("Source: HM Revenue and Customs – Pay As You Earn Real Time Information. Contains Ordnance Survey data Crown copyright and database rights [2015].","<br>","<br>", "Note: Estimates are based on where employees live.")
  
  #Run Map
  map_function(paye_la_stats_geo, boroughs, bins, map_palette, map_title)

```


<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>


### Change by industry

```{r paye industry stats,echo=FALSE}

# Find industries with lowest and highest growth in recent data

  sector_thresh_paye <- 25000
  
  lowest_growth_ind_paye <- paye_nuts1ind_stats %>% 
    filter(new_date == max(new_date) & geography_name == "London" & emp_level>=sector_thresh_paye & !is.na(change_perc_since_feb20)) %>%
    slice_min(change_perc_since_feb20,n=3) 
  
  highest_growth_ind_paye <- paye_nuts1ind_stats %>% 
    filter(new_date == max(new_date) & geography_name == "London" & emp_level>=sector_thresh_paye & !is.na(change_perc_since_feb20)) %>%
    slice_max(change_perc_since_feb20,n=3) 
  
  ## Find industry names and growth figures
  
  for (measure in c("lowest","highest")) {
    
    dat_temp <- eval(as.name(paste0(measure,"_growth_ind_paye")))
    
    for (num_la in 1:3) {

      growth_x_name <- dat_temp %>% filter(row_number()==num_la) %>% pull(industry_name_simple)
      
      growth_x_p <- perc_form(dat_temp %>% filter(row_number()==num_la) %>% select(change_perc_since_feb20),d = 1)
      
      assign(paste0(measure,"_growth_ind_paye_",num_la,"_name"),growth_x_name)
      assign(paste0(measure,"_growth_ind_paye_",num_la,"_p"),growth_x_p)
      remove(growth_x_name)
      remove(growth_x_p)
    }
    remove(dat_temp)
  }
  
```

<br/>

Some industries have yet to recover from the pandemic. Comparing **`r newest_period_paye_ind`** to pre-pandemic (February 2020) levels:

- `r highest_growth_ind_paye_1_name` saw the highest percentage growth in payrolled employees at `r highest_growth_ind_paye_1_p`%, followed by `r highest_growth_ind_paye_2_name` (`r highest_growth_ind_paye_2_p`%) and `r highest_growth_ind_paye_3_name` (`r highest_growth_ind_paye_3_p`%).

-  On the other hand, `r lowest_growth_ind_paye_1_name` saw the lowest percentage growth with `r condi_text(lowest_growth_ind_paye_1_p,"noun")` of `r abs2(lowest_growth_ind_paye_1_p)`%, followed by `r lowest_growth_ind_paye_2_name` (`r condi_text(lowest_growth_ind_paye_2_p,"up")` by `r abs2(lowest_growth_ind_paye_2_p)`%) and `r lowest_growth_ind_paye_3_name` (`r condi_text(lowest_growth_ind_paye_3_p,"up")` by `r abs2(lowest_growth_ind_paye_3_p)`%).

<a href="#top">Back to top</a>


```{r paye industry region, echo=FALSE, fig.cap=paye_seccap, out.width='100%' }
  
  pal <- gla_pal(gla_theme = "default", palette_type = "highlight", n = c(1, 1))
  theme_set(theme_gla(gla_theme = "default"))
  
  # Select industries large in London, here 25k
  lon_large_inds_paye <- paye_nuts1ind_stats %>%
    filter( emp_level >=sector_thresh_paye 
            & geography_name =="London" & new_date == max(new_date)) %>% 
    select(industry_name_simple)
  
  lon_large_inds_paye <- unlist(lon_large_inds_paye)
  
  paye_industry_bar <- paye_nuts1ind_stats %>%
    filter(new_date == max(new_date) & industry_name_simple %in% lon_large_inds_paye) %>% 
    group_by(geography_name) %>% #To ensure chart ranking is correct, make rank based on London and assign within sector
    mutate(chart_ranking = case_when(
      geography_name=="London" ~ dense_rank(desc(change_perc_since_feb20)))) %>% 
    group_by(industry_name_simple) %>% 
    mutate(chart_ranking=min(chart_ranking,na.rm=TRUE)) %>% 
    ungroup() %>% 
    arrange(chart_ranking) %>% 
    ggplot(mapping = aes(x =  factor(reorder(industry_name_simple,-chart_ranking)), 
                         y = change_perc_since_feb20, 
                         colour = factor(reorder(geography_name,desc(geography_name))) ,#since horizontal bar reverses orders, we need to reverse too
                         fill=factor(reorder(geography_name,desc(geography_name))),
                         width = 0.5,
                         text = paste(industry_name_simple, "\n",
                                      geography_name, "\n",
                                      "Change: ", perc_form(change_perc_since_feb20),"%", "\n",
                                      sep = ""))) +
    geom_bar(stat = "identity", position = position_dodge()) +
    geom_hline(aes(yintercept=0), colour="gray45") +
    scale_color_manual(values = rev(pal), aesthetics = "colour")+
    scale_fill_manual(values = rev(pal), aesthetics = "fill")+
    theme_set(theme_gla(gla_theme = "default", y_label_length=100)) + #GLA theme and removes lines below y-axis labels
    scale_y_continuous(limits = c(-14, 12), labels = dollar_format(prefix = "", 
                                                                   suffix = "%", 
                                                                   largest_with_cents = 1)) +
    theme(plot.margin = unit(c(1,1,1,1), "cm"))+
    scale_x_discrete(expand = c(0,0)) +
    theme(panel.grid.major.y = element_blank(), #removed y grid
          panel.grid.minor.y = element_blank(), # removes small y grid
          panel.grid.major.x = element_line( size=.5 ), # removes x grid
          axis.text.y = ggplot2::element_text( # ensures  y axis labels stay outside chart
            hjust = 0, vjust = 0.5,
            margin = ggplot2::margin(t = 0, r = 0, b = 0, l = 0,
                                     unit = "pt")),
          axis.ticks.length.y = ggplot2::unit(x = 0, units = "pt")) + #removes ticks from y axis
    coord_flip()+
    guides(colour=guide_legend(reverse=TRUE),
           fill=guide_legend(reverse=TRUE)) +
    labs(title = paste0("Payrolled employees, change since February 2020 by selected sector"),
         subtitle = paste0("Latest data for period ", newest_period_paye_ind),
         caption = "\nSource: HM Revenues and Customs – Pay As You Earn Real Time Information.\n\nNote: Estimates are based on where employees live. Chart excludes sectors with fewer than 25,000 payrolled employees in London.")+
    theme(plot.caption = element_text(color = rgb(166,166,166,maxColorValue = 255)))
  
  ggsave(paste0(IMAGES,Sys.Date(),"_","paye_industries",".svg"), device = "svg", width = 8, height = 8, units = "in")
  ggsave(paste0(IMAGES,Sys.Date(),"_","paye_industries",".png"), device = "png", width = 8, height = 8, units = "in")
  
  
  #knitr::include_graphics(paste0(IMAGES,Sys.Date(),"_","wfj_industries",".svg"))
  
  ggplotly(paye_industry_bar,tooltip = "text") %>% 
    ggla_plotly_settings()    %>% 
    layout(title = list(text = paste0("<b>","Payrolled employees, change since February 2020 by selected sector","</b>",
                                      "<br>",
                                      "<sup>",
                                      paste0("Latest data for period ", newest_period_paye_ind),
                                      "</sup>",
                                      "<br>"),
                        font = list(size = 22),
                        y = .95, xref = "plot"),
           xaxis = list(tickfont = list(size = 15), showgrid=TRUE),
           yaxis = list(tickfont = list(size = 15), showgrid=FALSE),
           legend = list(font = list(size = 15),title=list(text="")),
           margin = list(l = 160)) %>% 
    reverse_legend_labels()%>% 
    plotly_modebar
  
  
  paye_seccap <- paste0("Source: HM Revenue and Customs – Pay As You Earn Real Time Information.","<br>","<br>", "Note: Estimates are based on where employees live. Excludes sectors with fewer than 25,000 payrolled employees in London.")

```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>



### Change by age group


```{r paye age stats,echo=FALSE}
  
  #Max change in London
  paye_age_max_perc_lon <- perc_form(paye_nuts1age_stats %>% ungroup() %>% filter(new_date == max(new_date) & area == "London" & !(age_group %in% c("Aged: 0-17"))) %>% filter(change_perc_since_feb20 == max(change_perc_since_feb20) ) %>% select(change_perc_since_feb20),d = digits_perc)
  
  paye_age_max_age_lon <- as.character(paye_nuts1age_stats %>% ungroup() %>% filter(new_date == max(new_date) & area == "London" & !(age_group %in% c("Aged: 0-17"))) %>% filter(change_perc_since_feb20 == max(change_perc_since_feb20) ) %>% select(age_group))
  
  paye_age_max_l_perc_uk <- perc_form(paye_nuts1age_stats %>% ungroup() %>% filter(new_date == max(new_date) & area == "United Kingdom" & !(age_group %in% c("Aged: 0-17"))) %>% filter(age_group == paye_age_max_age_lon ) %>% select(change_perc_since_feb20),d = digits_perc)
  
  #Min change in London
  paye_age_min_perc_lon <- perc_form(paye_nuts1age_stats %>% ungroup() %>% filter(new_date == max(new_date) & area == "London" & !(age_group %in% c("Aged: 0-17"))) %>% filter(change_perc_since_feb20 == min(change_perc_since_feb20) ) %>% select(change_perc_since_feb20),d = digits_perc)
  
  paye_age_min_age_lon <- as.character(paye_nuts1age_stats %>% ungroup() %>% filter(new_date == max(new_date) & area == "London" & !(age_group %in% c("Aged: 0-17"))) %>% filter(change_perc_since_feb20 == min(change_perc_since_feb20) ) %>% select(age_group))
  
  paye_age_min_l_perc_uk <- perc_form(paye_nuts1age_stats %>% ungroup() %>% filter(new_date == max(new_date) & area == "United Kingdom" & !(age_group %in% c("Aged: 0-17"))) %>% filter(age_group == paye_age_min_age_lon ) %>% select(change_perc_since_feb20),d = digits_perc)

```

<br/>

Recent changes in the number of payrolled employees vary by age group. Comparing **`r newest_period_paye_age`** to pre-pandemic (February 2020) levels:

- The number of payrolled employees "`r paye_age_max_age_lon`" in London was `r condi_text(paye_age_max_perc_lon,"up")` by `r abs2(paye_age_max_perc_lon)`%, while the number of employees "`r paye_age_min_age_lon`" was `r condi_text(paye_age_min_perc_lon,"up")` by `r abs2(paye_age_min_perc_lon)`%.

- For the UK overall, the number of payrolled employees in those two age groups saw `r condi_text(paye_age_max_l_perc_uk,"noun")` of `r abs2(paye_age_max_l_perc_uk)`% and no change, respectively.

<a href="#top">Back to top</a>

```{r paye age region, echo=FALSE, fig.cap=paye_agecap, out.width='100%' }
  
  ### Change in PAYE employees by age group since Feb2020
  pal <- gla_pal(gla_theme = "default", palette_type = "highlight", n = c(1, 1))
  theme_set(theme_gla(gla_theme = "default"))
  
  
  paye_age_bars <- paye_nuts1age_stats %>%
    group_by(area) %>% 
    filter( (new_date == max(new_date)) & 
              !(age_group %in% c("Aged: 0-17"))
            & area %in% c("London", "United Kingdom")) %>%  
    mutate(chart_ranking = rank((age_group))) %>% 
    ungroup() %>% 
    ggplot(mapping = aes(x = reorder(age_group,chart_ranking), 
                         y = change_perc_since_feb20, 
                         colour = area , fill=area,
                         text = paste(age_group, "\n",
                                      area, "\n",
                                      "Change: ", perc_form(change_perc_since_feb20),"%", "\n",
                                      sep = ""))) +
    geom_bar(stat = "identity", position = position_dodge(), width = 0.4)+
    geom_hline(aes(yintercept=0), colour="gray45") +
    scale_colour_manual(values = pal) +
    scale_fill_manual(values = pal) +
    theme_set(theme_gla(gla_theme = "default")) +
    scale_y_continuous(limits = c(-4, 6), labels = dollar_format(prefix = "", 
                                                                 suffix = "%", 
                                                                 largest_with_cents = 1)) +
    theme(axis.text.x = element_text( hjust=1, vjust=0.5)) +
    theme(plot.margin = unit(c(1,1,1,1), "cm"))+
    labs(title = "Payrolled employees, change since February 2020 by age group",
         subtitle = paste0("Latest data for period ", newest_period_paye_age),
         caption = "\nSource: HM Revenue and Customs – Pay As You Earn Real Time Information.\n\nNote: Estimates are based on where employees live.") +
    theme(plot.caption = element_text(color = rgb(166,166,166,maxColorValue = 255)))
  
  ggsave( paste0(IMAGES,Sys.Date(),"_","paye_age_feb20.svg"), device = "svg", width = 8, height = 8, units = "in")
  ggsave( paste0(IMAGES,Sys.Date(),"_","paye_age_feb20.png"), device = "png", width = 8, height = 8, units = "in")
  
  
  ggplotly(paye_age_bars,tooltip = "text") %>% 
    ggla_plotly_settings()    %>% 
    layout(title = list(text = paste0("<b>","Payrolled employees, change since February 2020 by age group","</b>",
                                      "<br>",
                                      "<sup>",
                                      paste0("Latest data for period ", newest_period_paye_age),
                                      "</sup>",
                                      "<br>"),
                        font = list(size = 22),
                        y = .95, xref = "plot"),
           xaxis = list(tickfont = list(size = 15)),
           yaxis = list(tickfont = list(size = 15)),
           legend = list(font = list(size = 15),title=list(text=""))) %>% 
    plotly_modebar
  
  
  paye_agecap <- paste0("Source: HM Revenue and Customs – Pay As You Earn Real Time Information.","<br>","<br>", "Note: Estimates are based on where employees live.")

  
```



<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>


``` {r paye_nat_figures, echo=FALSE}

  last_nat_paye_date <- format(max(paye_nat_long$date_day),"%B %Y")
  
  paye_EUnat_london_current <- value_form(pull_dtpt(dt=paye_nat_long,flt="nationality == 'EU' & region== 'London' & date_day == max(date_day)",vr="employee_count"),s=signif_thous)
  
  paye_ROWnat_london_current <- value_form(pull_dtpt(dt=paye_nat_long,flt="nationality == 'ROW' & region== 'London' & date_day == max(date_day)",vr="employee_count"),s=signif_thous)
  
  paye_EUnat_london_feb20 <- perc_form(pull_dtpt(dt=paye_nat_long,flt="nationality == 'EU' & region== 'London' & date_day == max(date_day)",vr="p_change_feb20"))
  
  paye_ROWnat_london_feb20 <- perc_form(pull_dtpt(dt=paye_nat_long,flt="nationality == 'ROW' & region== 'London' & date_day == max(date_day)",vr="p_change_feb20"))
  
  paye_EUnat_uk_feb20 <- perc_form(pull_dtpt(dt=paye_nat_long,flt="nationality == 'EU' & region== 'United Kingdom' & date_day == max(date_day)",vr="p_change_feb20"))
  
  paye_ROWnat_uk_feb20 <- perc_form(pull_dtpt(dt=paye_nat_long,flt="nationality == 'ROW' & region== 'United Kingdom' & date_day == max(date_day)",vr="p_change_feb20"))

```


## Payrolled employees by nationality {.tabset}
<br/>

The ONS has released data on the number of payrolled employees by nationality, categorised either as UK, EU or Rest of the World (ROW) nationals. Note this data only covers employee numbers up until **`r last_nat_paye_date`**. At that time, the number of payrolled employees with EU nationality living in London was `r paye_EUnat_london_current`, while the number of ROW nationals was `r paye_ROWnat_london_current`.

From February 2020 to `r last_nat_paye_date`:

- The change in total payrolled employees in London was significantly impacted by changes in the number of EU and ROW nationals. The number of payrolled EU nationals `r condi_text(paye_EUnat_london_feb20, t= "increase")` by `r abs2(paye_EUnat_london_feb20)`%, while the number of ROW nationals `r condi_text(paye_ROWnat_london_feb20, t= "increase")` by `r abs2(paye_ROWnat_london_feb20)`%.

- In the UK overall, the same groups saw `r condi_text(paye_EUnat_uk_feb20, t= "noun")`  of `r abs2(paye_EUnat_uk_feb20)`% and `r condi_text(paye_ROWnat_uk_feb20, t= "noun")` of `r abs2(paye_ROWnat_uk_feb20)`%, respectively. 
 
 <a href="#top">Back to top</a>

<br/>

### Change since February 2020

``` {r paye_nat_bar, echo=FALSE, fig.cap=paye_natbar, out.width='100%'}
 # Change by nationality in London and UK bar chart
  nats_sorted <- c("Total","EU","ROW","UK")
  pal <- gla_pal(gla_theme = "default", palette_type = "highlight", n = c(1, 1))
  pal_named <- setNames(object=pal,nm = c("London","United Kingdom"))
  theme_set(theme_gla(gla_theme = "default"))
  
  paye_nat_bars <- paye_nat_long %>%
    filter(region %in% c("London","United Kingdom") & date_day==max(date_day)) %>% 
    ggplot(mapping = aes(x = factor(nationality,levels = nats_sorted), 
                         y = p_change_feb20, 
                         colour = region , fill=region,
                         text = paste("Nationality: ", nationality, "\n",
                                      region, "\n",
                                      "Change: ", perc_form(p_change_feb20),"%", "\n",
                                      sep = ""))) +
    geom_bar(stat = "identity", position = position_dodge(), width = 0.4)+
    ggla_axisat0() +
    scale_colour_manual(values = pal_named) +
    scale_fill_manual(values = pal_named) +
    theme_set(theme_gla(gla_theme = "default")) +
    scale_y_continuous( labels = dollar_format(prefix = "", 
                                               suffix = "%", 
                                               largest_with_cents = 1)) +
    theme(axis.text.x = element_text( hjust=1, vjust=0.5)) +
    theme(plot.margin = unit(c(1,1,1,1), "cm"))+
    labs(title = paste0("Payrolled employees, change from February 2020 to ",last_nat_paye_date),
         subtitle = paste0("By nationality, latest data for period ", last_nat_paye_date),
         caption = "\nSource: HM Revenue and Customs – Pay As You Earn Real Time Information.\n\nNote: Estimates are based on where employees live.") +
    theme(plot.caption = element_text(color = rgb(166,166,166,maxColorValue = 255)))
  
  ggsave( paste0(IMAGES,Sys.Date(),"_","paye_nat_feb20_bar.svg"), device = "svg", width = 8, height = 8, units = "in")
  ggsave( paste0(IMAGES,Sys.Date(),"_","paye_nat_feb20_bar.png"), device = "png", width = 8, height = 8, units = "in")
  
  ggplotly(paye_nat_bars,tooltip = "text") %>% 
    ggla_plotly_settings()    %>% 
    layout(title = list(text = paste0("<b>","Payrolled employees, change from February 2020 to ",last_nat_paye_date,"</b>",
                                      "<br>",
                                      "<sup>",
                                      paste0("By nationality, latest data for period ", last_nat_paye_date),
                                      "</sup>",
                                      "<br>"),
                        font = list(size = 22),
                        y = .95, xref = "plot"),
           xaxis = list(tickfont = list(size = 15)),
           yaxis = list(tickfont = list(size = 15)),
           legend = list(font = list(size = 15),title=list(text=""))) %>% 
    plotly_modebar
  

paye_natbar <- paste0("Source: HM Revenue and Customs – Pay As You Earn Real Time Information and Migrant Worker Scan (MWS).","<br>","<br>", "Note: Estimates are based on where employees live. The nationality of a worker is registered when applying for a National Insurance number and may therefore not reflect subsequent changes of nationality.")

```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

### Trends in London

``` {r paye_nat_index, echo=FALSE, fig.cap=paye_natindex, out.width='100%'}

 ## Charts
  # - Index trend chart for London
  pal <- c(gla_colours$blue_light,gla_colours$turquoise_core,"#cccccc") # blue, turquoise and grey
  pal_named <- setNames(object=pal,nm = c("EU","ROW","UK"))
  
  scale_size <- setNames(object = rep(2* mm_to_pt,3),nm =  c("EU","ROW","UK"))
    
  paye_nat_london_trend <- paye_nat_long %>% 
    filter(region=="London" & nationality %in% c("EU","ROW","UK") & date_day>="2020-02-01") %>% 
    ggplot(mapping = aes(x = date_day, y = index_feb20, 
                           group = nationality, 
                           text = paste(
                             "Nationality: ", nationality, "\n",
                             date, "\n",
                             "Index: ", value_form(index_feb20,d=1,s=3), "\n",
                             sep = ""))) +
      ggla_line(aes(colour = nationality, size = nationality)) +
      scale_colour_manual(values = pal_named) +
      scale_size_manual(values = scale_size) +
      geom_hline(aes(yintercept=0), colour="gray45") +
      coord_cartesian(clip = 'off') +
      scale_y_continuous(limits = c(85, 105),labels = dollar_format(prefix = "", 
                                                                  largest_with_cents = 1,
                                                                  suffix = "")) +
      scale_x_date(date_labels = "%b %Y") +
      theme(plot.margin = unit(c(1,1,1,1), "cm"))+
      labs(title = paste0("Payrolled employees, trends in London"),
           subtitle = paste0("By nationality, indexed to Feb 2020"),
           caption = "\nSource: HM Revenue and Customs – Pay As You Earn Real Time Information.\n\nNote: Estimates are based on where employees live.") +
      theme(plot.caption = element_text(color = rgb(166,166,166,maxColorValue = 255)))
  
  ggsave( paste0(IMAGES,Sys.Date(),"_","paye_nat_feb20_index.svg"), device = "svg", width = 8, height = 8, units = "in")
  ggsave( paste0(IMAGES,Sys.Date(),"_","paye_nat_feb20_index.png"), device = "png", width = 8, height = 8, units = "in")
  
  ggplotly(paye_nat_london_trend,tooltip = "text") %>% 
    ggla_plotly_settings()    %>% 
    layout(title = list(text = paste0("<b>","Payrolled employees, trends in London","</b>",
                                      "<br>",
                                      "<sup>",
                                      paste0("By nationality, indexed to Feb 2020, latest data for period ", last_nat_paye_date),
                                      "</sup>",
                                      "<br>"),
                        font = list(size = 22),
                        y = .95, xref = "plot"),
           xaxis = list(tickfont = list(size = 15)),
           yaxis = list(tickfont = list(size = 15)),
           legend = list(font = list(size = 15),title=list(text="")),
           hovermode = "x") %>% 
    plotly_modebar
  
  
 
  

paye_natindex <- paste0("Source: HM Revenue and Customs – Pay As You Earn Real Time Information and Migrant Worker Scan (MWS).","<br>","<br>", "Note: Estimates are based on where employees live.  The nationality of a worker is registered when applying for a National Insurance number and may therefore not reflect subsequent changes of nationality.")

```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

```{r, emp_figures, echo=FALSE}
  
  emp_rate_uk <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom" & value_type_name == "Level" & measures == 20207)  %>% select(total_in_employment_aged_16_to_64),d = digits_perc)
  
  emp_pp_quart_uk <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom"  & value_type_name == "Change on quarter" & measures == 20207)  %>% select(total_in_employment_aged_16_to_64),d = digits_perc)
  
  emp_pp_year_uk <-perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom"  & value_type_name == "Change on year" & measures == 20207) %>% select(total_in_employment_aged_16_to_64),d = digits_perc)

```

## Employment rate

<br/>

Employment measures the number of people aged 16 and over in paid work and those who had a job that they were temporarily away from. The employment rate is the proportion of people aged between 16 and 64 years who are in employment.

For `r lfs_lastmonth` to `r lfs_newmonth`:

-   The 16-64 employment rate in London was estimated at `r emp_rate`%; this represents `r condi_text(emp_pp_quart,"noun")` of `r abs2(emp_pp_quart)` percentage points (pp) on the previous quarter and `r condi_text(emp_pp_year,"noun")` of `r abs2(emp_pp_year)`pp from a year earlier.

-   The overall UK employment rate was estimated at `r emp_rate_uk`%, `r condi_text(emp_pp_quart_uk,"noun")` of `r abs2(emp_pp_quart_uk)`pp on the quarter and `r condi_text(emp_pp_year_uk,"noun")` of `r abs2(emp_pp_year_uk)`pp on the year.

<a href="#top">Back to top</a>

```{r employ chart, echo=FALSE,fig.cap = empcap, out.width='100%'}
  ### The below should be the relevant chart
  employment_rate <- GLALineChart(data_set = lfs_stats, y_var = total_in_employment_aged_16_to_64, 
                           title = "Employment rate (% of working age population)",
                           caption = "\nSource: ONS Labour Force Survey.\n\nNote: The margin of error is +/- 1.3% for London and +/- 0.4% for the UK.",
                           chart_name = "employment", y_limits = c(70,80), nudge_y = c(-0.5, +0.5))
  
  
  #knitr::include_graphics(paste0(IMAGES,Sys.Date(),"_","employment",".svg"))
  
  ggplotly(employment_rate,tooltip = "text") %>% 
    ggla_plotly_settings()    %>% 
    layout(title = list(text = paste0("<b>","Employment rate (% of working age population)","</b>",
                                      "<br>"),
           font = list(size = 22),
           y = .95, xref = "plot"),
           xaxis = list(tickfont = list(size = 15)),
           yaxis = list(tickfont = list(size = 15)),
           legend = list(font = list(size = 15),title=list(text="")),
           hovermode = "x") %>% 
    plotly_modebar
  
  empcap <- paste0("Source: ONS Labour Force Survey.","<br>","<br>", "Note: The margin of error is +/- 1.4% for London and +/- 0.4% for the UK.")


```

```{r, unem_figures, echo=FALSE}

  unem_pp_quart_uk <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom" & value_type_name == "Change on quarter" & measures == 20207) %>% select(total_unemployed_aged_16_and_over),d = digits_perc)
  
  
  unem_pp_year_uk <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom" & value_type_name == "Change on year" & measures == 20207)  %>% select(total_unemployed_aged_16_and_over),d = digits_perc)

```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

## Unemployment rate

<br/>

Unemployment measures people without a job who have been actively seeking work within the last four weeks and are available to start work within the next two weeks. The unemployment rate is the proportion of the economically active population who are unemployed.

For `r lfs_lastmonth` to `r lfs_newmonth`:

-   The unemployment rate for London was estimated at `r unem_rate`%; this represents `r condi_text(unem_pp_quart,"noun")` of `r abs2(unem_pp_quart)`pp on the previous quarter and `r condi_text(unem_pp_year,"noun")` of `r abs2(unem_pp_year)`pp from a year earlier.

-   The UK unemployment rate was estimated at `r unem_rate_uk`%, `r condi_text(unem_pp_quart_uk,"noun")` of `r abs2(unem_pp_quart_uk)`pp from the previous quarter and `r condi_text(unem_pp_year_uk,"noun")` of `r abs2(unem_pp_year_uk)`pp on the year.

<a href="#top">Back to top</a>

```{r unemployed chart, echo=FALSE,fig.cap = unempcap, out.width='100%'}

  unemployment_rate <- GLALineChart( data_set = lfs_stats, y_var = total_unemployed_aged_16_and_over, 
                               title = "Unemployment rate (% of economically active)",
                               caption = "\nSource: ONS Labour Force Survey.\n\nNote: The margin of error is +/- 0.8% for London and +/- 0.2% for the UK.",
                               chart_name = "unemployment", y_limits = c(0,8), nudge_y = 0.5)
  
  #knitr::include_graphics(paste0(IMAGES,Sys.Date(),"_","unemployment",".svg"))
  
  ggplotly(unemployment_rate,tooltip = "text") %>% 
    ggla_plotly_settings()    %>% 
    layout(title = list(text = paste0("<b>","Unemployment rate (% of economically active)","</b>",
                                      "<br>"),
           font = list(size = 22),
           y = .95, xref = "plot"),
           xaxis = list(tickfont = list(size = 15)),
           yaxis = list(tickfont = list(size = 15)),
           legend = list(font = list(size = 15),title=list(text="")),
           hovermode = "x") %>% 
    plotly_modebar
  
  unempcap <- paste0("Source: ONS Labour Force Survey.","<br>","<br>", "Note: The margin of error is +/- 0.8% for London and +/- 0.2% for the UK.")
  


```

```{r inact_figures,echo=FALSE}
  
  inact_rate <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London" & value_type_name == "Level" & measures == 20207)  %>% select(total_economically_inactive_aged_16_to_64),d= digits_perc)
    
  inact_pp_quart <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London" & value_type_name == "Change on quarter" & measures == 20207) %>% select(total_economically_inactive_aged_16_to_64),d= digits_perc)
    
  inact_pp_year <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "London" & value_type_name == "Change on year" & measures == 20207)  %>% select(total_economically_inactive_aged_16_to_64),d= digits_perc)
  
  inact_rate_uk <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom" & value_type_name == "Level" & measures == 20207)  %>% select(total_economically_inactive_aged_16_to_64),d= digits_perc)
    
  inact_pp_quart_uk <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom" & value_type_name == "Change on quarter" & measures == 20207) %>% select(total_economically_inactive_aged_16_to_64),d= digits_perc)
    
  inact_pp_year_uk <- perc_form(lfs_stats %>% filter(sex_name == "Total" & date == max(date) & geography_name == "United Kingdom" & value_type_name == "Change on year" & measures == 20207)  %>% select(total_economically_inactive_aged_16_to_64),d= digits_perc)

```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

## Economic inactivity rate

<br/>

The economic inactivity rate is the proportion of 16 to 64 year olds not in work and either not looking for or unable to work. This group includes some students, people who are looking after family/home, and people who are too ill to work (most of whom are long-term sick).

For `r lfs_lastmonth` to `r lfs_newmonth`:

-   The rate of economic inactivity in London was estimated at `r inact_rate`%; this represents `r condi_text(inact_pp_quart,"noun")` of `r abs2(inact_pp_quart)`pp on the previous quarter and `r condi_text(inact_pp_year,"noun")` of `r abs2(inact_pp_year)`pp from a year earlier.

-   The UK rate of economic inactivity was `r inact_rate_uk`%. This represents `r condi_text(inact_pp_quart_uk,"noun")` of `r abs2(inact_pp_quart_uk)`pp on the previous quarter but no change on the year.

<a href="#top">Back to top</a>

```{r inactive chart, echo=FALSE,fig.cap = inactcap, out.width='100%'}

  econ_inactive <- GLALineChart( data_set = lfs_stats, y_var = total_economically_inactive_aged_16_to_64, 
                              title =  "Economic inactivity (% of working age population)",
                               caption = "\nSource: ONS Labour Force Survey.\n\nNote: The margin of error is not published for London, the UK margin is +/- 0.4%.",
                              chart_name = "inact", y_limits = c(15,25), nudge_y = c(-0.5, +0.5))
  
  #knitr::include_graphics(paste0(IMAGES,Sys.Date(),"_","inact",".svg"))
  
  
  ggplotly(econ_inactive,tooltip = "text") %>% 
    ggla_plotly_settings()    %>% 
    layout(title = list(text = paste0("<b>","Economic inactivity (% of working age population)","</b>",
                                      "<br>"),
           font = list(size = 22),
           y = .95, xref = "plot"),
           xaxis = list(tickfont = list(size = 15)),
           yaxis = list(tickfont = list(size = 15)),
           legend = list(font = list(size = 15),title=list(text="")),
           hovermode = "x") %>% 
    plotly_modebar
  
  inactcap <- paste0("Source: ONS Labour Force Survey.","<br>","<br>", "Note: The margin of error is not published for London, the UK margin is +/- 0.4%.")

```

```{r claim_figures, echo=FALSE}

  claim_period <- claimant_count_stats_long %>%  filter(date == max(date)) %>% filter(row_number()==1) %>% mutate(date = format(date, '%B %Y'))  %>% select(date)
  
  claim_level <- value_form(claimant_count_stats_long %>% filter( sex_name == "Total" & geography_name == "London" & date == max(date) & age_name == "All categories: Age 16+" & measure_name == "claimant_count") %>% select(measure_value),s = signif_thous)
    
  claim_rate <- perc_form(claimant_count_stats_long %>% filter( sex_name == "Total" & geography_name == "London" & date == max(date) & age_name == "All categories: Age 16+" & measure_name == "claimants_as_a_proportion_of_residents_aged_16_64") %>% select(measure_value),d = digits_perc)
  
  claim_d_year <- value_form(claimant_count_stats_long %>% filter( sex_name == "Total" & geography_name == "London" & date == max(date) & age_name == "All categories: Age 16+" & measure_name == "claimant_count") %>% select(change_year),s = signif_thous)
    
  claim_p_year <- perc_form(claimant_count_stats_long %>% filter( sex_name == "Total" & geography_name == "London" & date == max(date) & age_name == "All categories: Age 16+" & measure_name == "claimant_count") %>% select(change_year_percent),d = digits_perc)
    
  claim_d_year_uk <- value_form(claimant_count_stats_long %>% filter( sex_name == "Total" & geography_name == "United Kingdom" & date == max(date) & age_name == "All categories: Age 16+" & measure_name == "claimant_count") %>% select(change_year),s = signif_thous)
    
  claim_p_year_uk <- perc_form(claimant_count_stats_long %>% filter( sex_name == "Total" & geography_name == "United Kingdom" & date == max(date) & age_name == "All categories: Age 16+" & measure_name == "claimant_count") %>% select(change_year_percent),d = digits_perc)

```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

## Claimant count {.tabset}

<br/>

The claimant count is a measure of the number of people claiming benefits principally for the reason of being unemployed. It includes people claiming Jobseeker's Allowance and those claiming Universal Credit who are required to seek work.

The latest data for `r claim_period` show that:

-   There were `r claim_level` people claiming unemployment-related benefits in London - equal to a claimant unemployment rate (as a proportion of residents aged 16 to 64) of `r claim_rate`%.

-   The number of claimants in London has `r condi_text(claim_d_year,"increase")` by `r abs2(claim_d_year)`, or `r abs2(claim_p_year)`%, on the year, compared to `r condi_text(claim_d_year,"noun")` of `r abs2(claim_d_year_uk)`, or `r abs2(claim_p_year_uk)`%, in the UK overall.

<a href="#top">Back to top</a>

### Rate of residents aged 16-64

```{r claimant chart, echo=FALSE,fig.cap = claimcap, out.width='100%'}
  
  ### Create claimant count charts
  
  claimant_count_percent <- GLALineChart( data_set = claimant_count_stats_wide, y_var = claimants_as_a_proportion_of_residents_aged_16_64, 
                                          suffix = "%", lfs_or_claims = "claims", title =  "Claimant count (as a % of residents aged 16 to 64)",
                                          caption = "\nSource: ONS Claimant Count by sex and age (NSA).\n\nNote: may include some employed claimants on low hours or earnings.",
                                          chart_name = "claimant_count_per", y_limits = c(0,10), nudge_y = 0.5)
  
  #knitr::include_graphics(paste0(IMAGES,Sys.Date(),"_","claimant_count_per",".svg"))
  
  ggplotly(claimant_count_percent,tooltip = "text") %>% 
    ggla_plotly_settings()    %>% 
    layout(title = list(text = paste0("<b>","Claimant count (as a % of residents aged 16 to 64)","</b>",
                                      "<br>"),
                        font = list(size = 22),
                        y = .95, xref = "plot"),
           xaxis = list(tickfont = list(size = 15)),
           yaxis = list(tickfont = list(size = 15)),
           legend = list(font = list(size = 15),title=list(text="")),
           hovermode = "x") %>% 
    plotly_modebar
  
  claimcap <- paste0("Source: ONS Claimant Count by sex and age (NSA).","<br>","<br>", "Note: may include some employed claimants on low hours or earnings.")
  
  ### Manually plot total London claimants
  
  new_release_text <- claimant_count_stats_wide$date_name[claimant_count_stats_wide$date == max(claimant_count_stats_wide$date)]
  
  
  claimant_count_total <- claimant_count_stats_wide %>%
    filter( geography_name == "London" & sex_name == "Total" & !is.na(claimant_count) 
            & age_name == "All categories: Age 16+") %>% 
    ggplot(mapping = aes(x = date, y = claimant_count, 
                         colour = geography_name, group = geography_name)) +
    ggla_line(aes(size= geography_name)) +
    scale_size_manual(values = c(4 * mm_to_pt, 2 * mm_to_pt)) +
    scale_colour_manual(values = pal) +
    ggla_highlight(filter_type = "end") +
    ggla_highlight(mapping = aes(label = format(claimant_count, big.mark = ",")),
                   geom = GeomGLATextHighlight, filter_type = "end",  size = 4.5,
                   position = position_nudge(y = 20000),check_overlap = TRUE)+
    coord_cartesian(clip = 'off') +
    ggla_axisat0() +
    scale_y_continuous(expand = c(0, 0), labels = dollar_format(prefix = "", 
                                                                suffix = "", 
                                                                largest_with_cents = 1), 
                       limits = c(0,750000)) +
    scale_x_date( date_breaks = "2 months",
                  date_labels = "%b'%y",
                  expand = expansion( mult = c(0.1,0.02))) +
    theme(plot.margin = unit(c(1,1,1,1), "cm"))+
    labs(title =  "Claimant count in London",
         subtitle = paste0("Latest data for period ", new_release_text),
         caption = "\nSource: ONS Claimant Count by sex and age (NSA).\n\nNote: May include some employed claimants on low hours or earnings.")+
    theme(plot.caption = element_text(color = rgb(166,166,166,maxColorValue = 255)))
  
  ggsave( paste0(IMAGES,Sys.Date(),"_","claimant_count",".svg"), device = "svg", width = 8, height = 8, units = "in")
  ggsave( paste0(IMAGES,Sys.Date(),"_","claimant_count",".png"), device = "png", width = 8, height = 8, units = "in") 
  
  # knitr::include_graphics(paste0(IMAGES,Sys.Date(),"_","claimant_count",".svg"))
  
  ### Claimant count bar charts (edit variable names)
  
  new_release_cc <- max(claimant_count_stats_wide$date)
  year_ago_cc <- ymd(new_release_cc) - years(1)


```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

### Year-on-year percentage change

```{r claimant increase, echo = FALSE,fig.cap = claimbarcap, out.width='100%'}
  
  ### Increases in claimants by age
  pal <- gla_pal(gla_theme = "default", palette_type = "highlight", n = c(1, 1))
  theme_set(theme_gla(gla_theme = "default"))
  
  claimant_count_bar <- claimant_count_stats_long %>%
    filter(measure_name=="claimant_count") %>% 
    group_by(sex_name, geography_name, date) %>% 
    mutate(chart_ranking = case_when( age_name == "All categories: Age 16+" ~ 1,
                                      TRUE ~ rank((age_name))+1)) %>% 
    ungroup() %>% 
    filter( sex_name == "Total" & (date == max(date)) & !(age_name %in% c("Aged 16-17", "Aged 18-21", "Aged 18-24", "Aged 50+" , "Age unknown (clerical claims)", "Aged 25-49", "Aged 65+"))) %>% 
    ggplot(mapping = aes(x = reorder(age_name,chart_ranking), 
                         y = change_year_percent, 
                         colour = geography_name , fill=geography_name,
                         text = paste(age_name, "\n",
                                      geography_name, "\n",
                                      "Change in count: ", perc_form(change_year_percent),"%", "\n",
                                      sep = ""))) +
    geom_bar(stat = "identity", position = position_dodge(), width = 0.4) +
    geom_hline(aes(yintercept=0), colour="gray45") +
    scale_colour_manual(values = pal) +
    scale_fill_manual(values = pal) +
    theme_set(theme_gla(gla_theme = "default")) +
    scale_y_continuous(limits = c(-50, 10), labels = dollar_format(prefix = "", 
                                                                   suffix = "%", 
                                                                   largest_with_cents = 1)) +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 8)) +
    theme(axis.text.x = element_text( hjust=1, vjust=0.5)) +
    theme(plot.margin = unit(c(1,1,1,1), "cm"))+
    labs(title = "Year-on-year percentage change in claimant count",
         subtitle = paste0("Latest data for period ", format(new_release_cc,"%B %Y")),
         caption = "\nSource: ONS Claimant Count by sex and age (NSA).\n\nNote: May include some employed claimants on low hours or earnings.") +
    theme(plot.caption = element_text(color = rgb(166,166,166,maxColorValue = 255)))
  
  ggsave( paste0(IMAGES,Sys.Date(),"_","claimant_age_increase.svg"), device = "svg", width = 8, height = 8, units = "in")
  ggsave( paste0(IMAGES,Sys.Date(),"_","claimant_age_increase.png"), device = "png", width = 8, height = 8, units = "in")
  
  #knitr::include_graphics(paste0(IMAGES,Sys.Date(),"_","claimant_age_increase",".svg"))
  
  ggplotly(claimant_count_bar,tooltip = "text") %>% 
    ggla_plotly_settings()    %>% 
    layout(title = list(text = paste0("<b>","Year-on-year percentage change in claimant count","</b>",
                                      "<br>",
                                      "<sup>",
                                      paste0("Latest data for period ", format(new_release_cc,"%B %Y")),
                                      "</sup>",
                                      "<br>"),
                        font = list(size = 22),
                        y = .95, xref = "plot"),
           xaxis = list(tickfont = list(size = 15)),
           yaxis = list(tickfont = list(size = 15)),
           legend = list(font = list(size = 15),title=list(text=""))) %>% 
    plotly_modebar
  
  
  claimbarcap <- paste0("Source: ONS Claimant Count by sex and age (NSA).","<br>","<br>", "Note: may include some employed claimants on low hours or earnings.")


```

```{r wfj_figures, echo=FALSE}

wfj_newmonth <- wfj_stats %>%  filter(new_date == max(new_date)) %>% filter(row_number()==1) %>% mutate(new_date = format(new_date, '%B %Y'))  %>% select(new_date)

wfj_lastmonth <- wfj_stats %>% filter(new_date == (ymd(max(new_date)) %m+% - months(3))) %>% filter(row_number()==1) %>% mutate(new_date = format(new_date, '%B %Y'))  %>% select(new_date)

wfj_level <- value_form(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% mutate(obs_value=obs_value/1000000) %>% select(obs_value),s=signif_mil, d=digits_mil)


wfj_p_quart_uk <- perc_form(100*(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "United Kingdom" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd(max(new_date)) %m+% - months(3)) & geography_name == "United Kingdom" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value))/(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd(max(new_date)) %m+% - months(3)) & geography_name == "United Kingdom" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value)),d = digits_perc)


wfj_p_dec19_uk <- perc_form(100*(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "United Kingdom" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd("2019-12-01")) & geography_name == "United Kingdom" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value))/(wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == (ymd("2019-12-01")) & geography_name == "United Kingdom" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value)),d = digits_perc)

wfj_emp_d_dec19 <- value_form(wfj_stats %>% filter(item_name == "employee jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "employee jobs" & new_date == (ymd("2019-12-01")) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value),s = signif_thous)
  
wfj_emp_p_dec19 <- perc_form(100*(wfj_stats %>% filter(item_name == "employee jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "employee jobs" & new_date == (ymd("2019-12-01")) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value))/(wfj_stats %>% filter(item_name == "employee jobs" & new_date == (ymd("2019-12-01")) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value)),d = digits_perc)

wfj_self_d_dec19 <- value_form(wfj_stats %>% filter(item_name == "self-employment jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "self-employment jobs" & new_date == (ymd("2019-12-01")) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value),s = signif_thous)
  
wfj_self_p_dec19 <- perc_form(100*(wfj_stats %>% filter(item_name == "self-employment jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value) - wfj_stats %>% filter(item_name == "self-employment jobs" & new_date == (ymd("2019-12-01")) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value))/(wfj_stats %>% filter(item_name == "self-employment jobs" & new_date == (ymd("2019-12-01")) & geography_name == "London" & measures_name == "Value" & industry_name == "Total") %>% select(obs_value)),d = digits_perc)


```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

## Workforce Jobs {.tabset}

<br/>

### Number of workforce jobs

The Workforce Jobs series provides quarterly estimates of the number of jobs in the UK and is the ONS' preferred source of jobs broken down by region of workplace and industry.

The latest data for `r wfj_newmonth` show that:

-   The total number of workforce jobs in London was estimated at `r wfj_level` million, `r condi_text(wfj_p_quart,"noun")` of approximately `r abs2(wfj_p_quart)`% from `r wfj_lastmonth` against a UK average of `r wfj_p_quart_uk`%.

-   The latest estimate represents `r condi_text(wfj_d_dec19,"noun")` of approximately `r abs2(wfj_d_dec19)`, or `r abs2(wfj_p_dec19)`%, from December 2019 levels (the pre-pandemic peak in London) compared with `r condi_text(wfj_p_dec19_uk,"noun")` of `r abs2(wfj_p_dec19_uk)`% across the UK on average.

-   Total jobs includes both employee and self-employment jobs alongside other types of employment; since December 2019, the latest employee jobs estimate has `r condi_text(wfj_emp_d_dec19,"increase")` by `r abs2(wfj_emp_d_dec19)` (`r abs2(wfj_emp_p_dec19)`%) while the self-employment jobs estimate has `r condi_text(wfj_self_d_dec19,"increase")` by `r abs2(wfj_self_d_dec19)` (`r abs2(wfj_self_p_dec19)`%).

<a href="#top">Back to top</a>

```{r workforce jobs trend, echo=FALSE,fig.cap = wfjcap, out.width='100%'}
  
  
  pal <- gla_pal(gla_theme = "default", palette_type = "highlight", n = c(1, 1))
  theme_set(theme_gla(gla_theme = "default"))
  
  # Make the wfj stats wide to have employee and self-employee jobs in label.
  wfj_stats_wide <- wfj_stats %>% 
    filter( new_date >= "2010-03-01" & measures_name == "Value" & geography_name == "London" & industry_name == "Total" ) %>% 
    select(date,new_date,geography_name,industry_name,item_name,obs_value) %>% 
    pivot_wider(id_cols=c(date,new_date,geography_name,industry_name),names_from = item_name,values_from=obs_value) %>% 
    clean_names()
  
  wfj_time_chart <- wfj_stats_wide %>%
    ggplot(mapping = aes(x = new_date, y = total_workforce_jobs /1000,
                         colour = geography_name, group = geography_name,
                         text = paste( format(new_date,'%B %Y'), "\n",
                                       "Total jobs: ", value_form(total_workforce_jobs /1000,s=3),"k", "\n",
                                       "Employee jobs: ", value_form(employee_jobs /1000,s=3),"k", "\n",
                                       "Self-employed jobs: ", value_form(self_employment_jobs /1000,s=3),"k", "\n",
                                       sep = ""))) +
    ggla_line(aes(size= geography_name)) +
    scale_size_manual(values = c(4 * mm_to_pt, 2 * mm_to_pt)) +
    scale_colour_manual(values = pal) +
    ggla_highlight(filter_type = "end") +
    ggla_highlight(mapping = aes(label = paste0(format(total_workforce_jobs /1000, digits = 1, big.mark=","))),
                   geom = GeomGLATextHighlight, filter_type = "end",  size = 4.5,
                   position = position_nudge(y = c(200,0)))+
    coord_cartesian(clip = 'off') +
    scale_y_continuous(expand = c(0, 0), labels = dollar_format(prefix = "",
                                                                largest_with_cents = 1),
                       limits = c(4500,6500)) +
    scale_x_date( date_breaks = "18 months",
                  date_labels = "%b'%y",
                  expand = expansion( mult = c(0.1,0.02))) +
    theme(plot.margin = unit(c(1,1,1,1), "cm"))+
    labs(title =  "Total number of Workforce jobs in London",
         subtitle = paste0("Seasonally adjusted (thousands), latest data for ", wfj_stats$date_name[wfj_stats$new_date == max(wfj_stats$new_date)]),
         caption = "\nSource: ONS Workforce Jobs.\n\nNote: The margin of error for all jobs is +/- 0.9% for London and +/- 0.3% for the UK.")+
    theme(plot.caption = element_text(color = rgb(166,166,166,maxColorValue = 255)))
  
  ggsave( paste0(IMAGES,Sys.Date(),"_","wfj_jobs",".svg"), device = "svg", width = 8, height = 8, units = "in")
  ggsave( paste0(IMAGES,Sys.Date(),"_","wfj_jobs",".png"), device = "png", width = 8, height = 8, units = "in")
  
  
  #knitr::include_graphics(paste0(IMAGES,Sys.Date(),"_","wfj_jobs",".svg"))
  
  ggplotly(wfj_time_chart,tooltip = "text", dynamicticks=TRUE) %>% 
    ggla_plotly_settings()    %>% 
    layout(title = list(text = paste0("<b>","Total number of Workforce jobs in London","</b>",
                                      "<br>",
                                      "<sup>",
                                      paste0("Seasonally adjusted (thousands), latest data for ", wfj_stats$date_name[wfj_stats$new_date == max(wfj_stats$new_date)]),
                                      "</sup>",
                                      "<br>"),
                        font = list(size = 22),
                        y = .95, xref = "plot"),
           xaxis = list(tickfont = list(size = 15)),
           yaxis = list(tickfont = list(size = 15)),
           legend = list(font = list(size = 15),title=list(text=""))) %>% 
    plotly_modebar
  
  
  wfjcap <- paste0("Source: ONS Workforce Jobs.","<br>","<br>", "Note: The margin of error for all jobs is +/- 0.9% for London and +/- 0.3% for the UK.")


```

```{r,echo=FALSE}
# Find industries with lowest and highest growth in recent data
  
  sector_thresh <- 25000
  
  lowest_growth <- wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & obs_value>=sector_thresh & !is.na(perc_change_dec19_within_date_rank)) %>% slice_max(perc_change_dec19_within_date_rank,n=3) 
  
  highest_growth <- wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & obs_value>=sector_thresh & !is.na(perc_change_dec19_within_date_rank)) %>% slice_min(perc_change_dec19_within_date_rank,n=3) 
  
  lowest_growth_q <- wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & obs_value>=sector_thresh & !is.na(perc_change_quart_within_date_rank)) %>% slice_max(perc_change_quart_within_date_rank,n=3) 
  
  highest_growth_q <- wfj_stats %>% filter(item_name == "total workforce jobs" & new_date == max(new_date) & geography_name == "London" & measures_name == "Value" & obs_value>=sector_thresh & !is.na(perc_change_quart_within_date_rank)) %>% slice_min(perc_change_quart_within_date_rank,n=3) 
  
  ## Find industry names and growth figures
  ## by year
  
  low_1_name <- lowest_growth %>% filter(row_number()==1) %>% select(industry_name_simple)
    
  low_1_d <- value_form(lowest_growth %>% filter(row_number()==1) %>% select(change_since_dec19),s = signif_thous)
    
  low_1_p <- perc_form(lowest_growth %>% filter(row_number()==1) %>% select(change_perc_since_dec19),d = digits_perc)
  
  low_2_name <- lowest_growth %>% filter(row_number()==2) %>% select(industry_name_simple)
    
  low_2_d <- value_form(lowest_growth %>% filter(row_number()==2) %>% select(change_since_dec19),s = signif_thous)
    
  low_2_p <- perc_form(lowest_growth %>% filter(row_number()==2) %>% select(change_perc_since_dec19),d = digits_perc)
  
  low_3_name <- lowest_growth %>% filter(row_number()==3) %>% select(industry_name_simple)
    
  low_3_d <- value_form(lowest_growth %>% filter(row_number()==3) %>% select(change_since_dec19),s = signif_thous)
    
  low_3_p <- perc_form(lowest_growth %>% filter(row_number()==3) %>% select(change_perc_since_dec19),d = digits_perc)
  
  
  high_1_name <- highest_growth %>% filter(row_number()==1) %>% select(industry_name_simple)
    
  high_1_d <- value_form(highest_growth %>% filter(row_number()==1) %>% select(change_since_dec19),s = signif_thous)
    
  high_1_p <- perc_form(highest_growth %>% filter(row_number()==1) %>% select(change_perc_since_dec19),d = digits_perc)
  
  high_2_name <- highest_growth %>% filter(row_number()==2) %>% select(industry_name_simple)
    
  high_2_d <- value_form(highest_growth %>% filter(row_number()==2) %>% select(change_since_dec19),s = signif_thous)
    
  high_2_p <- perc_form(highest_growth %>% filter(row_number()==2) %>% select(change_perc_since_dec19),d = digits_perc)
  
  high_3_name <- highest_growth %>% filter(row_number()==3) %>% select(industry_name_simple)
    
  high_3_d <- value_form(highest_growth %>% filter(row_number()==3) %>% select(change_since_dec19),s = signif_thous)
    
  high_3_p <- perc_form(highest_growth %>% filter(row_number()==3) %>% select(change_perc_since_dec19),d = digits_perc)

```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

### Workforce jobs by sector

Between December 2019 and `r wfj_newmonth`\*:

-   `r low_1_name` recorded the largest percentage decrease in jobs in London (`r low_1_d` or `r low_1_p`%); this was followed by `r low_2_name` (`r low_2_d` or `r low_2_p`%) and `r low_3_name` (`r low_3_d` or `r low_3_p`%).

-   The following industry groups shown in the chart saw the largest positive percentage change in jobs in London:

    -   `r high_1_name` (`r high_1_d` or `r high_1_p`%)
    -   `r high_2_name` (`r high_2_d` or `r high_2_p`%)
    -   `r high_3_name` (`r high_3_d` or `r high_3_p`%)

<a href="#top">Back to top</a>

<foot>\*Note: excluding sectors with less than `r value_form(sector_thresh,s = signif_thous)` jobs in London. </foot>

```{r wfj sector, echo=FALSE,fig.cap = wfjseccap, out.width='100%'}
  
  pal <- gla_pal(gla_theme = "default", palette_type = "highlight", n = c(1, 1))
  theme_set(theme_gla(gla_theme = "default"))
  
  # TODO
  
  # Select industries large in London, e.g. 20k+
  lon_large_inds <- wfj_stats %>%
    filter( item_name == "total workforce jobs" & new_date == max(new_date) & measures_name == "Value" & obs_value>=sector_thresh 
            & geography_name =="London") %>% 
    select(industry_name_simple)
  
  lon_large_inds <- unlist(lon_large_inds)
  
  wjf_industry_bar <- wfj_stats %>%
    filter( item_name == "total workforce jobs" & new_date == max(new_date) & measures_name == "Value" & industry_name_simple %in% lon_large_inds
            & geography_name %in% c("London", "United Kingdom")) %>% 
    group_by(geography_name) %>% #To ensure chart ranking is correct, make rank based on London and assign within sector
    mutate(chart_ranking = case_when( 
      industry_name_simple == "Total" & geography_name=="London" ~ 1,
      !(industry_name_simple == "Total") & geography_name=="London" ~ dense_rank(desc(change_perc_since_dec19))+1,
      TRUE ~ 99))  %>% 
    group_by(industry_name_simple) %>% 
    mutate(chart_ranking=min(chart_ranking)) %>% 
    ungroup() %>% 
    arrange(chart_ranking) %>% 
    ggplot(mapping = aes(x =  factor(reorder(industry_name_simple, -chart_ranking)), 
                         y = change_perc_since_dec19, 
                         colour = factor(reorder(geography_name,desc(geography_name))) ,#since horizontal bar reverses orders, we need to reverse too
                         fill=factor(reorder(geography_name,desc(geography_name))),
                         text = paste(industry_name_simple, "\n",
                                      geography_name, "\n",
                                      "Change: ", perc_form(change_perc_since_dec19),"%", "\n",
                                      sep = ""))) +
    geom_bar(stat = "identity", position = position_dodge(), width=0.5) +
    geom_hline(aes(yintercept=0), colour="gray45") +
    geom_vline(xintercept=c(length(lon_large_inds) - 0.5),colour="gray45")+ #adds line below total
    scale_color_manual(values = rev(pal), aesthetics = "colour")+
    scale_fill_manual(values = rev(pal), aesthetics = "fill")+
    theme_set(theme_gla(gla_theme = "default", y_label_length=100)) + #GLA theme and removes lines below y-axis labels
    scale_y_continuous(limits = c(-18, 12), labels = dollar_format(prefix = "", 
                                                                   suffix = "%", 
                                                                   largest_with_cents = 1)) +
    theme(plot.margin = unit(c(1,1,1,1), "cm"))+
    scale_x_discrete(expand = c(0,0)) +
       theme(panel.grid.major.y = element_blank(), #removed y grid
        panel.grid.minor.y = element_blank(), # removes small y grid
        panel.grid.major.x = element_line( size=.5 ), # removes x grid
        axis.text.y = ggplot2::element_text( # ensures  y axis labels stay outside chart
              hjust = 0, vjust = 0.5,
              margin = ggplot2::margin(t = 0, r = 0, b = 0, l = 0,
                                       unit = "pt")),
        axis.ticks.length.y = ggplot2::unit(x = 0, units = "pt")) + #removes ticks from y axis
    coord_flip()+
    guides(colour=guide_legend(reverse=TRUE),
           fill=guide_legend(reverse=TRUE)) +
    labs(title = paste0("Percentage change in Workforce Jobs by industry group"),
         subtitle = paste0("Seasonally adjusted, between December 2019 and ", wfj_stats$date_name[wfj_stats$new_date == max(wfj_stats$new_date)]),
         caption = "\nSource: ONS Workforce Jobs.\n\nNote: The margin of error for all jobs is +/- 0.9% for London and +/- 0.3% for the UK.")+
    theme(plot.caption = element_text(color = rgb(166,166,166,maxColorValue = 255)))
  
  ggsave(paste0(IMAGES,Sys.Date(),"_","wfj_industries",".svg"), device = "svg", width = 8, height = 8, units = "in")
  ggsave(paste0(IMAGES,Sys.Date(),"_","wfj_industries",".png"), device = "png", width = 8, height = 8, units = "in")
  
  
  #knitr::include_graphics(paste0(IMAGES,Sys.Date(),"_","wfj_industries",".svg"))
  
  ggplotly(wjf_industry_bar,tooltip = "text") %>% 
    ggla_plotly_settings()    %>% 
    layout(title = list(text = paste0("<b>","Percentage change in Workforce Jobs by industry group","</b>",
                                      "<br>",
                                      "<sup>",
                                      paste0("Seasonally adjusted, between December 2019 and ", wfj_stats$date_name[wfj_stats$new_date == max(wfj_stats$new_date)]),
                                      "</sup>",
                                      "<br>"),
                        font = list(size = 22),
                        y = .95, xref = "plot"),
           xaxis = list(tickfont = list(size = 15), showgrid=TRUE),
           yaxis = list(tickfont = list(size = 15), showgrid=FALSE),
           legend = list(font = list(size = 15),title=list(text="")),
           margin = list(l = 160)) %>% 
    reverse_legend_labels() %>% 
    plotly_modebar
  
  
  wfjseccap <- paste0("Source: ONS Workforce Jobs.","<br>","<br>", "Note: The margin of error for all jobs is +/- 0.9% for London and +/- 0.3% for the UK.")

```

<hr  width="100%" style="background-color: rgb(134,139,142);height: 1.0px;margin: 0;"/>

<br/>

## Summary of headline indicators

<br/>

```{r flextable indicators, echo = FALSE, ft.align="left", out.width='100%'}
  
  #-------------------------------------------------------------------------------
  #  Create the headline stats table as data frame
  #-------------------------------------------------------------------------------
  
  
  condi_arrow <- function(value) {
    arrow <- case_when( value < 0 ~ "u",
                        value == 0 ~ "v",
                        value > 0 ~ "t")
    return(arrow)
  }
  
  # Create rows to use as subheaders in table
  sub_head_rows <- tibble("geography_name"=c("Employment","Unemployment","Economic Inactivity"),"indicator"=c("emp_","unem_","inact_"),tab_rank=1)
  
  headline_rates <- lfs_stats %>% 
    filter(sex_name == "Total" & geography_name %in% c("London", "United Kingdom") & date == max(date)) %>% 
    mutate(measures_name = case_when(measures == 20100 ~ "num",
                                     measures == 20207 ~ "rate"),
           val_type = case_when(value_type_name == "Level" ~ "le",
                                value_type_name == "Change on quarter" ~ "cq",
                                value_type_name == "Change on year" ~ "cy"),
           geography_name = case_when(geography_name == "United Kingdom" ~ "UK" ,
                                      TRUE ~ "London")) %>% 
    select("date_name", "geography_name","measures_name","val_type","total_in_employment_aged_16_and_over","total_unemployed_aged_16_and_over",
           "total_economically_inactive_aged_16_and_over","total_in_employment_aged_16_to_64","total_economically_inactive_aged_16_to_64") %>% 
    pivot_longer(names_to = "obs_type_help", values_to = "obs_value", cols = "total_in_employment_aged_16_and_over":"total_economically_inactive_aged_16_to_64") %>% 
    mutate(obs_type = case_when(obs_type_help == "total_in_employment_aged_16_and_over" ~ "emp_16_plus",
                                obs_type_help == "total_unemployed_aged_16_and_over" ~ "unem_16_plus",
                                obs_type_help == "total_economically_inactive_aged_16_and_over" ~ "inact_16_plus",
                                obs_type_help == "total_in_employment_aged_16_to_64" ~ "emp_16_64",
                                obs_type_help == "total_economically_inactive_aged_16_to_64" ~ "inact_16_64")) %>%
    filter((measures_name=="num" & obs_type %in% c("emp_16_plus","unem_16_plus","inact_16_64")) | 
             (measures_name=="rate" & obs_type %in% c("emp_16_64","unem_16_plus","inact_16_64"))) %>% #Should make the right combination for table
    mutate(indicator =  str_extract(obs_type,".*?_")) %>% 
    mutate(col_name = paste(measures_name,val_type, sep = "_")) %>% 
    select("date_name","geography_name","obs_value","indicator","col_name") %>% 
    pivot_wider(names_from = "col_name", values_from = "obs_value") %>% 
    mutate(num_cq2 = condi_arrow(num_cq),
           num_cy2 = condi_arrow(num_cy),
           rate_cq2 = condi_arrow(rate_cq),
           rate_cy2 = condi_arrow(rate_cy),
           num_le = round(num_le/1000,0),
           num_cq = round(num_cq/1000,0),
           num_cy = round(num_cy/1000,0),
           tab_rank = case_when(geography_name == "London" ~ 2,
                                geography_name == "UK" ~ 3)) %>% 
    bind_rows(sub_head_rows) %>% 
    arrange(match(indicator,c("emp_","unem_","inact_")), tab_rank) %>% # Row order
    select("date_name","geography_name","indicator","num_le","num_cq","num_cq2","num_cy","num_cy2","rate_le","rate_cq","rate_cq2","rate_cy","rate_cy2") 
  
  #-------------------------------------------------------------------------------
  # Format table appearance manually
  #-------------------------------------------------------------------------------
  
  std_border = fp_border(width = 1)
  no_border = fp_border(width = 0)
  
  headline_rates_table <-  flextable(headline_rates, 
                                     col_keys=c("geography_name","num_le","num_cq","num_cq2","num_cy","num_cy2","rate_le","rate_cq","rate_cq2","rate_cy","rate_cy2")) %>% 
    colformat_num(
      big.mark = ",", decimal.mark = ".", digits=0) %>% 
    font(fontname='Arial', part = "all") %>% 
    # Setting colors and font of change variables to make indicator arrows
    font(j = c("num_cq2", "num_cy2", "rate_cq2", "rate_cy2"), fontname ='Marlett') %>% 
    color( i= ~ indicator %in% "emp_" & num_cq2 %in% "t", j = ~ num_cq2, color = "green") %>%  # Specifically for employment
    color( i= ~ indicator %in% "emp_" & num_cq2 %in% "u", j = ~ num_cq2, color = "red") %>% 
    color( i= ~ indicator %in% "emp_" & num_cy2 %in% "t", j = ~ num_cy2, color = "green") %>% 
    color( i= ~ indicator %in% "emp_" & num_cy2 %in% "u", j = ~ num_cy2, color = "red") %>% 
    color( i= ~ indicator %in% "emp_" & rate_cq2 %in% "t", j = ~ rate_cq2, color = "green") %>% 
    color( i= ~ indicator %in% "emp_" & rate_cq2 %in% "u", j = ~ rate_cq2, color = "red") %>% 
    color( i= ~ indicator %in% "emp_" & rate_cy2 %in% "t", j = ~ rate_cy2, color = "green") %>% 
    color( i= ~ indicator %in% "emp_" & rate_cy2 %in% "u", j = ~ rate_cy2, color = "red") %>%  
    color( i= ~ !(indicator %in% "emp_") & num_cq2 %in% "t", j = ~ num_cq2, color = "red") %>% # For the others
    color( i= ~ !(indicator %in% "emp_") & num_cq2 %in% "u", j = ~ num_cq2, color = "green") %>% 
    color( i= ~ !(indicator %in% "emp_") & num_cy2 %in% "t", j = ~ num_cy2, color = "red") %>% 
    color( i= ~ !(indicator %in% "emp_") & num_cy2 %in% "u", j = ~ num_cy2, color = "green") %>% 
    color( i= ~ !(indicator %in% "emp_") & rate_cq2 %in% "t", j = ~ rate_cq2, color = "red") %>% 
    color( i= ~ !(indicator %in% "emp_") & rate_cq2 %in% "u", j = ~ rate_cq2, color = "green") %>% 
    color( i= ~ !(indicator %in% "emp_") & rate_cy2 %in% "t", j = ~ rate_cy2, color = "red") %>% 
    color( i= ~ !(indicator %in% "emp_") & rate_cy2 %in% "u", j = ~ rate_cy2, color = "green") %>% 
    color( i= ~ num_cq2 %in% "v", j = ~ num_cq2, color = "orange") %>% # If equal
    color( i= ~ num_cy2 %in% "v", j = ~ num_cy2, color = "orange") %>% 
    color( i= ~ rate_cq2 %in% "v", j = ~ rate_cq2, color = "orange") %>% 
    color( i= ~ rate_cy2 %in% "v", j = ~ rate_cy2, color = "orange") %>% 
    #color( i= ~ geography_name %in% "London", j = ~ geography_name, color = rgb(109,167,222, maxColorValue = 255)) %>% 
    #color( i= ~ geography_name %in% "UK", j = ~ geography_name, color = rgb(204,204,204, maxColorValue = 255)) %>% 
    bold(i= ~ geography_name %in% c("Employment","Unemployment","Economic Inactivity"), j = ~ geography_name) %>% 
    add_header_row(colwidths = c(1 ,5, 5),
                   values = c(paste0("Headline estimates for ",headline_rates %>% filter(!is.na(date_name)) %>% filter(row_number()==1)  %>% select(date_name)), "Number* (thousands)","Rate** (%)")) %>% 
    align(part = "header", align = "center") %>% 
    set_header_labels( "geography_name"="","num_le"="Latest Estimate","num_cq"="Change on Quarter","num_cq2"="","num_cy"="Change on Year","num_cy2"="","rate_le"="Latest Estimate","rate_cq"="Change on Quarter","rate_cq2"="","rate_cy"="Change on Year","rate_cy2"="") %>% 
    bold(i=1, part = "header") %>% 
    add_footer_lines("Source: ONS Labour Force Survey. \n \nNotes: All figures are seasonally adjusted. *Numbers are based on total population (16+, male and female), except for inactivity which is 16-64. **Rates are based on working age population (16–64, male and female), except for the unemployment rate which is age 16+.") %>% 
    color( part = "footer", color = rgb(166,166,166,maxColorValue = 255)) %>%
    fontsize(size = 14, part = "all") %>% 
    # Cell widths
    width(j=1, width=6, unit="cm") %>% 
    width(j= ~ num_cq2 + num_cy2 + rate_cq2 + rate_cy2, width=1, unit="cm") %>% 
    # Merging of header cells
    merge_at(i=2,j=3:4,part="header") %>%  
    merge_at(i=2,j=5:6,part="header") %>%  
    merge_at(i=2,j=8:9,part="header") %>%  
    merge_at(i=2,j=10:11,part="header") %>%  
    merge_at(i=1:2,j=1,part="header") %>%  
    # Custom borders
    border_outer(part="header") %>% 
    border_outer(part="body") %>% 
    hline(i=1,part="header", border=no_border) %>% 
    hline(i= ~ geography_name %in% c("Employment","Unemployment","Economic Inactivity") ,part="body", border=std_border) %>% 
    hline(i= ~ geography_name %in% c("UK") ,part="body", border=std_border) %>% 
    vline(border = std_border, part = "all", j = ~ geography_name ) %>% 
    vline(border = std_border, part = "all", j = ~ num_cy2 ) %>% 
    fix_border_issues()
  
  knit_print(headline_rates_table)


```

<a href="#top">Back to top</a>

</font>

::: {.tocify-extend-page data-unique="tocify-extend-page" style="height: 0;"}
:::
